# Annexation

## Introduction

Consider the following expression:

"the boy's book"

This expression establishes a relation of _belonging_ between the two nouns: (i)\ "the boy", and (ii)\ "the book". It says that the book _belongs_ to the boy.

Arabic expresses this meaning using an construction called _annexation_. In this chapter we will learn about this construction.

## Forming the annexation

The word "annexation" is used when a new thing is added on to an existing base. We use the term "annexation" in Arabic grammar when one word is added on to another word by being placed right before it. Here is an example of an annexation:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=\txarb{ٱلْغُلَامِ} \txarb{كِتَابُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below=0.5ex] {\scriptsize base};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below=0.5ex] {\scriptsize annexe};

\path [decoration={text effects along path, 
  text={``the boy's book"},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate, below=4.5ex] (0,0);


%\draw [decoration={brace, mirror, raise=3ex}, decorate] 
%   (word-1.south west) -- (word-2.south east)
%      node [midway, below=3ex] {annexation};
\end{tikzpicture}
```
<!--"the boy's book"  -->

The annexation construction consists to two nouns:

1. The _annexe_ noun: This is the first noun in the annexation construction.
1. The _base_ noun: This is the second noun in the annexation construction.

The annexe [كِتَاب]{.ar} belongs to the base [ٱَلْغُلَام]{.ar}. 
You can use the alphabetical  order (A,\ B) to help you remember that the **a**nnexe noun comes before the **b**ase noun.

<!--
Take note of the following points regarding the annexation construction:

+ The base noun in an annexation shall always be in the i-state.
+ The state of the annexe noun is variable and shall depend on its function in the sentence.
+ The annexe noun in an annexation shall neither have an [n]{.trn}-mark nor the prefix [ٱَلْ]{.ar}.

## Remove

In Arabic we may chain two or more nouns in a sequence to establish an "of" relationship between them, in order to express, for example,

+ "the mother of the boy", or, equivalently" "the boy's mother".
+ "a house of a man", or, equivalently, "a man's house".
+ "the sword of Zayd", or equivalently, "Zayd's sword".

In English, as you can see there we can express this relationship using the word "of" or using apostrophe-s "'s" between two nouns. In some phrases, using apostrophe-s is more natural than using the word "of".

In Arabic we express these types of phrases by chaining the two nouns together thus:

[أُمُّ ٱلْغُلَامِ]{.ar}  
[Eummu -lgulAmi]{.trn}  
"the mother of the boy", or, equivalently" "the boy's mother"

We can identify noun-chains by noting that the first noun in the chain has neither an [n]{.trn} mark, nor an [ٱَلْ]{.ar}, and the second noun in the chain will be in the i-state. 

The first noun in the chain is the noun before "of" and the second noun in the chain is the noun after "of".
-->
## State of the annexe and base nouns

The base noun in an annexation is always in the i-state. The annexe noun may be in any state, depending on its function in the sentence. For example,
<!--
   + [أُمُّ ٱلْغُلَامِ مُعَلِّمَةٌ.]{.ar}  
     [Eummu -lgulAmi mueallimatun.]{.trn}  
     "The mother of the boy is a teacher.", or, equivalently, "The boy's mother is a teacher."
   + [وَجَدَ أُمَّ ٱلْغُلَامِ.]{.ar}  
     [wajada Eumma -lgulAmi.]{.trn}  
     "He found the boy's mother." (Note that [أُمَّ]{.ar} is in the a-state.)
   + [كَتَبَتْ إِلَىٰ أُمِّ ٱلْغُلَامِ.]{.ar}  
     [katabat EilA Eummi -lgulAmi.]{.trn}  
     "She wrote to the boy's mother." (Note that [أُمِّ]{.ar} is in the i-state.)
-->

[كِتَابُ ٱلْغُلَامِ ثَقِيلٌ.]{.ar}  
"The boy's book is heavy."  
(The annexe noun is in the u-state.)

[أَخَذَتِ ٱلْجَارِيَةُ كِتَابَ ٱلْغُلَامِ.]{.ar}  
"The girl took the boy's book."  
(The annexe noun is in the a-state.)

[كَتَبَ ٱلْمُعَلِّمُ فِي كِتَابِ ٱلْغُلَامِ.]{.ar}  
"The teacher~m~ wrote in the boy's book."  
(The annexe noun is in the i-state.)

## Definiteness of the annexation

Consider again the annexation expression we have been using so far:

[كِتَابُ ٱلْغُلَامِ]{.ar}  
"the boy's book"  

The base noun [ٱَلْغُلَام]{.ar} is definite because it is prefixed by [ٱَلْ]{.ar} "the".
Therefore we have translated it as "the boy".
The annexe noun [كِتَاب]{.ar} is not made definite by [ٱَلْ]{.ar}.
Nor is it made indefinite by an [n]{.trn}-mark. 
<!--Its definiteness is determined by the base noun. -->
Rather, it is defined by the base noun.
Because the base noun [ٱَلْغُلَام]{.ar} is definite, therefore the annexe noun [كِتَاب]{.ar} is also definite.
The entire annexation is definite.
<!--If the base noun is definite (as it is above) then the annexe noun shall also be definite.
And if the base noun is indefinite (as below) then the annexe noun shall also be indefinite.-->

Consider now the case when the base noun is indefinite.

[كِتَابُ غُلَامٍ]{.ar}  
"a boy's book"  

In the above example, the base noun [غُلَامٍ]{.ar} is indefinite because it has the [n]{.trn}-mark\ [◌ٍ]{.ar} and because it does not prefixed by [ٱَلْ]{.ar}. Therefore we have translated it as "a boy".
The annexe noun [كِتَاب]{.ar} has neither an [n]{.trn}-mark, nor the prefix [ٱَلْ]{.ar}.
It is defined by the base noun. Because the base noun [غُلَامٍ]{.ar} is indefinite, therefore the annexe noun [كِتَاب]{.ar} is also indefinite.
The entire annexation is indefinite.

We will see soon, if [#allAh]{.trn2} wills, why the definiteness of the annexe noun is important.

<!--

As you can see above English has two constructs to translate the annexation:

i. using "'s"
ii. using "of"

Sometimes both of the above constructs will fit for the translation, and sometimes only one will fit. For convenience, we will often translate with only one of them.

The annexe noun in an annexation can have neither an [n]{.trn} mark, nor an [ٱَلْ]{.ar}. 
The base noun is permitted to have an [n]{.trn} mark or an [ٱَلْ]{.ar} prefix. 
The definiteness of the base noun then determines the definiteness of the annexation. If the second noun is definite, then both the annexe and the base nouns be definite. And if the base noun is indefinite, then both the nouns in the chain will be indefinite. 

[بَيْتُ مُحَمَّدٍ]{.ar}  
[baytu muHammadin]{.trn}  
"the  house of [#muHammad]{.trn2}"

[بَيْتُ ٱلرَّجُلِ]{.ar}  
[baytu -rrajuli]{.trn}  
"The house of the man"

[بَيْتُ رَجُلٍ]{.ar}  
[baytu rajulin]{.trn}  
"a house of a man"
-->

Here are some examples of definite and indefinite annexations.

[لَبِسَ ٱلطِّفْلُ قَمِيصَ رَجُلٍ.]{.ar}  
"The child wore _a_ man's shirt."

[أَخَذَ أَمِيرُ ٱلْجَيْشِ رَايَةَ ٱلْمَلِكِ وَرَفَعَهَا.]{.ar}  
"_The_ army's commander took _the_ king's flag and raised it."

[جَلَسَ ٱلرَّجُلُ فِي ظِلِّ شَجَرَةٍ.]{.ar}  
"The man sat in _a_ tree's shade."

### Translating the annexation using "of"

So far we have been using the English "'s" to translate the Arabic annexation. Examples:

[بَيْتُ رَجُلٍ]{.ar}  
"a man's house"

[بَيْتُ ٱلرَّجُلِ]{.ar}  
"the man's house"

Instead of using "'s" we may use "of" as well. For example:

[بَيْتُ رَجُلٍ]{.ar}  
"a/the house of a man"

[بَيْتُ ٱلرَّجُلِ]{.ar}  
"a/the house of the man"

Note that the annexe noun "house" may be prefixed with either "a" or "the". This will depend on what is more natural in English. Often time both will fit. Here are some examples:

[لَبِسَ ٱلطِّفْلُ قَمِيصَ رَجُلٍ.]{.ar}  
"The child wore a/the shirt of a man."

[أَخَذَ أَمِيرُ ٱلْجَيْشِ رَايَةَ ٱلْمَلِكِ وَرَفَعَهَا.]{.ar}  
"_The_ commander of the army took _the_ flag of the king and raised it."

[جَلَسَ ٱلرَّجُلُ فِي ظِلِّ شَجَرَةٍ.]{.ar}  
"The man sat in _the_ shade of a tree."

[فَتَحَ ٱلِّصُّ شُبَّاكَ ٱلْبَيْتِ وَدَخَلَ ٱلْبَيْتَ.]{.ar}  
"The thief opened _a/the_ window of the house and entered the house."

It is important to understand that translating the annexe noun into English with "a" or "the" is purely for the reason of obtaining a natural translation. This does not affect whether or not the annexe noun is grammatically considered definite in Arabic.

As we mentioned earlier, the definiteness of the annexe noun in Arabic depends only on the definiteness of the base noun. 
If the base noun is definite then the annexe noun shall be considered definite as well.
And if the base noun is indefinite then the annexe noun shall be considered indefinite as well.

The need to maintain this distinction will become apparent in the next section.

If the base noun is definite, and it is desired to make the annexe noun grammatically indefinite, then it is necassary to break the annexation, and use a prepositional phrase instead, usually with the preposition [لِ]{.ar}, which, here, will mean "of". Example:

[ذَهَبَ ٱلْغُلَامُ إِلَىٰ بَيْتٍ لِلرَّجُلِ.]{.ar}  
"The boy went to a house of the man."

[فَتَحَ ٱلِّصُّ شُبَّاكًّا مِنَ ٱلْبَيْتِ وَدَخَلَ ٱلْبَيْتَ.]{.ar}  
"The thief opened a window of the house and entered the house."

## Broken plurals and [At]{.trn} sound plurals in annexations {#broken-plurals-and-at-sound-plurals-in-annexations}

There is no special rules for broken plurals and [At]{.trn} sound plurals in annexations. They behave just like singular nouns. Remember only that [At]{.trn} plurals end with [◌ٍ]{.trn} and [◌ِ]{.ar} in the a-state. Here are some examples:

[حَيَوَانَاتُ ٱلْغَابَةِ وَحْشَةٌ.]{.ar}  
"The animals of the forest are wild."

<!--
[كَتَبَتْ طَالِبَاتُ ٱلْمَدْرَسَةِ بِأَقْلَامِ مُعَلِّمَاتٍ.]{.ar}  
"The school's students~f~ wrote with teachers'~f~ pens."
-->

[قَرَأَتْ طَالِبَاتُ ٱلْمَدْرَسَةِ صَفَحَاتِ ٱلْكُتُبِ]{.ar}  
"The school's students~f~ read the pages of the books."

[فِي ٱلْخِزَانَةِ أَقْلَامُ مُعَلِّمَاتٍ.]{.ar}  
"In the cupboard are teachers'~f~ pens.

Contrary to broken plurals and [At]{.trn} plurals, duals and [Un]{.trn} sound plurals behave differently in annexations. We will deal with them in section\ \@ref(duals-and-sound-un-plurals-in-annexations)

## Describers in an annexation

### Describing the base noun

Consider the following expression:

[كِتَابُ ٱلْجَارِيَةِ]{.ar}  
"the girl's book"

Now say that we want to form an adjectival noun phrase "the small girl's book". 
Basically, we want to describe the base noun [ٱَلْجَارِيَة]{.ar} "the girl" with the adjectival noun [صَغِير]{.ar} "a small one". 
Here is how we will express this in Arabic:

[كِتَابُ ٱلْجَارِيَةِ ٱلصَّغِيرَةِ]{.ar}  
"the small girl's book"

In the manner we are already familiar with, we place the describer [صَغِير]{.ar} "a small one" after the describee
[ٱَلْجَارِيَة]{.ar} "the girl"
and match the describer with the describee in definiteness, state, gender and number (singular, dual, or plural).

Similarly, if we had an indefinite annexation, we would get:

[كِتَابُ جَارِيَةٍِ صَغِيرَةٍ]{.ar}  
"a small girl's book"

Here are some more examples:

[لَعِبَتِ ٱلْجَارِيَةُ فِي حَدِيقَةِ ٱلْبَيْتِ ٱلْكَبِيرِ.]{.ar}  
"The girl played in the garden of the big house."

[قَرَأَ ٱلْغُلَامُ سُورَةَ ٱلْقُرْآنِ ٱلْكَرِيمِ.]{.ar}  
"The boy read the [sUrah]{.trn2} of the Noble [#qurEAn]{.trn2}."

[جَلَسَ ٱلرَّجُلُ فِي ظِلِّ شَجَرَةٍ عَرِيضَةٍ وَسِيعَةٍ.]{.ar}  
"The man sat in the shade of a wide broad tree."

### Describing the annexe noun

Consider, again, the same annexation:

[كِتَابُ ٱلْجَارِيَةِ]{.ar}  
"the girl's book"

Say, now, that we want to describe the annexe noun [كِتَاب]{.ar} "book" with the adjectival noun [صَغِير]{.ar} "a small one". Normally, nothing can come between the annexe noun and the base noun in an annexation. So, the describer needs to be placed, again, after the base noun.
However, this time it will match the annexe noun, not the base noun, in state, definiteness, gender, and number. So we get:

[كِتَابُ ٱلْجَارِيَةِ ٱلصَّغِيرُ]{.ar}  
"the girl's small book"

Note how the describer 
[ٱَلصَّغِيرُ]{.ar} matches the annexe noun [كِتَابُ]{.ar} in state and gender.
Note also how the describer is definite with an [ٱَلْ]{.ar}. This is because it is matching the annexe noun [كِتَابُ]{.ar} in definiteness.
The annexe noun [كِتَاب]{.ar} is definite, not with [ٱَلْ]{.ar}, but rather because of the definite base noun [ٱَلْجَارِيَةِ]{.ar} "the girl".
We've already learned this rule in 
section\ \@ref(definiteness-of-the-annexation)
above.

Similarly, if we describe the annexe noun [كِتَاب]{.ar} in an indefinite annexation, we get:

[كِتَابُ جَارِيَةٍ صَغِيرٌ]{.ar}  
"a girl's small book"

This time
the describer 
[صَغِيرٌ]{.ar}
is indefinite with an [un]{.trn}-mark [◌ٌ]{.ar}. 
This is because 
the annexe noun [كِتَابُ]{.ar} is indefinite. It is indefinite because base noun [جَارِيَةٍ]{.ar} "a girl" is indefinite.

Now, you might be foreseeing a problem. What if the annexe noun and the base noun have the same gender, and the annexe too is in the i-state? For example, in the sentence:

[ذَهَبَ ٱلْغُلَامُ إِلَىٰ بَيْتِ ٱلرَّجُلِ ٱلْكَبِيرِ.]{.ar}  
"The boy went to the big/old man's house."  
or  
"The boy went to the man's big house."  

How do we know whether the describer [كَبِير]{.ar} is meant to describe the annexe noun [بَيْتِ]{.ar} or the base noun [ٱَلرَّجُل]{.ar}?
The annexe noun [بَيْتِ]{.ar} and the base noun [ٱَلرَّجُل]{.ar} are both masculine, singular, definite, and in the i-state.

The answer is that in such cases, context will have to be clear to tell us which of the two meanings is intended.
If the context makes it clear then there is no harm in using such a sentence for either of the two meanings.
<!--https://www.m-a-arabia.com/vb/showthread.php?t=40547
دفع اللبس في احتمالية النعت للمضاف والمضاف إليه-->

Also, sometimes, the meaning of the describer is such that it will likely apply to only one of the two nouns. For example,

[ذَهَبَ ٱلْغُلَامُ إِلَىٰ بَيْتِ ٱلرَّجُلِ ٱلْكَرِيمِ.]{.ar}  
"The boy went to a noble/generous man's house."  

In the sentence above the describer [كَرِيم]{.ar} "noble/generous" is likely to apply to a man, and not to a house.

If, however, the context is not clear, and the meaning of the describer can apply to both the annexe noun and the base noun, then the describer is likely to apply to the base noun and not to the annexe noun. So then, this interpretation is more likely:

[ذَهَبَ ٱلْغُلَامُ إِلَىٰ بَيْتِ ٱلرَّجُلِ ٱلْكَبِيرِ.]{.ar}  
"The boy went to the big/old man's house."  

In order to apply a describer to the annexe noun in such a case, it is better to break the annexation and form a prepositional phrase instead, usually with the preposition [لِ]{.ar}, which, here, will mean "of". Example:

[ذَهَبَ ٱلْغُلَامُ إِلَىٰ ٱلْبَيْتِ ٱلْكَبِيرِ لِلرَّجُلِ .]{.ar}  
"The boy went to the big house of the man."  

Here are some more examples:

[لَعِبَتِ ٱلْجَارِيَةُ بِكُرَةِ ٱلْغُلَامِ ٱلْحَمرَاءِ.]{.ar}  
"The girl played with the boy's red ball."  
(Note that [حَمْرَاء]{.ar} feminine to match [كُرَة]{.ar}.)

[سَقَطَتْ وَرَقَةُ ٱلشَّجَرَةِ ٱلْخَضْرَاءُ عَلَىٰ مَاءِ ٱلنَّعْرِ ٱلْعَرِيضِ.]{.ar}  
"The green leaf of the tree fell on the water of the broad river."  
(Note that [خَضْرَاء]{.ar} is in the u-state to match [وَرَقَة]{.ar})

[حَمَلَ ٱلْغُلَامُ حَقِيبَةَ ٱلْمَدْرَسَةِ ٱلثَّقِيلَةَ.]{.ar}  
"The boy carried the heavy school-bag."  
(literally: the heavy bag of the school).

[كَتَبَ ٱلرَّجُلُ عَلَىٰ صَفْحَةِ كِتَابٍ بَيْضَاءَ.]{.ar}  
"The man wrote on the white page of a book."  
(Note that [بَيْضَاءَ]{.ar} is feminine to match [صَفْحَة]{.ar}. However, also note that it has an a-mark [◌َ]{.ar} in the i-state because it is semi-flexible.)

<!--
[ذَهَبَ ٱلطَّالِبُ إِلَىٰ مَدْرَسَةِ مَدِينَةٍ عَظِيمَةٍ.]{.ar}
"The student went to a great city's school."  
or  
"The student went to a great school of a city."  
-->

## Semi-flexible nouns in an annexation

Remember that semi-flexible nouns don't take [n]{.trn}-marks and that when indefinite, the i-state is indicated by an [a]{.trn}-mark [◌َ]{.ar}. 
But when definite with [ٱَلْ]{.ar} then they behave just like fully-flexible nouns.
Example of the semi-flexible noun 
[صَحْرَاء^2^]{.ar} "a desert":

State  | Indefinite | Definite
:------|:-----------|:---------
u-state| [صَحْرَاءُ]{.ar} | [ٱَلصَّحْرَاءُ]{.ar}
a-state| [صَحْرَاءَ]{.ar} | [ٱَلصَّحْرَاءَ]{.ar}
i-state| [صَحْرَاءَ]{.ar} | [ٱَلصَّحْرَاءِ]{.ar}

We will now see how semi-flexible nouns behave in an annexation.

### A semi-flexible noun as the base noun

Here are examples of the semi-flexible noun [صَحْرَاء^2^]{.ar} "a desert" as the base noun in an annexation:

[ٱَلْقَرْيَةُ فِي وَسَطِ ٱلصَّحْرَاءِ.]{.ar}  
"The village is in the middle of the desert."

[شَرِبَ ٱلْأَعْرَابِيُّ مَاءً مِنْ بِئْرِ صَحْرَاءَ.]{.ar}  
"The bedouin drank some water from a desert's well."

As you can see, when 
[صَحْرَاء^2^]{.ar}
is definite, then its i-state is indicate by an [i]{.trn}-mark [◌ِ]{.ar}, just like fully-flexible nouns. 
However, when it is indefinite, then 
its i-state is indicate by an [a]{.trn}-mark [◌َ]{.ar}.

This is consistent with the general behavior of semi-flexible nouns that we are familiar with.

### A semi-flexible noun as the annexe noun

Contrary from expected behavior, a semi-flexible annexe noun, even when indefinite, takes an [i]{.trn}-mark\ [◌ِ]{.ar} in the i-state instead of an [a]{.trn}-mark\ [◌َ]{.ar}. Example,

[قَدِمَ ٱلْأَعْرَابِيُّ مِنْ صَحْرَاءِ أَرْضٍ بَعِيدَةٍ.]{.ar}  
"The bedouin came from the desert of a far land."

In the above example,
[صَحْرَاء^2^]{.ar} "a desert" is indefinite because it is the annexe noun to an indefinite base noun [أَرْض]{.ar} "a land". 
It is in the i-state because it is preceded by the preposition [مِنْ]{.ar} "from". 
Nevertheless, it takes an [i]{.trn}-mark [صَحْرَاءِ أَرْضٍ]{.ar}, not an [a]{.trn}-mark, which would be incorrect: $\times$\ [صَحْرَاءَ أَرْضٍ]{.ar}.

## Annexations with more than two nouns

So far we have seen annexations with two nouns. Annexations may be arbitrarily long. Here is an example of a noun-chain with more than two nouns:

<!--
[مِفْتَاحُ بَابِ ٱلْبَيْتِ]{.ar}  
"the house's door's key"
-->

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=\txarb{ٱلْبَيْتِ} \txarb{بَابِ} \txarb{مِفْتَاحُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-3.south west) -- (word-3.south east)
      node [midway, below=0.5ex] {\scriptsize $a_1$};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below=0.5ex] {\scriptsize $b_1$};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below=2.75ex] {\scriptsize $a_2$};

\draw [decoration={brace, mirror}, decorate, below=3ex] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below=2.75ex] {\scriptsize $b_2$};


\path [decoration={text effects along path, 
  text={``the house's door's key"},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate, below=6.75ex] (0,0);


%\draw [decoration={brace, mirror, raise=3ex}, decorate] 
%   (word-1.south west) -- (word-2.south east)
%      node [midway, below=3ex] {annexation};
\end{tikzpicture}
```

The above annexation consists of three nouns. It may be divided into two sub-annexations:

i. [مِفْتَاحُ بَابِ]{.ar} "door's key". Its annexe noun $a_1$\ is [مِفْتَاح]{.ar} and its base noun\ $b_1$ is [بَابِ]{.ar}.
i. [بَابِ ٱلْبَيْتِ]{.ar} "the house's door". Its annexe noun $a_2$\ is [بَابِ]{.ar} and its base noun\ $b_2$ is [ٱلْبَيْتِ]{.ar}.

The noun [بَاب]{.ar} "door" is common to both sub-annexations. It is the base noun of the first sub-annexation
[مِفْتَاحُ بَابِ]{.ar} "door's key".
At the same time, it is also the annexe noun of the second sub-annexation
[بَابِ ٱلْبَيْتِ]{.ar} "the house's door".

Only the final base noun may have [ٱَلْ]{.ar} or an [n]{.trn} mark. If the final base noun has [ٱَلْ]{.ar} (as above) then all the nouns in the annexation are definite. 

And if the final base noun is indefinite, as in the example below, then all the nouns in the annexation are indefinite.

[مِفْتَاحُ بَابِ بَيْتٍ]{.ar}  
"a house's door's key"

All the nouns except the first annexe noun must be in the i-state. Consistent with 
section\ \@ref(a-semi-flexible-noun-as-the-annexe-noun)
if a semi-flexible noun is any of the annexe nouns and is in the i-state, then its i-state is indicated by an [a]{.trn}-mark [◌َ]{.ar}. Example:

[مِنْ بِئْرِ صَحْرَاءِ أَرْضٍ]{.ar}  
"from the well of the desert of a land"

## Pronouns as base nouns

Consider the expression:

"his book"

This expression is very similar to the annexation:

[كِتَابُ ٱلْغُلَامِ]{.ar}  
"the boy's book" 

The difference is that we would like to replace the base noun [ٱَلْغُلَام]{.ar} "the boy" with the pronoun "his". For this we use the attached pronoun [هُ]{.ar}. When we place this pronoun as the base noun, we get:

[كِتَابُهُ]{.ar}  
"his book"

This annexation follows the same rules as the other annexations we have been studying so far:

+ The annexe noun may be in any state, depending on its function in the sentence.
+ The base noun is in the i-state. But because the base noun is a pronoun, and pronouns are rigid nouns (see section\ \@ref(rigidity-of-pronouns) that don't change their ending based on their state, therefore it's i-state will not be apparent.

Here are some examples of this annexation used in sentences:

[كِتَابُهُ ثَقِيلٌ.]{.ar}  
"His book is heavy"

[قَرَأَ ٱلرَّجُلُ كِتَابَهُ.]{.ar}  
"The man read his book."

[كَتَبَ ٱلْمُعَلِّمُ فِي كِتَابهِ.]{.ar}  
"The teacher~m~ wrote in his book."

## Duals and [Un]{.trn} sound plurals in annexations {#duals-and-sound-un-plurals-in-annexations}

We have already dealt with broken plurals and [At]{.trn} sound plurals in annexations in 
section\ \@ref(broken-plurals-and-at-sound-plurals-in-annexations).

In this section we will deal with 
duals and [Un]{.trn} sound plurals in annexations.

### Duals and [Un]{.trn} sound plurals as base nouns

As base noun, duals and [Un]{.trn} sound plurals behave no differently than other nouns. Being base nouns they will be in the i-state and this shall be indicated by:

i. [◌َيْنِ]{.ar} [-ayni]{.trn} for duals
i. [◌ِينَ]{.ar} [-Ina]{.trn} for [Un]{.trn} sound plurals

Here are some examples:

[لَجِئَ ٱلْمَظْلُومُنَ ٱلضُّعَفَاءُ فِي بِلَادِ ٱلْمُسْلِمِينَ ٱلْآمِنَةِ.]{.ar}  
"The weak wronged ones took refuge in the secure lands of the Muslims."

[أُخُتُ ٱلْغُلَامَيْنِ ٱلطَّوِيلَيْنِ صَغِيرَةِ.]{.ar}  
"The tall boys'~2~ sister is little."

[هِيَ طَالِبَةُ مُعَلِّمَتَيْنِ كَرِيمَتَيْنِ.]{.ar}  
"She is the student~f~ of noble teachers~2,f~."

### Duals and [Un]{.trn} sound plurals as annexe nouns

When duals and [Un]{.trn} sound plurals are annexe nouns, then their final [ن]{.ar} is treated as an [n]{.trn}-mark and is, therefore, deleted before annexing them to a base noun. For example:

[بَيْتَا ٱلرَّجُلِ]{.ar}  
"the man's houses~2~"  
not  
$\times$\ [بَيْتَانِ ٱلرَّجُلِ]{.ar}

Note, also, that because the base noun [ٱَلرَّجُلِ]{.ar} begins with a connecting [hamzah]{.trn} [ٱ]{.ar}, therefore the long vowel [A]{.trn} at the end of [بَيْتَا]{.trn} is pronounced as a short vowel [a]{.trn}, thus:

[bayta -rrajuli]{.trn}  
not  
$\times$\ [baytA -rrajuli]{.trn}

If the dual annexe noun were in the i-state then the final [ي]{.ar} gets an [i]{.trn}-mark [◌ِ]{.trn} if there is following connecting [hamzah]{.trn}. Example:




Compare this with describing the base noun with the same adjectival noun.

[كِتَابُ ٱلْجَارِيَةِ ٱلصَّغِيرَةِ]{.ar}  
"the girl's small book"

As you can see, the adjectival noun [صَغِير]{.ar} "a small one" is in the same location, but its gender and state are different in both expressions because they match a different describee. 
The definiteness, however, is the same in the two expressions because the annexe noun has the same definiteness as the base noun.



If the second noun in a chain is to be described using a qualitative noun, then just like we have learned, we will place the qualitative noun after the second noun in the chain and match it definiteness, state, gender, and number (depending on whether it denotes a rational or non-rational being). Examples: 

[بَيْتُ رَجُلٍ كَبِيرٍ]{.ar}  
[baytu rajulin kabIrin]{.trn}  
"a house of an old/big man" = "an old/big man's house"

[بُيُوتُ ٱلرِّجالِ ٱلْكِبَارِ]{.ar}  
[buyUtu -rrijAli -lkibAri]{.trn}  
"houses of the old/big men" = "the old/big men's houses"

If the first noun in a chain is to be described using a qualitative noun, then we will actually place the qualitative noun after the second noun, not the first noun, in the chain. This way we don't break the chain of the two nouns. The describer will match the first noun definiteness, state, gender, and number (depending on whether it denotes a rational or non-rational being). Examples: 

[بَيْتُ رَجُلٍ كَبِيرٌ]{.ar}  
[baytu rajulin kabIrun]{.trn}  
"a big house of a man" = "a man's big house"

[بُيُوتُ  ٱلْمَرْأَةِ ٱلْكَبِيرَةُ]{.ar}  
[buyUtu -lmarEati -lkabIratu]{.trn}  
"the big houses of the woman" = "the woman's big houses"

Consider, however, the sentence:

[ذَهَبَ إِلَىٰ بَيْتِ رَجُلٍ كَبِيرٍ.]{.ar}  
[pahaba EilA bayti rajulin kabIrin.]{.trn}  

In this sentence both the first noun in the chain [بَيْتِ]{.ar} [bayti]{.trn} and second noun [رَجُلٍ]{.ar} [rajulin]{.trn} are indefinite, singular, masculine, and in the i-state. So the describer [كَبِيرٍ]{.ar} [kabirin]{.trn} could be applicable to either. And the meaning of the sentence could be either,

"He went to a man's big house."

or,

"He went to a big/old man's house."

The listener would know from context which of the two meanings is intended. If the speaker wished to explicitly disambiguate the meanings without relying on context, he could use the preposition [لِ]{.ar} instead of using a noun-chain. Here the preposition [لِ]{.ar} gives the meaning of "of", rather than "for".

[ذَهَبَ إِلَىٰ بَيْتٍ كَبِيرٍ لِرَجُلٍ.]{.ar}  
[pahaba EilA baytin kabIrin lirajulin.]{.trn}  
"He went to a big house of a man." = "He went to a man's big house."

and,

[ذَهَبَ إِلَىٰ بَيْتٍ لِرَجُلٍ  كَبِيرٍ.]{.ar}  
[pahaba EilA baytin kabIrin lirajulin.]{.trn}  
"He went to a big/old man's house."

## Duals and plurals in noun chains

### Duals

Remember that duals are expressed by suffixing [ـَانِ]{.ar} [Ani]{.trn} to the noun when it is in the u-state and [ـَيْنِ]{.ar} [ayni]{.trn} when it is in the a-state and i-state.

If a dual is the second noun in a chain, then it is not treated any differently from any other noun. For example,

[بَيْتُ ٱلرَّجُلَيْنِ]{.ar}  
[baytu -rrajulayni]{.trn}  
"the house of the two men"

If a dual is the first noun in a chain, then the [ن]{.ar} in the dual suffix is dropped. For example,

+ [بَيْتَا ٱلرَّجُلِ وَاسِعَانِ.]{.ar}  
  [bayta -rrajuli wAsieAni.]{.trn}  
  "The two houses of the man are spacious."

  Note that [بَيْتَا ٱلرَّجُلِ ]{.ar} [bayta -rrajuli]{.trn} is technically pronounced identically to [بَيْتَ ٱلرَّجُلِ]{.ar} [bayta -rrajuli]{.trn}. The [Ealif]{.trn} in [بَيْتَا]{.ar} is not pronounced because of the connecting [hamzah]{.trn} after it. Some say that extra stress may be placed on the vowel [ta]{.trn} in order to convey that it is the dual: [bay#t#a -rrajuli]{.trn}.

+ [ذَهَبَ إِلَىٰ مَدِنَتَيْ بَلَدٍ.]{.ar}  
  [pahaba EilA madInatay baladin.]{.trn}  
  "He went to two cities of a country."
+ [فَتَحَ بَابَيِ ٱلْبَيْتِ.]{.ar}  
  [fataHa bAbayi -lbayti.]{.trn}  
  "He opened the two doors of the house."

  Note that a helper [i]{.trn} vowel is placed on the [ي]{.ar} in [بَابَيِ]{.ar} [bAbayi]{.trn} instead of the $\emptyset$-mark because of the connecting [hamzah]{.trn} after it.

### Plurals

We have learned three kinds of plurals:

1. The sound [Un]{.trn} plural.
2. The sound [At]{.trn} plural.
3. The broken plural.

The sound [Un]{.trn} plural behaves similarly to the dual in that when it is the first noun in a chain, the [ن]{.ar} is dropped. Examples:

+ [مُعَلِّمُو ٱلْغُلَامِ فِي ٱلْمَدْرَسَةِ.]{.ar}  
  [mueallimu -lgulAmi fi -lmadrasati.]{.trn}  
  "The teachers of the boy are in the school."

  Note, again, that [مُعَلِّمُو ٱلْغُلَامِ]{.ar} is technically pronounced identically to [مُعَلِّمُ ٱلْغُلَامِ]{.ar} [mueallimu -lgulAmi]{.trn}, except some say that extra stress may be placed on the vowel [مُ]{.ar} in the case of the plural [muealli#m#u -lgulAmi]{.trn}. This can serve to differentiate it from the singular when heard orally:

  [مُعَلِّمُ ٱلْغُلَامِ فِي ٱلْمَدْرَسَةِ.]{.ar}  
  [mueallimu -lgulAmi fi -lmadrasati.]{.trn}  
  "The teacher of the boy is in the school."

+ [وَجَدَ مُعَلِّمي غُلَامٍ.]{.ar}  
  [wajada mueallimI gulAmin.]{.trn}  
  "He found teachers of a boy."

+ [ذَهَبَ إِلَىٰ مُعَلِّمِي ٱلْغُلَامِ.]{.ar}  
  [pahaba EilA mueallimi -lgulAmi.]{.trn}  
  "He went to the teachers of the boy."

  Here too, some say to place extra stress on the vowel [مِ]{.ar} [muealli#m#i -lgulAmi]{.trn} in order to differentiate it to the listener from the singular:

+ [ذَهَبَ إِلَىٰ مُعَلِّمِ ٱلْغُلَامِ.]{.ar}  
  [pahaba EilA mueallimi -lgulAmi.]{.trn}  
  "He went to the teacher of the boy."

The other two types of plurals: the sound [At]{.trn} plural and the broken plural are not treated any differently from other nouns when used in chains. Examples:

+ [بُيُوتِ الرَّجُلِ وَاسِعَةٌ.]{.ar}  
  [buyUtu -rrajuli wAsieatun.]{.trn}  
  "The houses of the man are spacious."

+ [ذَهَبَتِ ٱلْمَعَلِّمَةُ إِلَىٰ طَالِبَاتِ ٱلْمَدْرَسَةِ.]{.ar}  
  [pahabati -lmueallimatu EilA TAlibAti -lmadrsati.]{.trn}  
  "The (female) teacher went to the (girl) students of the school."

## Non-fully changing nouns in noun-chains

Remember that a non-fully changing noun without [ٱَلْ]{.ar} does not take an [n]{.trn}-mark and its last letter keeps an [a]{.trn} mark instead of an [i]{.trn} mark when in the i-state. When definite with [ٱَلْ]{.ar}, it is similar to fully-changing nouns.

So when a non-fully changing noun without [ٱَلْ]{.ar} is the second noun in a noun-chain, it will have an [a]{.trn}-mark instead of an [i]{.trn} mark.

For example, [صَحْرَاءُ]{.ar} [SaHrAEu]{.trn} "a desert" is a non-fully-changing noun. "A desert's sky" will be expressed as:

[سَمَاءُ صَحْرَاءَ]{.ar}  
[samAEu SaHrAEa]{.trn}

not:

[سَمَاءُ صَحْرَاءِ]{.ar}  
[samAEu SaHrAEi]{.trn}

Similarly, [جَمِيلَةُ]{.ar} "[#jamIlah]{.trn2}" as a woman's name is a proper noun and is a non-fully-changing noun. "[#jamIlah]{.trn2}'s house" will be expressed as:

[بَيْتُ جَمِيلَةَ]{.ar}  
[baytu jamIlata]{.trn}  

not:

[بَيْتُ جَمِيلَةِ]{.ar}  
[baytu jamIlati]{.trn}  

However, as an exception, when an indefinite non-full-changing noun is the first noun in a noun-chain and is in the i-state, it will actually take an [i]{.trn}-mark on its last letter. For example, [صَحْرَاءُ]{.ar} [SaHrAEu]{.trn} "a desert" is a non-fully-changing noun.

"in the desert of [#makkah]{.trn2}" will be expressed as,

[فِي صَحْرَاءِ مَكَّةَ]{.ar}  
[fI SaHrAEi makkata]{.trn}  

As you can see, [صَحْرَاءِ]{.ar} [SaHrAEi]{.trn} has an [i]{.trn}-mark despite being a non-fully-changing noun, exemplifying the exceptional rule we have just stated. Also note, by the way, that [مَكَّةَ]{.ar} [makkata]{.trn} has an [a]{.trn}-mark on it's last letter in the i-state because it is a non-fully-changing noun without [ٱَلْ]{.ar} as the second noun in a noun-chain.

## Noun chains with more than two nouns

So far we have seen noun-chains with two nouns. Noun-chains can be arbitrarily long, although practically they will usually be less than five nouns long. Here is an example of a noun-chain with more than two nouns:

+ [مِفْتَاحُ بَابِ بَيْتِ ٱلرَّجُلِ]{.ar}  
  [miftAHu bAbi bayti -rrajuli]{.trn}  
  "the key of the door of the man's house"

+ [سَمَاءُ صَحْرَاءِ مَكَّةَ]{.ar}  
  [samAEu SaHrAEi makkata]{.trn}  
  "the sky of the desert of Makkah"

+ [كِتَابَا جَارِيَتَيْ مَدِينَةِ دَمِشْقَ]{.ar}  
  [kitAbA jAriyatay madInati damicqa]{.trn}  
  "the two books of the two girls of the city of Damascus"

+ [مِنْ وَرَقَتَيْ شَجَرَةِ حَدِيقَةٍ صَغِيرَتَيْنِ]{.ar}  
  [min waraqatay cajarati HadIqatin]{.trn}  
  "from two small leaves of a tree of a garden"

Note the following:

+ Only the last noun may have [ٱَلْ]{.ar} or an [n]{.trn} mark. 
+ The definiteness of the last noun determines the definiteness of the rest of the nouns in the chain. 
+ All the nouns except the first must be in the i-state. 
+ The [ن]{.ar} of the dual and the [Un]{.trn} plural is dropped for all nouns except the last.
+ Any non-fully-changing noun in the i-state in any position but the last will take an [i]{.trn}-mark on it's last letter. If a non-fully-changing noun is the last noun in the chain, and it does not have [ٱَلْ]{.ar} then it will take an [a]{.trn}-mark on it's last letter.
+ Describers for any of the nouns in the chain will appear at the end. Practically, only one of the nouns will have a describer.

## Mismatched definiteness in noun-chains

We have learned that if the second noun is definite (either by having [ٱَلْ]{.ar}, or by being a proper noun), then both the nouns in the chain will be definite. And if the second noun is indefinite, then both the nouns in the chain will be indefinite. 

What if we need to express an of-relationship where the two nouns are not both definite or both indefinite? For example, how then do we express phrases like "a house of the man" and "the house of a man"? 
In order to express the former, we can employ the preposition [لِ]{.ar}:

[ذَهَبَ إِلَىٰ بَيْتٍ لِلرَّجُلِ.]{.ar}  
[pahaba EilA baytin lirrajuli]{.trn}  
"He went to a house of the man."

Another way to express this is that we may interpret the phrase "a house of the man" to understand that the man has more than one house and that we are talking about one of them. In this case, we can use the plural to say,

[ذَهَبَ إِلَىٰ بَيْتٍ مِنْ بُيُوتِ ٱلرَّجُلِ]{.ar}  
[pahaba EilA baytin min buyUti -rrajuli.]{.trn}  
"He went to a house from the houses of the man."

This use of the plural, although fairly common, may be more appropriate in some cases than others, depending on the meaning of the sentence. 

The question remains: how to express a phrase of the type "the house of a man", where the first noun in definite and the second noun is indefinite?

In this case we will use the same expression as we did for "a house of a man":

[بَيْتُ رَجُلٍ]{.ar}  
[baytu rajulin]{.trn}  
"a house of a man"

Technically, "house" is indefinite in this expression, but it can still be used idiomatically to translate "the house of a man." It may help to use apostrophe-s in the translation where the definiteness of "house" is not apparent: "a man's house".

## Pronouns in a noun-chain

Consider the noun-chain:

[بَيْتُ ٱلرَّجُلِ]{.ar}  
[baytu -rrajuli]{.trn}  
"the house of the man" = "the man's house"

We can replace the noun "the man" with the attached pronoun [ـهُ]{.ar} [-hu]{.trn} "him" to get:

[بَيْتُهُ]{.ar}  
[baytuhu]{.trn}  
"the house of him" = "his house"

This can be done with all the attached pronouns:

|Arabic|English|
|:---|:----|
|[بَيْتُهُ]{.ar} [baytuhu]{.trn}    |  his house |
|[بَيْتُهَا]{.ar} [baytuhA]{.trn}   |  her house |
|[بَيْتُهُمَا]{.ar} [baytuhumA]{.trn}|  their~dual\ masc.~ house |
|[بَيْتُهُمَا]{.ar} [baytuhumA]{.trn}|  their~dual\ fem.~ house |
|[بَيْتُهُمْ]{.ar} [baytuhum]{.trn}  |  their~pl.\ masc.~ house |
|[بَيْتُهُنَّ]{.ar} [baytuhunna]{.trn}|  their~pl.\ fem.~ house |
|[بَيْتُكَ]{.ar} [baytuka]{.trn}    |  your~sing.\ masc.~ house |
|[بَيْتُكِ]{.ar} [baytuki]{.trn}    |  your~sing.\ fem.~ house |
|[بَيْتُكُمَا]{.ar} [baytukumA]{.trn}|  your~dual~ house |
|[بَيْتُكُمْ]{.ar} [baytukum]{.trn}  |  your~pl.\ masc.~ house |
|[بَيْتُكُنَّ]{.ar} [baytukunna]{.trn}|  your~pl.\ fem.~ house |
|[بَيْتِي]{.ar} [baytI]{.trn}      |  my house |
|[بَيْتُنَا]{.ar} [baytunA]{.trn}   |  our house |

Here are some notes regarding the use of pronouns in noun-chains.

+ An attached pronoun may be only the last noun in a noun-chain. It cannot be the first noun in a noun-chain.
+ A looped [tAE]{.trn} [ة]{.ar} in the first noun becomes an open [tAE]{.trn} [ت]{.ar} when a pronoun is attached as the second noun in the noun-chain. For example [مَدْرَسَةٌ]{.ar} [madrasatun]{.trn} "a school" becomes [مَدْرَسَتُهُ]{.ar} [madrasatuhu]{.trn} "his school".
+ The first noun in the noun-chain, before the attached pronoun, may be in any state. So, for example, we can have:
  + [دَخَلَ بَيْتَهُ.]{.ar} (a-state)  
    [daxala baytahu]{.trn}  
    "He entered his house."
  + [فِي بَيْتِهِ.]{.ar} (i-state)  
    [fI baytihi]{.trn}  
    "in his house"
+ With the pronoun [ـِي]{.ar} [-I]{.trn}, the first noun in the noun-chain will generally appear the same in all states. For example,
  + [بَيْتِي كَبِيرٌ]{.ar} (u-state)  
    [baytI kabIrun.]{.trn}  
    "My house is big."
  + [دَخَلَ بَيْتِي.]{.ar} (a-state)  
    [daxala baytI]{.trn}  
    "He entered my house."
  + [فِي بَيْتِي]{.ar} (i-state)  
    [fI baytI]{.trn}  
    "in my house"
+ We have mentioned that when a dual noun or sound [Un]{.trn} plural noun is first noun in a noun-chain its [ن]{.ar} is dropped. This rule remains valid when the last noun in the noun-chain is an attached pronoun. So we get:
  + [بَيْتَاهُ]{.ar}  
    [baytAhu]{.trn}  
    "his two houses"
  + [إِلَىٰ بَيْتَيْهِمَا]{.ar}  
    [EilA baytayhimA]{.trn}  
    "to their~dual~ two houses"
  + [مُعَلِّمُوكَ]{.ar}  
    [mueallimUka]{.trn}
    "your~sing.\ masc.~ (male) teachers"
  + [مِنْ مُعَلِّمِيهِ]{.ar}  
    [min mueallimIhi]{.trn}  
    "from his (male) teachers"

+ With dual nouns and sound [Un]{.trn} plural nouns, when the last noun in the noun-chain is [ـِي]{.ar} [-I]{.trn} "my", the [ـِي]{.ar} [-I]{.trn} becomes a [يَ]{.ar} [ya]{.trn} when preceded by an unvowelled [ا]{.ar} [Ealif]{.trn} or [ي]{.ar} [yAE]{.trn}. The [يَ]{.ar} gets doubled when preceded by an unvowelled [ي]{.ar}. Examples:
  + [مُعَلِّمَايَ طَيِّبَانِ.]{.ar}  
    [mueallimAya TayyibAni.]{.trn}  
    "My two (male) teachers are good."
  + [مِنْ مُعَلِّمَيَّ]{.ar}  
    [min mueallimayya]{.trn}  
    "from my two (male) teachers"
  + [إِلَىٰ مُعَلِّمِيَّ]{.ar}  
    [EilA mueallimiyya]{.trn}  
    "to my (male) teachers"
<!--There is an example of ya after unvoweled waw: fUya "my mouth" but famI and fiyya are more common so I won't mention this rule since I won't be providing an example-->

+ As a special rule, however, when the sound [Un]{.trn} plural is in the u-state and is the first-noun in a noun-chain followed by [ـِي]{.ar} [-I]{.trn} "my", the ending will not become [ـُويَ]{.ar} [-Uya]{.trn} as we might expect. Instead, it will become identical to the a-state and i-state [ـِيَّ.]{.ar} [-iyya]{.trn}. For example:
  + [مُعَلِّمِيَّ طَيِّبُونَ.]{.ar}  
    [mueallimiyya TayyibUna.]{.trn}  
    "My (male) teachers~pl.~ are good."

+ Pronouns are always definite. So when a pronoun is the last noun in a noun-chain, all preceding nouns will be definite. This is important to know because if a describer is to be used for any of the preceding nouns, it will be definite too and will therefore have [ٱَلْ]{.ar}. Examples:
  + [بَيْتِي ٱلْكَبِيرُ وَاسِعٌ.]{.ar}  
    [bayti -lkabIru wAsieun.]{.trn}  
    "My big house is spacious."
  + [قَرَءَا كِتَابَيْهِمَا ٱلْجَدِيدَيْنِ.]{.ar}  
    [qaraEA kitAbayhima -ljadIdayni.]{.trn}  
    "They~dual~ read their two new books."
  + [مُعَلِّمِيَّ ٱلْكِرَامُ فِي مَدْرَسَتِهِمُ ٱلْقَدِيمَةِ.]{.ar}  
    [mueallimiyya -lkirAmu fI madrasatihimu -lqadImati.]{.trn}  
    "My noble (male) teachers are in their old school."

  By the way, describers will obviously only be for nouns preceding the pronoun and never for the pronoun itself.

## Noun-chains with "and"

### Multiple nouns before "of" and one noun after "of"

In English we can have an expression like "the pen and the book of the boy" = "the boy's pen and book". In this sentence there are two nouns before "of" and one noun after "of".

In order to express this in Arabic, we will say:

[قَلَمُ ٱلْغُلَامِ وَكِتَابُهُ]{.ar}  
[qalamu -lgulAmi wa kitAbuhu]{.trn}  
"the boy's pen and his book" = "the boy's pen and book"

Note that the noun-chain is not broken by the insertion of [وَ]{.ar} [wa]{.trn} "and". Rather a second noun-chain is used and the two are separated by [وَ]{.ar} [wa]{.trn} "and". This is the preferred way of expressing such expressions.

There is another, less preferred way of expressing this. And this is by breaking the first noun-chain and inserting [وَ]{.ar} [wa]{.trn} "and":

[قَلَمُ وَكِتَابُ ٱلْغُلَامِ]{.ar}  
[qalamu wa kitAbu -lgulAmi]{.trn}  
"the boy's pen and book"

This second method is not considered as eloquent. Some even consider it incorrect. So we advise you to use the first method whenever possible.

#### With pronouns

If the second noun in the noun-chain is replaced with a pronoun then only the first method is allowed. For example,

[قَلَمُهُ وَكِتَابُهُ]{.ar}  
[qalamuhu wakitAbuhu]{.trn}  
"his pen and his book"

### One noun before "of" and multiple nouns after "of"

We can also have expressions like "the house of the boy and the girl". In this sentence there is one noun before "of" and two nouns after "of".

To express this in Arabic we will say:

[بَيْتُ ٱلْغُلَامِ وَٱلْجَارِيَةِ]{.ar}  
[baytu -lgulAmi wa-ljAriyati]{.trn}  
"the house of the boy and the girl"

Note that both 
[ٱلْغُلَامِ]{.ar} [EalgulAmi]{.trn} and [ٱَلْجَارِيَةِ]{.ar} [EaljAriyati]{.trn} are in the i-state because they are both second nouns in the noun-chain.

#### With pronouns

If any or both of the second nouns in the noun-chain is replaced with a pronoun then the first noun must be repeated. For example,

[بَيْتُهُ وَبَيْتُ ٱلْجَارِيَةِ]{.ar}  
[baytuhu wabaytu -ljAriyati]{.trn}  
"his house and the girl's house"

[بَيْتُهُ وَبَيْتُهَا]{.ar}  
[baytuhu wabaytuhA]{.trn}  
"his house and her house"

However, it must be said that these expressions are ambiguous because they could be interpreted to mean that there are two houses: one that is "his" and the second that is "her's". One way to remove this ambiguity, when only a single house is intended, is to use the dual pronoun instead:

[بَيْتُهُمَا]{.ar}  
[baytuhumA]{.trn}  
"their~dual~ house"

## Non-possessive noun-chains

piece of bread, man of wealth, rule of law, nahru -lEurdunn

## Exercises

1. [في البيت باب كبير و باب صغير. دخل الرجل من باب البيت الكبير.]{.ar}  

