# a-state incomplete-action verbs

## Introduction

In 
chapter\ \@ref(u-state-incomplete-action-verbs)
we mentioned that incomplete action verbs have three states (like nouns). 
These states are called:

i. The u-state
i. The a-state
i. The [0]{.txt}-state

We introduced the u-state incomplete-action verb in
chapter\ \@ref(u-state-incomplete-action-verbs).
In this chapter we will study the 
a-state
incomplete-action verb.

The u-state incomplete-action verb makes a plain statement.
The a-state incomplete-action verb implies a wish or purpose.
The a-state incomplete-action verb is used after the following articles:

+ [أَنْ]{.ar} [Ean]{.trn}
+ [لَنْ]{.ar} [lan]{.trn}
+ [لِ]{.ar} [li]{.trn}
+ [كَيْ]{.ar} [kay]{.trn}
+ [حَتَّىٰ]{.ar} [HattA]{.trn}
+ [إِذَنْ]{.ar} [Eipan]{.trn}

We will go over these cases in this chapter.

## Forming the a-state incomplete-action verb

Here is the u-state incomplete action verb for the singular masculine absentee participant doer "he":

[يَفْعَلُ]{.ar}  
[yafealu]{.trn}  
"he does"

Note that, because it is in the u-state, the its final letter ends with a [u]{.trn}-mark [◌ُ]{.ar}.
In order to form the 
a-state
incomplete-action verb,
we change the [u]{.trn}-mark into a
a-mark [◌َ]{.ar}, thus:

[يَفْعَلَ]{.ar}  
[yafeala]{.trn}  

This is done for all participants whose doer pronoun is invisible and u-state verb ends with a [u]{.trn}-mark [◌ُ]{.ar}.

For participants whose doer pronoun is followed by an extra [ن]{.ar} in the u-state verb, this final [ن]{.ar} is dropped in order to form the 
a-state
incomplete-action verb.
So, for example, the u-state
incomplete-action verb:

[يَفْعَلَانِ]{.ar}  
[yafealAni]{.trn}  
"they~2,m~ do"

becomes, for the 
a-state:

[يَفْعَلَا]{.ar}  
[yafealA]{.trn}  

Here is the complete table of the 
a-state
incomplete-action verb
for all doer participants.

|Participant|Incomplete-action verb doer pronoun|u-state incomplete-action verb | a-state incomplete-action verb|
|:---|:--|:---|:---|
| he        |_invisible_         |[يَفْعَلُ]{.ar}    |[يَفْعَلَ]{.ar}    |
| she       |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1m~   |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1f~   |[ي]{.ar}            |[تَفْعَلِينَ]{.ar}  |[تَفْعَلِي]{.ar}   |
| I         |_invisible_         |[أَفْعَلُ]{.ar}    |[أَفْعَلَ]{.ar}    |
| they~2m~  |[ا]{.ar}            |[يَفْعَلَانِ]{.ar}  |[يَفْعَلَا]{.ar}   |
| they~2f~  |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| you~2~    |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| they~3m~  |[و]{.ar}            |[يَفْعَلُونَ]{.ar}  |[يَفْعَلُوا]{.ar}  |
| they~3f~  |[نَ]{.ar}            |[يَفْعَلْنَ]{.ar}   |[يَفْعَلْنَ]{.ar} (same)   |
| you~3m~   |[و]{.ar}            |[تَفْعَلُونَ]{.ar}  |[تَفْعَلُوا]{.ar}  |
| you~3f~   |[نَ]{.ar}            |[تَفْعَلْنَ]{.ar}   |[تَفْعَلْنَ]{.ar} (same)  |
| we        |_invisible_         |[نَفْعَلُ]{.ar}    |[نَفْعَلَ]{.ar}    |

Take note the following:

+ The u-state and a-state verbs are the same for the feminine plural absentee and addressee participants: 
  + [يَفْعَلْنَ]{.ar} (they~3f~)
  + [تَفْعَلْنَ]{.ar} (you~3f~)
+ The a-state verbs for the masculine plural absentee and addressee participants have a final silent [Ealif]{.trn2}: 
  + [يَفْعَلُوا]{.ar} (they~3m~) 
  + [تَفْعَلُوا]{.ar} (you~3m~)

<!--
### Doubled and weak roots

The following table lists the a-state incomplete-action verb for some roots with a weak final letter. These are the only roots that have any complications in forming the a-state. The rest of the doubled and weak root verbs are straightforward. For a complete listing, including the higher forms, see Appendix\ \@ref(chapter-verb-tables).

Root: [سعي]{.arroot}

|Participant|u-state verb | a-state verb
|:---|:--|:---|:---|
| he        |_invisible_         |[يَفْعَلُ]{.ar}    |[يَفْعَلَ]{.ar}    |
| she       |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1m~   |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1f~   |[ي]{.ar}            |[تَفْعَلِينَ]{.ar}  |[تَفْعَلِي]{.ar}   |
| I         |_invisible_         |[أَفْعَلُ]{.ar}    |[أَفْعَلَ]{.ar}    |
| they~2m~  |[ا]{.ar}            |[يَفْعَلَانِ]{.ar}  |[يَفْعَلَا]{.ar}   |
| they~2f~  |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| you~2~    |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| they~3m~  |[و]{.ar}            |[يَفْعَلُونَ]{.ar}  |[يَفْعَلُوا]{.ar}  |
| they~3f~  |[نَ]{.ar}            |[يَفْعَلْنَ]{.ar}   |[يَفْعَلْنَ]{.ar} (same)   |
| you~3m~   |[و]{.ar}            |[تَفْعَلُونَ]{.ar}  |[تَفْعَلُوا]{.ar}  |
| you~3f~   |[نَ]{.ar}            |[تَفْعَلْنَ]{.ar}   |[تَفْعَلْنَ]{.ar} (same)  |
| we        |_invisible_         |[نَفْعَلُ]{.ar}    |[نَفْعَلَ]{.ar}    |
-->

## The a-state incomplete-action verb after [أَنْ]{.ar} [Ean]{.trn}

[أَنْ]{.ar} [Ean]{.trn} "that" is the main article which causes the following
incomplete-action verb to be in the a-state.
The other articles that we listed in the introduction are all either derived from 
[أَنْ]{.ar} 
or include its meaning implicitly without expressing it.

### Basic usage of [أَنْ]{.ar} [Ean]{.trn} with the a-state incomplete-action verb

[أَنْ]{.ar} often follows verbs that have a meaning of wishing or hoping. For example,

[أَمَلَ ٱلطَّالِبُ أَنْ يَنْجَحَ.]{.ar}  
[Eamala -TTAlibu Ean yanjaH.]{.trn}  
"The student hoped that he succeed."

[لَا]{.ar} can be used to negate the following a-state incomplete-action verb.
[لَا]{.ar} combines with [أَنْ]{.ar} and assimilates with it to form [أَلَّا]{.ar} [EallA]{.trn} "that not".
For example,

[أَمَرَ ٱلْأَبُ ٱلِٱبْنَ أَلَّا يَكْسَلَ.]{.ar}  
[Eamara -lEabu li-bna EallA yaksal.]{.trn}  
"The father ordered the son that he not be lazy."

Other than this [لَا]{.ar},
[أَنْ]{.ar} must directly precede the following a-state incomplete-action verb and must not be separated from it.

### Grammatical equivalence of [أَنْ]{.ar} clause with a doing verbal noun

In grammatical theory,
[أَنْ]{.ar} 
and the following verb form a clause that is equivalent in meaning to the doing verbal-noun of the verb. So in the example,
[أَمَلَ ٱلطَّالِبُ أَنْ يَنْجَحَ.]{.ar}, the 
[أَنْ]{.ar} clause
([أَنْ يَنْجَحَ]{.ar})
is equivalent to the doing verbal noun [ٱلنَّجَاح]{.ar}.
So the sentence is grammatically equivalent to

[أَمَلَ ٱلطَّالِبُ ٱلنَّجَاحَ.]{.ar}  
[Eamala -TTAlibu -nnajAH.]{.trn}  
"The student hoped [for] success."

This grammatical equivalence of the 
[أَنْ]{.ar} clause
with a noun aloows the 
[أَنْ]{.ar} clause to take the place of a noun in various positions in a sentence.
So, in the above example, the 
[أَنْ]{.ar} clause
is in place of the direct doee of the verb [أَمَلَ]{.ar}:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَنْ يَنْجَحَ} \txarb{أَمَلَ ٱلطَّالِبُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below] {\scriptsize direct doee};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``The student hoped that he suceed.''};

\end{tikzpicture}
```

We show other examples below where the
[أَنْ]{.ar} clause
occurs in place of other noun positions.

As the subject:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{هُوَ ٱلْمَقْصِدُ} \txarb{أَنْ تَنْجَحَ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize subject};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``That you succeed is the purpose.''};

\end{tikzpicture}
```

As the information:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَنْ تَنْجَحَ} \txarb{ٱَلْمَقْصِدُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below] {\scriptsize information};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``The purpose is that you succeed.''};

\end{tikzpicture}
```

As a doer noun:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَلَّا يَرْجِعَ} \txarb{يَكْبُدُنِي},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below] {\scriptsize doer};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``That he not return affects me severely.''};

\end{tikzpicture}
```

In the i-state as the base noun in an annexation:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{عَلَيَّ} \txarb{أَنْ تَغْضَبَ} \txarb{لَبِثْتُ فِي ٱلْبَيْتِ مِنْ خَشْيَةِ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize i-state base noun};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``I lingered at home from fear that you be angry at me.''};

\end{tikzpicture}
```

In the i-state after a preposition:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَنْ يَأْكُلَ} \txarb{رَغِبَ ٱلْغُلَامُ فِي},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize i-state after preposition};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``The boy desired that he eat.''};

\end{tikzpicture}
```

### Option to drop the preposition before [أَنْ]{.ar}

In the above example the verb [رَغِبَ يَرْغَبُ]{.ar} takes an indirect doee after the preposition [فِي]{.ar}.
In such cases, where the 
[أَنْ]{.ar} clause
occurs after a preposition, it is common to drop the preposition as long as there is not resulting confusion in meaning.
So, we can also say (without the preposition [فِي]{.ar}) for the same meaning:

[رَغِبَ ٱلْغُلَامُ آَنْ يَأْكُلَ.]{.ar}  
"The boy desired that he eat."

### Uncommon usages of [أَنْ]{.ar}

Ocassionally, [أَنْ]{.ar} is used with the meaning "lest". For example:

[قَتَلْتُ ٱلثُّعْبَانَ أَنْ يَقْتُلَنِي.]{.ar}  
"I killed the serpent lest it kill me."

[أَنْ]{.ar} may also occur before a completed-action verb. Example:

[بَلَغَنِي أَنْ رَجَعْتَ.]{.ar}  
"That you have returned has reached me."

### Other types of [أَنْ]{.ar}

There are other types of [أَنْ]{.ar} in the Arabic language. They all have the basic meaning "that". But they are used in different grammatical ways.

The [أَنْ]{.ar} we have learned here is called the _verbal noun [أَنْ]{.ar}_ because of the equivalence of its clause with a doing verbal noun.

There is also another type of [أَنْ]{.ar} called the _lightened [أَنْ]{.ar}_ that we will learn in section\ \@ref(lightened-an).

There is also the _explanatory [أَنْ]{.ar}_ and the _extra [أَنْ]{.ar}_ that we will cover in chapter\ \@ref(types-of-an).

## The a-state incomplete-action verb after [لِ]{.ar} [li]{.trn}

The article [أَنْ]{.ar} may be attached to the preposition [لِ]{.ar} [li]{.trn} thus: [لِأَنْ]{.ar} [liEan]{.trn} to give the meaning "so that". For example:

[أَكَلَ لِأَنْ يَشْبَعَ.]{.ar}  
"He ate so that he be sated."

When [لِ]{.ar} is thus used, [أَنْ]{.ar} is optionally allowed to be dropped while its meaning is retained.
[لِ]{.ar} is then attached to the verb.
So we can say, for the same meaning:

[أَكَلَ لِيَشْبَعَ.]{.ar}  
"He ate so that he be sated."

But when using [لَا]{.ar} to negate the verb, then [أَنْ]{.ar} must be expressed, and the combination of [لِ]{.ar}, [أَنْ]{.ar}, and [لَا]{.ar} is written as [لِئَلَّا]{.ar} [liEallA]{.trn}. For example,

[شَرِبَ ٱلْمَاءَ لِئَلَّا يَعْطَشَ.]{.ar}  
"He drank the water so that he not be thirsty."

By the way, this [لِ]{.ar}, which is used with a-state incomplete-action verbs, is different from 
the [لِ]{.ar} we learned in section\ \@ref(indirect-commands), which is used with the [0]{.txt}-state incomplete-action verbs for indirect commands.

## The a-state incomplete-action verb after [كَيْ]{.ar} [kay]{.trn}

[كَيْ]{.ar} [kay]{.trn}
is a preposition similar to [لِ]{.ar} in meaning. It may be translated as "in order that", or also as "so that".
It is also used before the a-state incomplete-action verb.
The difference from [لِ]{.ar} is that, when [لِ]{.ar} is used with the a-state incomplete-action verb, expressing or dropping the [أَنْ]{.ar} was optional.
But with [كَيْ]{.ar}, dropping the [أَنْ]{.ar} is mandatory, while its meaning is retained. For example:

[أَكَلَ كَيْ يَشْبَعَ.]{.ar}  
"He ate in order that he be sated."

[لَا]{.ar} is used, as usual, to negate the verb and is attached to [كَيْ]{.ar} thus: [كَيْلَا]{.ar} [kaylA]{.trn}. Example:

[شَرِبَ ٱلْمَاءَ كَيْلَا يَعْطَشَ.]{.ar}  
"He drank the water in order that he not be thirsty."

The preposition [لِ]{.ar} may be combined with [كَيْ]{.ar} thus: [لِكَيْ]{.ar} [likay]{.trn}, for more or less the same meaning. For example:

[أَكَلَ لِكَيْ يَشْبَعَ.]{.ar}  
"He ate in order that he be sated."

With [لَا]{.ar} the whole combination is written as [لِكَيْلَا]{.ar} [likaylA]{.trn}. [أَنْ]{.ar} must again be not be expressed.

Example:
[شَرِبَ ٱلْمَاءَ لِكَيْلَا يَعْطَشَ.]{.ar}  
"He drank the water in order that he not be thirsty."

## The a-state incomplete-action verb after [حَتَّىٰ]{.ar} [HattA]{.trn}

The preposition 
[حَتَّىٰ]{.ar} [HattA]{.trn}
is unusual in that the incomplete-action verb following it may either be in the a-state or the u-state, with slightly different meanings.
When the incomplete-action verb following it is in the u-state then
[حَتَّىٰ]{.ar} gives the meaning "until".
And when the incomplete-action verb following it is in the a-state then
[حَتَّىٰ]{.ar} gives the meaning "for the result that".
As with [كَيْ]{.ar}, [أَنْ]{.ar} must be not be expressed while its meaning is retained. Examples:

[يَأْكُلُ حَتَّىٰ يَشْبَعَ.]{.ar} (a-state)  
"He eats for the result that he be sated."

[يَأْكُلُ حَتَّىٰ يَشْبَعُ.]{.ar} (u-state)
"He eats until he is sated."

## The a-state incomplete-action verb after [إِذَنْ]{.ar} [Eipan]{.trn}

TODO

