# a-state incomplete-action verbs

## Introduction

In 
chapter\ \@ref(u-state-incomplete-action-verbs)
we mentioned that incomplete action verbs have three states (like nouns). 
These states are called:

i. The u-state
i. The a-state
i. The [0]{.txt}-state

We introduced the u-state incomplete-action verb in
chapter\ \@ref(u-state-incomplete-action-verbs).
In this chapter we will study the 
a-state
incomplete-action verb.

The u-state incomplete-action verb makes a plain statement.
The a-state incomplete-action verb implies a wish or purpose.
The a-state incomplete-action verb is used after the following articles:

+ [أَنْ]{.ar} [Ean]{.trn}
+ [لَنْ]{.ar} [lan]{.trn}
+ [لِ]{.ar} [li]{.trn}
+ [كَيْ]{.ar} [kay]{.trn}
+ [حَتَّىٰ]{.ar} [HattA]{.trn}
+ [إِذَنْ]{.ar} [Eipan]{.trn}

We will go over these cases in this chapter.

## Forming the a-state incomplete-action verb

Here is the u-state incomplete action verb for the singular masculine absentee participant doer "he":

[يَفْعَلُ]{.ar}  
[yafealu]{.trn}  
"he does"

Note that, because it is in the u-state, the its final letter ends with a [u]{.trn}-mark [◌ُ]{.ar}.
In order to form the 
a-state
incomplete-action verb,
we change the [u]{.trn}-mark into a
a-mark [◌َ]{.ar}, thus:

[يَفْعَلَ]{.ar}  
[yafeala]{.trn}  

This is done for all participants whose doer pronoun is invisible and u-state verb ends with a [u]{.trn}-mark [◌ُ]{.ar}.

For participants whose doer pronoun is followed by an extra [ن]{.ar} in the u-state verb, this final [ن]{.ar} is dropped in order to form the 
a-state
incomplete-action verb.
So, for example, the u-state
incomplete-action verb:

[يَفْعَلَانِ]{.ar}  
[yafealAni]{.trn}  
"they~2,m~ do"

becomes, for the 
a-state:

[يَفْعَلَا]{.ar}  
[yafealA]{.trn}  

Here is the complete table of the 
a-state
incomplete-action verb
for all doer participants.

|Participant|Incomplete-action verb doer pronoun|u-state incomplete-action verb | a-state incomplete-action verb|
|:---|:--|:---|:---|
| he        |_invisible_         |[يَفْعَلُ]{.ar}    |[يَفْعَلَ]{.ar}    |
| she       |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1m~   |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1f~   |[ي]{.ar}            |[تَفْعَلِينَ]{.ar}  |[تَفْعَلِي]{.ar}   |
| I         |_invisible_         |[أَفْعَلُ]{.ar}    |[أَفْعَلَ]{.ar}    |
| they~2m~  |[ا]{.ar}            |[يَفْعَلَانِ]{.ar}  |[يَفْعَلَا]{.ar}   |
| they~2f~  |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| you~2~    |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| they~3m~  |[و]{.ar}            |[يَفْعَلُونَ]{.ar}  |[يَفْعَلُوا]{.ar}  |
| they~3f~  |[نَ]{.ar}            |[يَفْعَلْنَ]{.ar}   |[يَفْعَلْنَ]{.ar} (same)   |
| you~3m~   |[و]{.ar}            |[تَفْعَلُونَ]{.ar}  |[تَفْعَلُوا]{.ar}  |
| you~3f~   |[نَ]{.ar}            |[تَفْعَلْنَ]{.ar}   |[تَفْعَلْنَ]{.ar} (same)  |
| we        |_invisible_         |[نَفْعَلُ]{.ar}    |[نَفْعَلَ]{.ar}    |

Take note the following:

+ The u-state and a-state verbs are the same for the feminine plural absentee and addressee participants: 
  + [يَفْعَلْنَ]{.ar} (they~3f~)
  + [تَفْعَلْنَ]{.ar} (you~3f~)
+ The a-state verbs for the masculine plural absentee and addressee participants have a final silent [A]{.abar}: 
  + [يَفْعَلُوا]{.ar} (they~3m~) 
  + [تَفْعَلُوا]{.ar} (you~3m~)

<!--
### Doubled and weak roots

The following table lists the a-state incomplete-action verb for some roots with a weak final letter. These are the only roots that have any complications in forming the a-state. The rest of the doubled and weak root verbs are straightforward. For a complete listing, including the higher forms, see Appendix\ \@ref(chapter-verb-tables).

Root: [سعي]{.arroot}

|Participant|u-state verb | a-state verb
|:---|:--|:---|:---|
| he        |_invisible_         |[يَفْعَلُ]{.ar}    |[يَفْعَلَ]{.ar}    |
| she       |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1m~   |_invisible_         |[تَفْعَلُ]{.ar}    |[تَفْعَلَ]{.ar}    |
| you~1f~   |[ي]{.ar}            |[تَفْعَلِينَ]{.ar}  |[تَفْعَلِي]{.ar}   |
| I         |_invisible_         |[أَفْعَلُ]{.ar}    |[أَفْعَلَ]{.ar}    |
| they~2m~  |[ا]{.ar}            |[يَفْعَلَانِ]{.ar}  |[يَفْعَلَا]{.ar}   |
| they~2f~  |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| you~2~    |[ا]{.ar}            |[تَفْعَلَانِ]{.ar}  |[تَفْعَلَا]{.ar}   |
| they~3m~  |[و]{.ar}            |[يَفْعَلُونَ]{.ar}  |[يَفْعَلُوا]{.ar}  |
| they~3f~  |[نَ]{.ar}            |[يَفْعَلْنَ]{.ar}   |[يَفْعَلْنَ]{.ar} (same)   |
| you~3m~   |[و]{.ar}            |[تَفْعَلُونَ]{.ar}  |[تَفْعَلُوا]{.ar}  |
| you~3f~   |[نَ]{.ar}            |[تَفْعَلْنَ]{.ar}   |[تَفْعَلْنَ]{.ar} (same)  |
| we        |_invisible_         |[نَفْعَلُ]{.ar}    |[نَفْعَلَ]{.ar}    |
-->

## After [أَنْ]{.ar} [Ean]{.trn}

[أَنْ]{.ar} [Ean]{.trn} "that" is the main article which causes the following
incomplete-action verb to be in the a-state.
The other articles that we listed in the introduction are all either derived from 
[أَنْ]{.ar} 
or include its meaning implicitly without expressing it.

### Basic usage of [أَنْ]{.ar} [Ean]{.trn} with the a-state incomplete-action verb

[أَنْ]{.ar} often follows verbs that have a meaning of wishing or hoping. For example,

[أَمَلَ ٱلطَّالِبُ أَنْ يَنْجَحَ.]{.ar}  
[Eamala -TTAlibu Ean yanjaH.]{.trn}  
"The student hoped that he succeed."

[لَا]{.ar} can be used to negate the following a-state incomplete-action verb.
[لَا]{.ar} combines with [أَنْ]{.ar} and assimilates with it to form [أَلَّا]{.ar} [EallA]{.trn} "that not".
For example,

[أَمَرَ ٱلْأَبُ ٱلِٱبْنَ أَلَّا يَكْسَلَ.]{.ar}  
[Eamara -lEabu li-bna EallA yaksal.]{.trn}  
"The father ordered the son that he not be lazy."

Other than this [لَا]{.ar},
[أَنْ]{.ar} must directly precede the following a-state incomplete-action verb and must not be separated from it.

### Grammatical equivalence of [أَنْ]{.ar} clause with a doing verbal noun

In grammatical theory,
[أَنْ]{.ar} 
and the following verb form a clause that is equivalent in meaning to the doing verbal-noun of the verb. So in the example,
[أَمَلَ ٱلطَّالِبُ أَنْ يَنْجَحَ.]{.ar}, the 
[أَنْ]{.ar} clause
([أَنْ يَنْجَحَ]{.ar})
is equivalent to the doing verbal noun [ٱلنَّجَاح]{.ar}.
So the sentence is grammatically equivalent to

[أَمَلَ ٱلطَّالِبُ ٱلنَّجَاحَ.]{.ar}  
[Eamala -TTAlibu -nnajAH.]{.trn}  
"The student hoped [for] success."

This grammatical equivalence of the 
[أَنْ]{.ar} clause
with a noun aloows the 
[أَنْ]{.ar} clause to take the place of a noun in various positions in a sentence.
So, in the above example, the 
[أَنْ]{.ar} clause
is in place of the direct doee of the verb [أَمَلَ]{.ar}:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَنْ يَنْجَحَ} \txarb{أَمَلَ ٱلطَّالِبُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below] {\scriptsize direct doee};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``The student hoped that he suceed.''};

\end{tikzpicture}
```

We show other examples below where the
[أَنْ]{.ar} clause
occurs in place of other noun positions.

As the subject:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{هُوَ ٱلْمَقْصِدُ} \txarb{أَنْ تَنْجَحَ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize subject};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``That you succeed is the purpose.''};

\end{tikzpicture}
```

As the information:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَنْ تَنْجَحَ} \txarb{ٱَلْمَقْصِدُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below] {\scriptsize information};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``The purpose is that you succeed.''};

\end{tikzpicture}
```

As a doer noun:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَلَّا يَرْجِعَ} \txarb{يَكْبُدُنِي},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below] {\scriptsize doer};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``That he not return affects me severely.''};

\end{tikzpicture}
```

In the i-state as the base noun in an annexation:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{عَلَيَّ} \txarb{أَنْ تَغْضَبَ} \txarb{لَبِثْتُ فِي ٱلْبَيْتِ مِنْ خَشْيَةِ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize i-state base noun};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``I lingered at home from fear that you be angry at me.''};

\end{tikzpicture}
```

In the i-state after a preposition:

```{r,echo=FALSE,engine='tikz', fig.ext='pdf', engine.opts = list(template = if (knitr::is_html_output()) "srctex/tikz2pdf-lua-html.tex" else "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.45ex}]
\path [decoration={text effects along path, 
  text=.\txarb{أَنْ يَأْكُلَ} \txarb{رَغِبَ ٱلْغُلَامُ فِي},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize i-state after preposition};

\path (0,0) node [midway, below=6ex, right=-2ex, align=left] {``The boy desired that he eat.''};

\end{tikzpicture}
```

### Option to drop the preposition before [أَنْ]{.ar}

In the above example the verb [رَغِبَ يَرْغَبُ]{.ar} takes an indirect doee after the preposition [فِي]{.ar}.
In such cases, where the 
[أَنْ]{.ar} clause
occurs after a preposition, it is common to drop the preposition as long as there is not resulting confusion in meaning.
So, we can also say (without the preposition [فِي]{.ar}) for the same meaning:

[رَغِبَ ٱلْغُلَامُ آَنْ يَأْكُلَ.]{.ar}  
"The boy desired that he eat."

### Uncommon usages of [أَنْ]{.ar}

Ocassionally, [أَنْ]{.ar} is used with the meaning "lest". For example:

[قَتَلْتُ ٱلثُّعْبَانَ أَنْ يَقْتُلَنِي.]{.ar}  
"I killed the serpent lest it kill me."

[أَنْ]{.ar} may also occur before a completed-action verb. Example:

[بَلَغَنِي أَنْ رَجَعْتَ.]{.ar}  
"That you have returned has reached me."

### Other types of [أَنْ]{.ar}

There are other types of [أَنْ]{.ar} in the Arabic language. They all have the basic meaning "that". But they are used in different grammatical ways.

The [أَنْ]{.ar} we have learned here is called the _verbal noun [أَنْ]{.ar}_ because of the equivalence of its clause with a doing verbal noun.

There is also another type of [أَنْ]{.ar} called the _lightened [أَنْ]{.ar}_ that we will learn in section\ \@ref(lightened-an).

There is also the _explanatory [أَنْ]{.ar}_ and the _extra [أَنْ]{.ar}_ that we will cover in chapter\ \@ref(types-of-an).

## After [لِ]{.ar} [li]{.trn}

### The [لِ]{.ar} of purpose

The article [أَنْ]{.ar} may be attached to the preposition [لِ]{.ar} [li]{.trn} thus: [لِأَنْ]{.ar} [liEan]{.trn} to give the purpose of the following verb. This [لِ]{.ar} may be translated as "so that". For example:

[أَكَلَ لِأَنْ يَشْبَعَ.]{.ar}  
"He ate so that he be sated."

When [لِ]{.ar} is thus used, [أَنْ]{.ar} is optionally allowed to be dropped while its meaning is retained.
[لِ]{.ar} is then attached to the verb.
So we can say, for the same meaning:

[أَكَلَ لِيَشْبَعَ.]{.ar}  
"He ate so that he be sated."

But when using [لَا]{.ar} to negate the verb, then [أَنْ]{.ar} must be expressed, and the combination of [لِ]{.ar}, [أَنْ]{.ar}, and [لَا]{.ar} is written as [لِئَلَّا]{.ar} [liEallA]{.trn}. For example,

[شَرِبَ ٱلْمَاءَ لِئَلَّا يَعْطَشَ.]{.ar}  
"He drank the water so that he not be thirsty."

<!--
By the way, this [لِ]{.ar}, which is used with a-state incomplete-action verbs, is different from 
the [لِ]{.ar} we learned in section\ \@ref(indirect-commands), which is used with the [0]{.txt}-state incomplete-action verbs for indirect commands.
-->

By the way, the grammatical equivalence of [أَنْ]{.ar} and a following a-state incomplete-action verb with a verbal noun of doing applies also to when [لِ]{.ar} is used before (either an expressed or an implied) [أَنْ]{.ar}. So, for example, if we have a sentence:

[قَرَأَ ٱلْكِتَابَ لِيَعْلَمَ مَفْهُومَهُ.]{.ar} or  
[قَرَأَ ٱلْكِتَابَ لِأَنْ يَعْلَمَ مَفْهُومَهُ.]{.ar}  
"He read the book so that he know its meaning."

Then, grammatically, [أَنْ]{.ar} and what follows it may be expressed with the verbal noun of doing [عِلْم]{.ar} thus:

[قَرَأَ ٱلْكِتَابَ لِعِلْمِ مَفْهُومِهِ.]{.ar}  
"He read the book for the knowledge of its meaning."

### The [لِ]{.ar} of denial

There is a specific [لِ]{.ar}, called the [لِ]{.ar} of denial,  which is used with a-state incomplete-action verbs and the verb [كَانَ]{.ar} that we will discuss in section (TODO in [كَانَ]{.ar} chapter).

## After [كَيْ]{.ar} [kay]{.trn}

[كَيْ]{.ar} [kay]{.trn}
is a preposition similar to [لِ]{.ar} in meaning. It may be translated as "in order that", or also as "so that".
It is also used before the a-state incomplete-action verb.
The difference from [لِ]{.ar} is that, when [لِ]{.ar} is used with the a-state incomplete-action verb, expressing or dropping the [أَنْ]{.ar} was optional.
But with [كَيْ]{.ar}, dropping the [أَنْ]{.ar} is mandatory, while its meaning is retained. For example:

[أَكَلَ كَيْ يَشْبَعَ.]{.ar}  
"He ate in order that he be sated."

[لَا]{.ar} is used, as usual, to negate the verb and is attached to [كَيْ]{.ar} thus: [كَيْلَا]{.ar} [kaylA]{.trn}. Example:

[شَرِبَ ٱلْمَاءَ كَيْلَا يَعْطَشَ.]{.ar}  
"He drank the water in order that he not be thirsty."

The preposition [لِ]{.ar} may be combined with [كَيْ]{.ar} thus: [لِكَيْ]{.ar} [likay]{.trn}, for more or less the same meaning. For example:

[أَكَلَ لِكَيْ يَشْبَعَ.]{.ar}  
"He ate in order that he be sated."

With [لَا]{.ar} the whole combination is written as [لِكَيْلَا]{.ar} [likaylA]{.trn}. [أَنْ]{.ar} must again be not be expressed.

Example:

[شَرِبَ ٱلْمَاءَ لِكَيْلَا يَعْطَشَ.]{.ar}  
"He drank the water in order that he not be thirsty."

By the way, 
[كَيْ]{.ar} and a following a-state incomplete-action verb are not _directly_ replaced by
a verbal noun of doing.
So, for example, if we have a sentence:

[قَرَأَ ٱلْكِتَابَ كَيْ يَعْلَمَ مَفْهُومَهُ.]{.ar}  
"He read the book in order that he know its meaning."

Then [لِ]{.ar} is to be used in place of [كَيْ]{.ar} if we wish to replace it and what follows with the verbal noun of doing [عِلْم]{.ar} thus:

[قَرَأَ ٱلْكِتَابَ لِعِلْمِ مَفْهُومِهِ.]{.ar}  
"He read the book for the knowledge of its meaning."

## After [حَتَّىٰ]{.ar} [HattA]{.trn}

[حَتَّىٰ]{.ar} [HattA]{.trn}
is a particle that can be used in multiple ways.
Its basic meaning is "until" or "to the point of" or "even" where it indicates an extreme limit.

Before we discuss its use with a verb following it, we will take a short digression to discuss its use with a following noun.

### [حَتَّىٰ]{.ar} [HattA]{.trn} with a following noun

Consider the following sentence:

[أَكَلْتُ ٱلسَّمَكَةَ حَتَّىٰ رَأْسِهَا.]{.ar}  
"I ate the fish until its head."

[حَتَّىٰ]{.ar} "until",
here, is used as a preposition.
Therefore,
[رَأْس]{.ar}
is in the i-state, as the noun following a preposition.
The meaning of the sentence is that the fish was eaten all the way to its head. 
(Whether the head itself was eaten or not is ambiguous. The sentence itself admits both meanings.)

Consider now a variant of this sentence:

[أَكَلْتُ ٱلسَّمَكَةَ حَتَّىٰ رَأْسَهَا.]{.ar}  
"I ate the fish, even its head."

[رَأْس]{.ar},
here, is in the a-state because it is a direct doee of the verb [أَكَلَ]{.ar} "ate".
The particle
[حَتَّىٰ]{.ar} "even",
here, is only a connector between the direct doees in much the same way as [وَ]{.ar} "and" functions. For example,
[أَكَلْتُ ٱلسَّمَكَةَ وَرَأْسَهَا.]{.ar}
"I ate the fish and its head."

Consider now yet another variant of this sentence:

[أَكَلْتُ ٱلسَّمَكَةَ. حَتَّىٰ رَأْسُهَا [أَكَلْتُهَا].]{.ar}  
"I ate the fish. Even its head [I ate]."

Now
[رَأْس]{.ar}
is in the u-state because it is actually the subject of a new sentence, whose information is (an either expressed or implied) [أَكَلْتُهَا]{.ar} "I ate it".
[حَتَّىٰ]{.ar},
here, serves as an introductory particle to the second subject and does not affect the state of the following noun.

### [حَتَّىٰ]{.ar} [HattA]{.trn} with a following verb

<!-- See
Wright vol. ii. Section 15, p. 29C,D

Arik Sadan, "The Mood of The Verb following Ḥattā, according To Medieval arab grammarians"
in
The foundations of Arabic linguistics : Sibawayhi and early Arabic grammatical theory / edited by Amal
Elesha Marogy
-->

Just as
[حَتَّىٰ]{.ar}
is used for different purposes with a following noun,
so too is it used with different purposes with a verb following it.

#### [حَتَّىٰ]{.ar} with a following a-state incomplete-action verb

When the verb following
[حَتَّىٰ]{.ar}
is used with an expectation or purpose of a future action, then the verb is used as an a-state incomplete-action verb. 
This is done in the following two scenarios:

1. When [حَتَّىٰ]{.ar} is used to indicate an extreme point at which the action of the following verb would, or is meant to, occur. For example,

   [قَرَأْتُ ٱلْقُرْآنَ حَتَّىٰ أَخْتِمَهُ.]{.ar}  
   "I read the [#qurEAn]{.trn2} to the point of finishing it."

   [يَغْضَبُ حَتَّىٰ يَهْرَبُوا مِنْهُ.]{.ar}  
   "He becomes angry to the point of their fleeing from him."

   [غَضِبَ حَتَّىٰ لَا يَمْلِكَ نَفْسَهُ.]{.ar}  
   "He became angry to the point of not controlling himself."

   It is noteworthy that the use of [حَتَّىٰ]{.ar}, here, implies only that the following action is meant, or at the point of being expected, to occur. It doesn't actually state that the action actually occured, for something may have prevented it from occurring in reality.

   Note, also, that [لَا]{.ar} is not attached to [حَتَّىٰ]{.ar} in [حَتَّىٰ لَا]{.ar}.

   Also, similar to the case of [كَيْ]{.ar}, there is an assumed (but mandatorily unexpressed) [أَنْ]{.ar} which is the real cause of the following incomplete-action verb being in the a-state.
   In fact, 
   [حَتَّىٰ]{.ar},
   here, can be considered synonymous to [إِلَىٰ أَنْ]{.ar} "to [the point] that". So the above two examples can be considered similar in meaning to:

   [قَرَأْتُ ٱلْقُرْآنَ إِلَىٰ أَنْ أَخْتِمَهُ.]{.ar}  

   [يَغْضَبُ إِلَىٰ أَنْ يَهْرَبُوا مِنْهُ.]{.ar}  

   [غَضِبَ إِلَىٰ أَلَّا يَمْلِكَ نَفْسَهُ.]{.ar}  

2. When [حَتَّىٰ]{.ar} is used with the meaning "to such a purpose that". This is a similar meaning to [كَيْ]{.ar} "in order that".
   For example,

   [أَذْهَبُ إِلَيْهِ حَتَّىٰ يَأْمُرَنِي بِشَيْءٍ.]{.ar}  
   "I go to him to such a purpose that he order me [to do] something."

   [وَعَظَ ٱلْأَبُ ٱبْنَهُ حَتَّىٰ يَصْلُحَ.]{.ar}  
   "The father admonished his son to such a purpose that he be righteous."

   Again, there is an assumed (but mandatorily unexpressed) [أَنْ]{.ar} which is the real cause of the following incomplete-action verb being in the a-state.

Sometimes, the sentence itself may admit both of the above meanings. For example:

[يَأْكُلُ حَتَّىٰ يَشْبَعَ.]{.ar}  
"He eats to the point of being full."  
and/or  
"He eats to the purpose that he be full."

Context would be needed to determine which meaning or whether both meanings are intended.

#### [حَتَّىٰ]{.ar} with no effect on the following verb

If
[حَتَّىٰ]{.ar}
is not used with any expectation or purpose of a future action of the verb following it, then it has no effect on this verb.
(It goes without saying that an implicit [أَنْ]{.ar} is not assumed with [حَتَّىٰ]{.ar} in this case.)
In fact, the following verb may be a completed-action verb. For example:

[أَكَلْتُ ٱلطَّعَامَ حَتَّىٰ شَبِعْتُ.]{.ar}  
"I ate the food until I became full."

When used with a following incomplete-action verb, the verb is put in the u-state and the meaning is that the action of the verb _before_
[حَتَّىٰ]{.ar}
was done to such an extent that it caused
the action of the verb _following_
[حَتَّىٰ]{.ar}
to definitely occur.
The action before
[حَتَّىٰ]{.ar}
must necessarily be a past action, and the action following
[حَتَّىٰ]{.ar}
must necessarily be a present (not a future) action.
For example,

[أَكَلْتُ ٱلطَّعَامَ حَتَّىٰ أَشْبَعُ.]{.ar}  
"I ate the food to such an extent that I am (being) full."

[غَضِبَ حَتَّىٰ يَهْرَبُونَ مِنْهُ.]{.ar}  
"He became so angry that they are fleeing from him."

[غَضِبَ حَتَّىٰ لَا يَمْلِكُ نَفْسَهُ.]{.ar}  
"He became so angry that he is not controlling himself."

Compare these examples with the corresponding ones in the previous sub-section that have an a-state incomplete action verb.


<!--The preposition 
[حَتَّىٰ]{.ar} [HattA]{.trn}
is unusual in that the incomplete-action verb following it may either be in the a-state or the u-state, with slightly different meanings.
When the incomplete-action verb following it is in the u-state then
[حَتَّىٰ]{.ar} gives the meaning "until".
And when the incomplete-action verb following it is in the a-state then
[حَتَّىٰ]{.ar} gives the meaning "for the result that".
A mandatorily unexpressed [أَنْ]{.ar} is assumed with
[حَتَّىٰ]{.ar} 
when the verb following it is in the a-state.
Examples:

[يَأْكُلُ حَتَّىٰ يَشْبَعَ.]{.ar} (a-state, unexpressed [أَنْ]{.ar} assumed)  
"He eats for the result that he be sated."

[يَأْكُلُ حَتَّىٰ يَشْبَعُ.]{.ar} (u-state, no [أَنْ]{.ar})  
"He eats until he is sated."

[حَتَّىٰ]{.ar} 
is also used with completed-action verbs. Example:

[أَكَلَ حَتَّىٰ شَبِعَ.]{.ar} (completed-action, no [أَنْ]{.ar})  
"He ate until he was sated."
-->

## After [لَنْ]{.ar} [lan]{.trn}

[لَا]{.ar} and [أَنْ]{.ar} are combined to form [لَنْ]{.ar} [lan]{.trn} with the meaning "shall not". [لَنْ]{.ar} is used with the a-state incomplete-action verb to emphatically negate the future.

[لَنْ تَذْهَبَ.]{.ar}  
"You~1m~ shall not go."

## After [إِذَنْ]{.ar} [Eipan]{.trn}

TODO

## After [وَ]{.ar}, [فَ]{.ar}, [أَوْ]{.ar}, and [ثُمَّ]{.ar}

### As connectors

If the connectors
[وَ]{.ar}, [فَ]{.ar}, [أَوْ]{.ar}, and [ثُمَّ]{.ar}
occur after an a-state incomplete-action verb, then 
a second a-state incomplete-action verb
(that doesn't have its own [أَنْ]{.ar}, etc.)
may be either in the a-state or the u-state.
For example,

[أَرْغَبُ أَنْ أَحْضُرَ ٱلْمَجْلِسَ وَأَسْمَعَ.]{.ar} ([أَسْمَعَ]{.ar} in a-state)  
"I desire that I attend the session and [that] I listen."

or

[أَرْغَبُ أَنْ أَحْضُرَ ٱلْمَجْلِسَ وَأَسْمَعُ.]{.ar} ([أَسْمَعَ]{.ar} in u-state)  
"I desire that I attend the session and I will listen."

<!--
 كتاب شرح المفصل لابن يعيش
 vol. 4, p. 259
 https://shamela.ws/book/13301/1623
-->

### With special meanings

[وَ]{.ar}, [فَ]{.ar}, [أَوْ]{.ar}, and [ثُمَّ]{.ar} also cause the following incomplete-action verb to be in the a-state in their own right, not simply as connectors. This is discussed in more detail in chapter TODO.

