# Pointing nouns

## Introduction

Consider the following expression:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=this book,
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize pointing noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize pointed-to noun};

\end{tikzpicture}
```

The word "this" is what we will call a _pointing noun_. We call it this because we can imagine standing next to a book and pointing to it and saying "this book".

The word "book" here is similarly called the _pointed-to_ noun. It refers to the object being pointed to.

## The pointing nouns in Arabic

There are two types of pointing nouns:

i. Near pointing nouns: "this-one" (singular) and "these-ones" (dual and plural).
i. Far  pointing nouns: "that-one" (singular) and "those-ones" (dual and plural).

The following are the pointing nouns in Arabic:

| Participant |State| Near pointing  noun || Far pointing  noun ||
|:----|:-|--:|:----|--:|:----|
|sing. masc.|all|[هَـٰذَا]{.ar}  |this one~m~    |[ذَ ٰلِكَ]{.ar}  |that one~m~  |
|sing. fem. |all|[هَـٰذِهِ]{.ar}  |this one~f~    |[تِلْكَ]{.ar}   |that one~f~  |
|dual masc. |u  |[هَـٰذَانِ]{.ar} |these ones~2,m~|[ذَ ٰنِكَ]{.ar}  |those ones~2,m~|
|dual masc. |a,i|[هَـٰذَيْنِ]{.ar} |these ones~2,m~|[ذَيْنِكَ]{.ar}  |those ones~2,m~|
|dual fem.  |u  |[هَاتَانِ]{.ar} |these ones~2,f~|[تَانِكَ]{.ar}  |those ones~2,f~|
|dual fem.  |a,i|[هَاتَيْنِ]{.ar} |these ones~2,f~|[تَيْنِكَ]{.ar}  |those ones~2,f~|
|plural     |all|[هَـٰؤُلَاءِ]{.ar}|these ones~3~  |[أُولَـٰئِكَ]{.ar}|those ones~3~  |

Note the following:

+ Many of the pointing nouns contain small [Ealif]{.trn} [◌ٰ]{.ar}. For most of them, this is how they must be written. It would be incorrect to write [هَـٰذَا]{.ar} [hApA]{.trn} as [هَاذَا]{.ar}.
+ All the near pointing nouns begin with a [ه]{.ar}. And all the far pointing nouns end with [ك]{.ar}.
+ The [و]{.ar} in [أُولَـٰئِكَ]{.ar} [EulAEika]{.trn} is silent and not pronounced. That is, the first syllable has a short vowel [u]{.trn}, not the long vowel [U]{.trn}.
+ Most of the pointing nouns are rigid nouns. That is: their endings are not modified for their state. 

  The dual pointing nouns, however, are flexible nouns, for example: [هَـٰذَانِ]{.ar}  (u-state)   / [هَـٰذَيْنِ]{.ar}  [hApayni]{.trn} (a- and i-states).

+ The pointing nouns for the plural are the same for both masculine and feminine genders.

## Definiteness of pointing nouns

The pointing nouns share some similarities with pronouns [هُوَ]{.ar}, [هِيَ]{.ar}, etc. Just like pronouns, pointing nouns, too, are definite nouns even though they don't have [ٱَلْ]{.ar}.

Remember, however, from section\ \@ref(describers-with-annexations-to-pronouns), that pronouns may not be describees.
Pointing nouns are different from pronouns in this regard. It is allowed to describe a pointing noun with a describer in a noun phrase.

Both these facts will prove useful in the next section.

## The pointing noun phrase

Remember from chapter\ \@ref(adjectival-nouns-and-descriptive-noun-phrases) that a descriptive noun-phrase consists of a describer and a describee. The describer follows the describer and matches it in definiteness, state, gender, and number. 

Here is an example of a descriptive noun-phrase in a sentence.

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=\txarb{ٱلصَّغِيرَينِ} \txarb{ٱلْكِتَابَيْنِ} \txarb{قَرَأْتُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize\uppercase{describer}};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize\uppercase{describee}};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-2.south east)
      node [midway, below=9ex] {\scriptsize\uppercase{descriptive noun-phrase}};

\path [decoration={text effects along path, 
  text={``I read the small books''},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate, below=14ex] (0,0);


\end{tikzpicture}
```




## Remove

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=This is a book,
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize pointing noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-4.south west) -- (word-4.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize pointed-to noun};

\end{tikzpicture}
```


Pointing nouns are the Arabic equivalents of words like "this", "that", "these", "those". We call them pointing nouns because a person saying "This man.", for example, could be pointing to a man near him. Pointing nouns are used in sentences like:

+ "This is a house."
+ "These women are good."
+ "That is a book."

The pointing nouns in Arabic are:


Note the following regarding the pointing nouns shown above:

+ Many of the pointing nouns contain small [Ealif]{.trn} [◌ٰ]{.ar}. For most of them, this is how they must be written. It would be incorrect to write [هَـٰذَا]{.ar} [hApA]{.trn} as [هَاذَا]{.ar}.
+ The [و]{.ar} in [أُولَـٰئِكَ]{.ar} [EulAEika]{.trn} is silent and not pronounced. That is, the first syllable has a short vowel [u]{.trn}, not the long vowel [U]{.trn}.
+ The pointing nouns are the same in all states except for the dual, where [ـَانِ]{.ar} [-Ani]{.trn} in the u-state becomes [ـَيْنِ]{.ar} [-ayni]{.trn} in the a-state and i-state.
+ The pointing nouns for the plural are the same for both masculine and feminine genders.

### Using pointing nouns with nouns or phrases not beginning with [ٱَلْ]{.ar}

If a pointing noun comes before a noun or phrase that does not begin with [ٱَلْ]{.ar} then the expression is a complete sentence. The pointing noun is the subject of sentence, and the other noun or phrase is the information of the sentence. Here are some examples:

+ [هَـٰذَا رَجُلٌ.]{.ar}  
  [hApA rajulun.]{.trn}  
  "This~masc.~ is a man."
+ [هَـٰذَانِ مُعَلِّمَانِ كَرِيمَانِ.]{.ar}  
  [hApAni mueallimAni karImAni.]{.trn}  
  "These~dual\ masc.~ are noble (male) teachers."
+ [هَـٰؤُلَاءِ مُعَلِّمُو غُلَامٍ.]{.ar}  
  [hAEulAEi mueallimU gulAmin.]{.trn}  
  "These~pl.~ are (male) teachers of a boy."
+ [تِلْكَ فَاطِمَةُ.]{.ar}  
  [tilka fATimatu.]{.trn}  
  "That~fem.~ is [#faTimah]{.trn2}."
+ [تَانِكَ مُعَلِّمَتَا ٱلْجَارِيَةِ.]{.ar}  
  [tAnika mueallimata -ljAriyati.]{.trn}  
  "Those~dual\ fem.~ are the (female) teachers of the girl."
+ [هَـٰذِهِ رِسَالَتِي.]{.ar}  
  [hApihi risAlatI.]{.trn}  
  "This is my message."

In the first three examples above, the information of the sentence is an indefinite noun or noun-phrase. 
In the fourth example, the information is a proper noun which is definite without [ٱَلْ]{.ar}.
In the last example, the information is a noun-chain whose last noun is definite. Therefore the first noun is definite as well, although, again, without [ٱَلْ]{.ar}.

So you can see that the rule we have stated, that if a pointing noun comes before a noun or phrase that does not begin with [ٱَلْ]{.ar} then the expression is a complete sentence, that this rule applies regardless of whether the noun or phrase is definite or indefinite.

## Using pointing nouns with nouns or phrases beginning with [ٱَلْ]{.ar}

### The pointing noun before the other noun

If a pointing noun comes before a noun that begins with [ٱَلْ]{.ar}, for example,  
[هَـٰذَا ٱَلرَّجُلُ]{.ar}  
[hApA -rrujulu]{.trn},  
then the expression can either:

 i. be a noun phrase: "this man", or
ii. be a complete sentence: "This is the man."

Here is another example:

[هَـٰذَانِ ٱلْمُعَلِّمَانِ ٱلْكَرِيمَانِ.]{.ar}  
[hApAni -lmueallimAni -lkarImAni.]{.trn}  
"These~dual\ masc.~ noble (male) teachers"  
or  
"These~dual\ masc.~ are the noble (male) teachers."

Context would be needed to determine which of the two meanings is intended.

In order to disambiguate the complete sentence from the noun phrase, the detached pronoun is often (but not always) placed between the pointing noun and the information of the sentence thus:

[هَـٰذَانِ هُمَا ٱلْمُعَلِّمَانِ ٱلْكَرِيمَانِ.]{.ar}  
[hApAni huma -lmueallimAni -lkarImAni.]{.trn}  
"These~dual\ masc.~ are the noble (male) teachers"  

Here are some examples in complete sentences:

+ [قَرَأْتُ ذَيْنِكَ ٱلْكِتَابَيْنِ.]{.ar}  
  [qaraEtu paynika -lkitAbayni.]{.trn}  
  "I read those two books."

### The pointing noun after the other noun

If a pointing noun comes after a noun that begins with [ٱَلْ]{.ar}, for example,  
[ٱَلرَّجُلُ هَـٰذَا]{.ar}  
[Earrujulu hApA]{.trn},  
then the expression can only be a noun-phrase: "this man". 

Despite the disambiguity in expressing the noun-phrase with this second method, the first method of placing the pointing noun before the other noun is more commonly used. That is, in order to express "this man",  
[هَـٰذَا ٱَلرَّجُلُ]{.ar}  
[hApA -rrujulu]{.trn}  
is more commonly used than  
[ٱَلرَّجُلُ هَـٰذَا]{.ar}  
[Earrujulu hApA]{.trn}.

Here are some examples in complete sentences:

## Pointing nouns in noun-phrases with definite nouns not beginning with [ٱَلْ]{.ar}

It remains to see how to express pointing nouns in phrases with defiinite nouns not beginning with [ٱَلْ]{.ar}. Such nouns include:

 i. proper nouns, and
ii. the first noun of a noun-chain whose last noun is definite.

In other words, we would like to express phrases like:

+ "that [#fATimah]{.trn2}", used for example in the sentence "This [#fATimah]{.trn2} is tall and that [#fATimah]{.trn2} is short."
+ "these~pl.~ (female) teachers of the girl"
+ "this message of mine"

We know that we cannot place the pointing noun in the beginning of the phrase because, as we saw, that would make it a complete sentence. For example,

+ [تِلْكَ فَاطِمَةُ.]{.ar}  
  [tilka fATimatu.]{.trn}  
  "That~fem.~ is [#faTimah]{.trn2}."
+ [تَانِكَ مُعَلِّمَتَا ٱلْجَارِيَةِ.]{.ar}  
  [tAnika mueallimata -ljAriyati.]{.trn}  
  "Those~dual\ fem.~ are the (female) teachers of the girl."
+ [هَـٰذِهِ رِسَالَتِي.]{.ar}  
  [hApihi risAlatI.]{.trn}  
  "This is my message."

So, we are left with placing the pointing noun at the end of the phrase, thus:

+ [فَاطِمَةُ تِلْكَ.]{.ar}  
  [fATimatu tilka]{.trn}  
  "that~fem.~ [#faTimah]{.trn2}"
+ [مُعَلِّمَتَا ٱلْجَارِيَةِ تَانِكَ]{.ar}  
  [mueallimata -ljAriyati tAnika]{.trn}  
  "those~dual\ fem.~ (female) teachers of the girl"
+ [رِسَالَتِي هَـٰذِهِ]{.ar}  
  [risAlatI hApihi]{.trn}  
  "this message of mine"

Here are some examples in complete sentences:

+ [فَاطِمَةُ هَـٰذِهِ طَوِيلَةٌ وَفَاطِمَةُ تِلْكَ قَصِيرَةٌ.]{.ar}  
  [fATimatu hApihi TawIlatun wafATimatu tilka qaSIratun.]{.trn}  
  "This [#fATimah]{.trn2} is tall and that [#fATimah]{.trn2} is short."

## Using pointing nouns with plural non-rational beings

### REMOVE: In noun-phrases with common nouns

A pointing noun can be used in a noun phrase for expressions like "this man". We can express this using the pointing noun 
[هَـٰذَا]{.ar} [hApA]{.trn} "this~masc.~ one" and the noun being pointed to [ٱَلرَّجُلُ]{.ar} [Earrujulu]{.trn} "the man". We put the two nouns together thus:

[هَـٰذَا ٱَلرَّجُلُ]{.ar}  
[hApA -rrujulu]{.trn}  
"this one, the man" = "this man"

We can also switch the order of the two nouns with the same meaning:

[ٱَلرَّجُلُ هَـٰذَا]{.ar}  
[Earujulu hApA]{.trn}  
"the man, this one" = "this man"

Both of these orders are permissible with the same meaning with the first order ([هَـٰذَا ٱَلرَّجُلُ]{.ar}) being more common.

These noun-phrases are very similar to the ones we learned for desccribers. In fact, we can add a describer to the noun-phrases above:

[هَـٰذَا ٱَلرَّجُلُ ٱلْكَرِيمُ]{.ar}  
[hApA -rrujulu -lkarImu]{.trn}  
"this one, the noble man" = "this noble man"

[ٱَلرَّجُلُ ٱلْكَرِيمُ هَـٰذَا]{.ar}  
[Earujulu -lkarImu hApA]{.trn}  
"the noble man, this one" = "this noble man"

These noun-phrases can be used in sentences too:

[هَـٰذَا ٱَلرَّجُلُ ٱلْكَرِيمُ عَظِيمٌ.]{.ar}  
[hApA -rrujulu -lkarImu eaPImun.]{.trn}  
"This one, the noble man, is great" = "This noble man is great."

Here are some more examples with some of the other pointing nouns:

+ [ضَرَبَتْ هَـٰذِهِ ٱلْمَرْأَةُ تِلْكَ ٱلْجَارِيَةَ.]{.ar}  
+ [قَرَأَ ذَ ٰنِكَ ٱلرَّجُلَانِ تَيْنِكَ ٱلرِّسَالَتَيْنِ ٱلطَّوِيلَتَيْنِ.]{.ar}
+ [هَاتَانِ ٱلْهِرَّتَانِ ٱلصَّغِيرَتَانِ لَكُمَا وَهَـٰذَانِ ٱلْكَلْبَانِ لَنَا.]{.ar}
+ [هَـٰؤُلَاءِ ٱلرِّجَالُ وَأُولَـٰئِكَ ٱلْغِلْمَانُ خَرَجُوا مِنَ ٱلْمَدْرَسَةِ.]{.ar}

### Independent pointing nouns in subject-information sentences

A pointing noun is considered definite even though it does not have [ٱَلْ]{.ar}. It can be used as a subject in subject information sentences. Here are some examples:

[هَـٰذَا رَجُلٌ]{.ar}  
[hApA rajulun]{.trn}  
"This is a man."

Such phrases consist of a pointing noun ("this") and the noun being pointed to ("man"). 

In Arabic, pointing nouns

In Arabic we cannot express a pointed-to noun like "man". We have to say either [رَجُلٌ]{.ar} [rajulun]{.trn} "a man" or [ٱَلرَّجُلُ]{.ar} [Earrujulu]{.trn} "the man". We choose the definite noun [ٱَلرَّجُلُ]{.ar} [Earrujulu]{.trn}

Both the pointing noun and the noun being pointed to are considered definite.

In English, the noun being pointed to does not have the word "the". However, in Arabic the noun being pointed to will have [ٱَلْ]{.ar}. 

Also, in English, the pointing noun comes before the noun being pointed to. In Arabic, the pointing noun may come before or 

### In subject-information sentences

A pointing noun is considered definite even though it does not have [ٱَلْ]{.ar}. It can be used as a subject in subject information sentences. Here are some examples:

+ 
[هَـٰذَا رَجُلٌ]{.ar}   [hApA]{.trn}     
[هَـٰذِهِ]{.ar}   [hApihi]{.trn}   
[هَـٰذَانِ]{.ar}  [hApAni]{.trn}   
[هَاتَانِ]{.ar}  [hAtAni]{.trn}   
[هَـٰؤُلَاءِ]{.ar} [hAEulAEi]{.trn} 
[ذَ ٰلِكَ]{.ar}   [pAlika]{.trn}   
[تِلْكَ]{.ar}    [tilka]{.trn}    
[ذَ ٰنِكَ]{.ar}   [pAnika]{.trn}   
[تَانِكَ]{.ar}   [tAnika]{.trn}   
[أُولَـٰئِكَ]{.ar} [EulAEika]{.trn} 










## TODO

hunA hunAka

hA huwa pA
hA hum EulAEik


