# Pointing nouns

## Introduction

Consider the following expression:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=this book,
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize pointing noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize pointed-to noun};

\end{tikzpicture}
```

The word "this" is what we will call a _pointing noun_. We call it this because we can imagine standing next to a book and pointing to it and saying "this book".

The word "book" here is similarly called the _pointed-to_ noun. It refers to the object being pointed to.

## The pointing nouns in Arabic

There are two types of pointing nouns:

i. Near pointing nouns: "this-one" (singular) and "these-ones" (dual and plural).
i. Far  pointing nouns: "that-one" (singular) and "those-ones" (dual and plural).

The following are the pointing nouns in Arabic:

| Participant |State| Near pointing  noun || Far pointing  noun ||
|:----|:-|--:|:----|--:|:----|
|sing. masc.|all|[هَـٰذَا]{.ar}  |this one~m~    |[ذَ ٰلِكَ]{.ar}  |that one~m~  |
|sing. fem. |all|[هَـٰذِهِ]{.ar}  |this one~f~    |[تِلْكَ]{.ar}   |that one~f~  |
|dual masc. |u  |[هَـٰذَانِ]{.ar} |these ones~2,m~|[ذَ ٰنِكَ]{.ar}  |those ones~2,m~|
|dual masc. |a,i|[هَـٰذَيْنِ]{.ar} |these ones~2,m~|[ذَيْنِكَ]{.ar}  |those ones~2,m~|
|dual fem.  |u  |[هَاتَانِ]{.ar} |these ones~2,f~|[تَانِكَ]{.ar}  |those ones~2,f~|
|dual fem.  |a,i|[هَاتَيْنِ]{.ar} |these ones~2,f~|[تَيْنِكَ]{.ar}  |those ones~2,f~|
|plural     |all|[هَـٰؤُلَاءِ]{.ar}|these ones~3~  |[أُولَـٰئِكَ]{.ar}|those ones~3~  |

Note the following:

+ Many of the pointing nouns contain small [Ealif]{.trn} [◌ٰ]{.ar}. For most of them, this is how they must be written. It would be incorrect to write [هَـٰذَا]{.ar} [hApA]{.trn} as [هَاذَا]{.ar}.
+ All the near pointing nouns begin with a [ه]{.ar}. And all the far pointing nouns end with [ك]{.ar}.
+ The [و]{.ar} in [أُولَـٰئِكَ]{.ar} [EulAEika]{.trn} is silent and not pronounced. That is, the first syllable has a short vowel [u]{.trn}, not the long vowel [U]{.trn}.
+ Most of the pointing nouns are rigid nouns. That is: their endings are not modified for their state. 

  The dual pointing nouns, however, are flexible nouns, for example: [هَـٰذَانِ]{.ar}  (u-state)   / [هَـٰذَيْنِ]{.ar}  [hApayni]{.trn} (a- and i-states).

+ The pointing nouns for the plural are the same for both masculine and feminine genders.

## Definiteness of pointing nouns

The pointing nouns share some similarities with pronouns [هُوَ]{.ar}, [هِيَ]{.ar}, etc. Just like pronouns, pointing nouns, too, are definite nouns even though they don't have [ٱَلْ]{.ar}.

Remember, however, from section\ \@ref(describers-with-annexations-to-pronouns), that pronouns may not be describees.
Pointing nouns are different from pronouns in this regard. It is allowed to describe a pointing noun with a describer in a noun phrase.

Both these facts will prove useful in the next section.

## Pointing noun for plurals of non-intelligent beings

Consistent with how we have been dealing with the so far,
, we can choose between the following pointing nouns for 
the plurals of non-intelligent beings:

||Near pointing noun |Far pointing noun|
|:----|:-|--:|:----|--:|:----|
|sing. fem. |all|[هَـٰذِهِ]{.ar}  |this one~f~    |[تِلْكَ]{.ar}   |that one~f~  |
|plural     |all|[هَـٰؤُلَاءِ]{.ar}|these ones~3~  |[أُولَـٰئِكَ]{.ar}|those ones~3~  |

The singular feminine pointing noun is usually preferred, unless the plural plural pointing noun is needed to indicate that there is more than one.
We will be giving examples throughout this chapter.

## The pointing noun phrase

Remember from chapter\ \@ref(adjectival-nouns-and-descriptive-noun-phrases) that a descriptive noun-phrase consists of a describer and a describee. The describer follows the describer and matches it in definiteness, state, gender, and number. 

Here is an example of a descriptive noun-phrase in a sentence.

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=.\txarb{ٱلصَّغِيرَينِ} \txarb{ٱلْكِتَابَيْنِ} \txarb{قَرَأْتُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-1.north west) -- (word-1.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize describer};

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize describee};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize descriptive noun-phrase};

\path (0,0) node [midway, below=7.5ex, right=0ex, align=left] {``I read the small books.'')};

\end{tikzpicture}
```

We will now see how this same descriptive noun-phrase can be used with pointing nouns.

### Pointing to a single noun

We will first deal with nouns that are single words, like [ٱَلْكِتَابَيْنِ]{.ar} above. 
In section\ \@ref(pointing-to-an-annexation)
below, we will deal with nouns that are part of an annexation, like [كِتَابَيِ ٱلرَّجُلِ]{.ar}.

#### The pointed-to noun is definite with [ٱَلْ]{.ar} {#phrase-single-pointed-to-noun-with-al}

Just like an adjectival noun, a pointing noun can be a describer in a noun-phrase. But remember from 
section\ \@ref(definiteness-of-pointing-nouns)
above that pointing nouns are definite.
So, if a pointing noun is a describer in a noun-phrase, the describee has to be definite too.
Example:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=.\txarb{هَـٰذَيْنِ} \txarb{ٱلْكِتَابَيْنِ} \txarb{قَرَأْتُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-1.north west) -- (word-1.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize describer};

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize describee};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-2.south east)
      node [midway, below] {\scriptsize descriptive noun-phrase};

\path (0,0) node [midway, below=10ex, right=0ex, align=left] {``I read these books.''\\(literally: ``I read the these-ones books.'')};

\end{tikzpicture}
```

In the above example, the pointed-to noun [ٱَلْكِتَابَيْنِ]{.ar} is the describee in a descriptive noun-phrase. It is definite, in the a-state, masculine, and dual.

The pointing noun [هَـٰذَيْنِ]{.ar} is its describer. It follows the describee and matches it being dual, in the a-state, masculine, and dual.

As a special case, when the pointed-to noun has [ٱَلْ]{.ar} (as in this case: [ٱَلْكِتَابَيْنِ]{.ar}), then the order of the pointing noun and the pointed to noun is permitted to be reversed. 
<!--The pointing noun can be the describee, and the pointed-to noun can be the describer.-->

The pointing noun is then a replacee (see section\ \@ref(the-replacement)), and the pointed-to noun is its replacement.

Example:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=.\txarb{ٱلْكِتَابَيْنِ} \txarb{هَـٰذَيْنِ} \txarb{قَرَأْتُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-1.north west) -- (word-1.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize replacement};

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize replacee};

\path (0,0) node [midway, below=7.5ex, right=0ex, align=left] {``I read these books.''\\(literally: ``I read the books: these-ones.'')};

\end{tikzpicture}
```

In the above example, the 
pointing noun [هَـٰذَيْنِ]{.ar} 
is a replacee. It is definite, in the a-state, masculine, and dual.

The 
pointed-to noun [ٱَلْكِتَابَيْنِ]{.ar} 
is its replacement. It follows the replacee and matches it being dual, in the a-state, masculine, and dual.

As a matter of fact, even though both orders are permitted, this reverse order of placing the pointing noun first and following it with the pointed-to noun is more common.

Here are some more examples of pointing noun phrases when the pointed-to noun is definite with [ٱَلْ]{.ar}:

[هَـٰذَا ٱَلرَّجُلُ ٱلْكَرِيمُ إِمَامٌ.]{.ar}  
[ٱَلرَّجُلُ ٱلْكَرِيمُ هَـٰذَا إِمَامٌ.]{.ar}  
"This noble man is an [imAm]{.trn2}."

#### The pointed-to noun is a proper noun

Remember that proper noun are definite nouns, even though they usually don't begin with [ٱَلْ]{.ar}. For example:

|||||
|--:|:---|--:|:---|
|[زَيْد]{.ar}     |Zayd   |[ٱَلْحَارِث]{.ar}  |[al-#HAriv]{.trn2}|
|[زَيْنَب^2^]{.ar} |Zaynab |[قُرَيْش]{.ar}    |[#qurayc]{.trn2}  |

Such names may also be part of a pointing noun phrase. 
If they don't begin with [ٱَلْ]{.ar} then only the [pointed-to noun first, then pointing noun] order is permitted.
Example:

[زَيْدٌ هَـٰذَا أَخُو زَيْنَبَ تِلْكَ.]{.ar}  
"This Zayd is that Zaynab's brother."

[قُرَيْشٌ هَـٰؤُلَاءِ سَكَنُوا بِمَكَّةَ.]{.ar}  
"These [#qurayc]{.trn2} dwelled in Mecca."

If the name begins with [ٱَلْ]{.ar} then both orders are permitted.

[هَـٰذَا ٱلْحَارِث]{.ar}  
[ٱلْحَارِث هَـٰذَا]{.ar}  
"this [al-#HAriv]{.trn2}"

<!--Wright, ii Sect 136 277B-->

### Pointing to an annexation

Consider the following expression:

"the man's book"

We can apply the pointing noun "this" to either "the book" or to "the man" in a pointing noun phrase. So we have two options:

i. "the book of this man"
i. "this book of the man"

Similarly, consider the following expression:

"Zayd's book"

We can, again, apply the pointing noun "this" to either "the book" or to "Zayd":

i. "the book of this Zayd"
i. "this book of Zayd"

In this section we will learn how to construct these pointing noun phrases in Arabic.
Arabic uses annexations to express the above meanings. So we will discuss annexations like:

[كِتَابُ ٱلرَّجُلِ]{.ar}  
"the book of the man"

and

[كِتَابُ زَيْدٍ]{.ar}  
"the book of Zayd"

Note that both the above annexations are definite because their base nouns are definite.

Indefinite annexations like 
[كِتَاب رَجُلٍ]{.ar} "a man's book"
cannot be used in pointing noun phrases.

#### The definite base noun begins with [ٱَلْ]{.ar}

We will first consider annexations where the definite base noun begins with [ٱَلْ]{.ar}, like:

[كِتَابُ ٱلرَّجُلِ]{.ar}  
"the book of the man"

##### Pointing to the base noun  

We would like to express the phrase:

"the book of this man"

In order to point to the base noun [ٱَلرَّجُل]{.ar} "the man" with the pointing noun [هَـٰذَا]{.ar} "this-one~m~", we can put the pointing noun either before or after the base noun, thus:

[كِتَابُ هَـٰذَا ٱلرَّجُلِ]{.ar}  
[كِتَابُ ٱلرَّجُلِ هَـٰذَا]{.ar}  
"the book of this man"

Both these pointing noun phrases give the same meaning: "the book of this man".
However, the first phrase
[كِتَابُ هَـٰذَا ٱلرَّجُلِ]{.ar}
is preferred, consistent with what we learned in 
section\ \@ref(phrase-single-pointed-to-noun-with-al), above.

The second phrase
[كِتَابُ ٱلرَّجُلِ هَـٰذَا]{.ar},
although correct, would only rarely be used with this meaning.
(In fact, it has another meaning: "this book of the man" which we will learn in
section\ \@ref(pointing-to-the-annexe-noun), below.)

Here is how these phrases could be used in complete sentences:

[كِتَابُ هَـٰذَا ٱلرَّجُلِ جَدِيدٌ.]{.ar}  
[كِتَابُ ٱلرَّجُلِ هَـٰذَا جَدِيدٌ.]{.ar}  
"The book of this man is new."

Before we give more examples, let's analyze these phrases in detail.

Consider the first pointing noun phrase:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=.\txarb{ٱلرَّجُلِ} \txarb{هَـٰذَا} \txarb{كِتَابُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize base-noun};

\draw [decoration={brace}, decorate] 
   (word-3.north west) -- (word-3.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize annexe-noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize replacement};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize replacee};

\path (0,0) node [midway, below=17ex, right=0ex, align=left] {``the book of this man''\\(literally: ``the book of this-one: the man'')};

\end{tikzpicture}
```

As you can see the pointing noun
[هَـٰذَا]{.ar} has taken the place of 
[ٱَلرَّجُل]{.ar} as the base noun in the annexation.
In addition to being the base noun,
[هَـٰذَا]{.ar} is also a replacee, whose replacement is 
[ٱَلرَّجُل]{.ar}.
The literal, word-for-word, translation of this phrase is:

"the book of this-one: the man"

The more natural translation is:

"the book of this man"

Consider, now, the second pointing noun phrase:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=.\txarb{هَـٰذَا} \txarb{ٱلرَّجُلِ} \txarb{كِتَابُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize base-noun};

\draw [decoration={brace}, decorate] 
   (word-3.north west) -- (word-3.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize annexe-noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize describer};

\draw [decoration={brace, mirror}, decorate] 
   (word-2.south west) -- (word-2.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize describee};

\path (0,0) node [midway, below=15ex, right=0ex, align=left] {``the book of this man''\\(literally: ``the book of the this-one man'')};

\end{tikzpicture}
```

[ٱَلرَّجُل]{.ar}, here, keeps its place as the base noun in the annexation.
In addition to being the base noun,
[ٱَلرَّجُل]{.ar}
is also a describee, whose describer is the pointing noun
[هَـٰذَا]{.ar}.
The literal, word-for-word, translation of this phrase is:

"the book of the this-one man"

The more natural translation is:

"the book of this man"

##### Pointing to the annexe noun  

Consider, again, the annexation:

[كِتَابُ ٱلرَّجُلِ]{.ar}  
"the book of the man"

We have already discussed how to point to the base noun
[ٱَلرَّجُل]{.ar}
in a pointing noun phrase.
Now, we would like to point to the annexe noun 
[كِتَاب]{.ar}
in a pointing noun phrase.

In other words, we would like to express the meaning:

"this book of the man"

<!--
There are three possible locations to place the pointing noun
[هَـٰذَا]{.ar}:

i. $\times$ [هَـٰذَا كِتَابُ ٱلرَّجُلِ]{.ar},
i. $\times$ [كِتَابُ هَـٰذَا ٱلرَّجُلِ]{.ar},
i. [كِتَابُ ٱلرَّجُلِ هَـٰذَا]{.ar},

The first two options are incorrect, and the third option is correct. We will explain why below.
-->

The way to express this in Arabic is

[كِتَابُ ٱلرَّجُلِ هَـٰذَا]{.ar}  
"this book of the man"

But wait! Didn't we see in section\ \@ref(pointing-to-the-base-noun) above that this expression has the meaning
"the book of this man"?

It turns out that this expression supports both meanings.

But it will generally only be used for the meaning: "this book of the man"

In order to express
"the book of this man"
we will typically use the expression [كِتَابُ هَـٰذَا ٱلرَّجُلِ]{.ar}.

Let's analyze the expression
[كِتَابُ ٱلرَّجُلِ هَـٰذَا]{.ar}
"this book of the man"
in detail:

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=.\txarb{هَـٰذَا} \txarb{ٱلرَّجُلِ} \txarb{كِتَابُ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize base-noun};

\draw [decoration={brace}, decorate] 
   (word-3.north west) -- (word-3.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize annexe-noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize describer};

\draw [decoration={brace, mirror}, decorate] 
   (word-3.south west) -- (word-3.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize describee};

\path (0,0) node [midway, below=15ex, right=0ex, align=left] {``this book of the man''\\(literally: ``the this-one book of the man'')};

\end{tikzpicture}
```

[كِتَاب]{.ar}, here, is both and annexe noun and a describee.
Its describer is the pointing noun
[هَـٰذَا]{.ar}.
The literal, word-for-word, translation of this phrase is:

"the this-one book of the man"

The more natural translation is:

"this book of the man"

Here is this pointing noun phrase in a complete sentence:

[كِتَابُ ٱلرَّجُلِ هَـٰذَا أَخْضَر.]{.ar}  
"This book of the man is green."

###### Ambiguity of this phrase {-}

A quick note about the ambiguity of this expression:

[كِتَابُ ٱلرَّجُلِ هَـٰذَا]{.ar}  
"this book of the man" (usual)  
"the book of this man" (rare)

The ambiguity of whether the pointing noun [هَـٰذَا]{.ar} points to the annexe noun 
[كِتَابُ]{.ar}
or the base noun
[ٱلرَّجُلِ]{.ar}
only exists because the annexe noun and the base noun match each other in gender and number: singular masculine.
If the annexe noun and the base noun were different in gender and number, then there would be no ambiguity. Examples:

[كِتَابَا ٱلرَّجُلِ هَـٰذَانِ]{.ar}  
"these books~2~ of the man"

[كِتَابُ ٱلرَّجُلَيْنِ هَـٰذَا]{.ar}  
"this book of the men~2~"

[كِتَابُ ٱلْمَرْأَةِ هَـٰذَا]{.ar}  
"this book of the woman"

[كِتَابُ ٱلْمَرْأَةِ هَـٰذِهِ]{.ar}  
"the book of this woman"

Here are some more examples of pointing to annexe nouns:

##### The base noun is a proper noun beginning with [ٱَلْ]{.ar}  

Consider the annexation:

[كِتَابُ ٱلزُّبَيْرِ]{.ar}  
"the book of [al-#zubayr]{.trn2}"

We can apply the preceding discussion of pointing to the annexe noun and base noun to this annexation as well. So we get:

[كِتَابُ هَـٰذَا ٱلزُّبَيْرِ]{.ar}  
"this book of al-Zubayr"

[كِتَابُ ٱلزُّبَيْرِ هَـٰذَا]{.ar}  
"this book of al-Zubayr" (usual)  
"the book of this al-Zubayr" (rare)

#### The definite base noun does not begin with [ٱَلْ]{.ar}

Consider, now, that the base noun is definite but does not begin with [ٱَلْ]{.ar}. There are two such types of nouns that we will discuss:

i. Proper nouns not beginning with [ٱَلْ]{.ar}
i. Pronouns

##### The base noun is a proper noun not beginning with [ٱَلْ]{.ar}  

<!--Consider, now, that the base noun is a proper noun that does not begin with [ٱَلْ]{.ar}, for example:-->

We will first deal with proper nouns  that don't begin with [ٱَلْ]{.ar}.
Consider the annexation:

[كِتَابُ زَيْدٍ]{.ar}  
"the book of Zayd"

Because the base noun [زَيْد]{.ar} does not begin with [ٱَلْ]{.ar}, any pointing nouns can come only after the entire annexation, thus:

[كِتَابُ زَيْدٍ هَـٰذَا]{.ar}

In theory, this supports two meanings:

i. "this book of Zayd"
i. "the book of this Zayd"

In practice, however, the first meaning 
("this book of Zayd")
is much more likely.
Pointing to a proper noun in a pointing noun phrase 
("the book of this Zayd")
is uncommon, generally.

##### The base noun is a pronoun

We have learned, in 
section\ \@ref(definiteness-of-pronouns),
that pronouns are always definite, despite not beginning with [ٱَلْ]{.ar}.

We have also learned, in
section\ \@ref(pronouns-as-base-nouns),
that a pronoun may be a base noun in an annexation.
Example:

[كِتَابُهُ]{.ar}  
"his book"

Neither the annexe noun [كِتَاب]{.ar}, nor the attached pronoun [هُ]{.ar} begin with [ٱَلْ]{.ar}.
So if we want to add the pointing noun [هَـٰذَا]{.ar} to this annexation to form a pointing noun phrase,
then we have to place it at the end, after the annexation, thus:

[كِتَابُهُ هَـٰذَا]{.ar}  

The pointing noun [هَـٰذَا]{.ar}, here, is a describee. But what is its describer?

We have also learned, in
section\ \@ref(describers-with-annexations-to-pronouns)
that pronouns may not be describees in a descriptive noun phrase.

So, we are left with only one option: the annexe noun [كِتَاب]{.ar} is the desceibee. And the meaning of the phrase is:

[كِتَابُهُ هَـٰذَا]{.ar}  
"this book of his"

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=\txarb{هَـٰذَا} \txarb{ـهُ} \txarb{كِتَابُـ},
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace}, decorate] 
   (word-2.north west) -- (word-2.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize base-noun};

\draw [decoration={brace}, decorate] 
   (word-3.north west) -- (word-3.north east)
      node [midway, rotate=90, anchor=south, right=0ex] {\scriptsize annexe-noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize describer};

\draw [decoration={brace, mirror}, decorate] 
   (word-3.south west) -- (word-3.south east)
      node [midway, rotate=90, anchor=south, left=0ex] {\scriptsize describee};

\path (0,0) node [midway, below=15ex, right=0ex, align=left] {``this book of his''};

\end{tikzpicture}
```


Here are some more examples:



[كِتَابُ ٱلرَّجُلِ]{.ar}  

We wwold like to be able to point to either the annexe noun
[كِتَابَا]{.ar} "books~2~"
or the base noun
[ٱلرَّجُلِ]{.ar} "the man"
in a pointing noun phrase.

By the way, we can only do this when then base noun (and therefore the annexe noun also) is definite.
If the base noun and annexe nouns are indefinite, then they may not be described with poinitng nouns because pointing nouns are always definite.

#### The pointed-to noun is the base noun

If the pointed-to noun is the base noun, then we may place

## Remove

```{r,echo=FALSE,engine='tikz', engine.opts = list(template = "srctex/tikz2pdf-lua.tex")}
\usetikzlibrary{decorations.text, decorations.pathreplacing}
\begin{tikzpicture}[nodes={text depth=0.25ex,text height=1.25ex}]
\path [decoration={text effects along path, 
  text=This is a book,
  text effects/.cd, 
    path from text, text along path,
    group letters, word count=\w,
    every word/.style={name=word-\w, execute at begin node=\strut}},
  decorate] (0,0);

\draw [decoration={brace, mirror}, decorate] 
   (word-1.south west) -- (word-1.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize pointing noun};

\draw [decoration={brace, mirror}, decorate] 
   (word-4.south west) -- (word-4.south east)
      node [midway, below=2ex, left=0.5ex, rotate=90] {\scriptsize pointed-to noun};

\end{tikzpicture}
```


Pointing nouns are the Arabic equivalents of words like "this", "that", "these", "those". We call them pointing nouns because a person saying "This man.", for example, could be pointing to a man near him. Pointing nouns are used in sentences like:

+ "This is a house."
+ "These women are good."
+ "That is a book."

The pointing nouns in Arabic are:


Note the following regarding the pointing nouns shown above:

+ Many of the pointing nouns contain small [Ealif]{.trn} [◌ٰ]{.ar}. For most of them, this is how they must be written. It would be incorrect to write [هَـٰذَا]{.ar} [hApA]{.trn} as [هَاذَا]{.ar}.
+ The [و]{.ar} in [أُولَـٰئِكَ]{.ar} [EulAEika]{.trn} is silent and not pronounced. That is, the first syllable has a short vowel [u]{.trn}, not the long vowel [U]{.trn}.
+ The pointing nouns are the same in all states except for the dual, where [ـَانِ]{.ar} [-Ani]{.trn} in the u-state becomes [ـَيْنِ]{.ar} [-ayni]{.trn} in the a-state and i-state.
+ The pointing nouns for the plural are the same for both masculine and feminine genders.

### Using pointing nouns with nouns or phrases not beginning with [ٱَلْ]{.ar}

If a pointing noun comes before a noun or phrase that does not begin with [ٱَلْ]{.ar} then the expression is a complete sentence. The pointing noun is the subject of sentence, and the other noun or phrase is the information of the sentence. Here are some examples:

+ [هَـٰذَا رَجُلٌ.]{.ar}  
  [hApA rajulun.]{.trn}  
  "This~masc.~ is a man."
+ [هَـٰذَانِ مُعَلِّمَانِ كَرِيمَانِ.]{.ar}  
  [hApAni mueallimAni karImAni.]{.trn}  
  "These~dual\ masc.~ are noble (male) teachers."
+ [هَـٰؤُلَاءِ مُعَلِّمُو غُلَامٍ.]{.ar}  
  [hAEulAEi mueallimU gulAmin.]{.trn}  
  "These~pl.~ are (male) teachers of a boy."
+ [تِلْكَ فَاطِمَةُ.]{.ar}  
  [tilka fATimatu.]{.trn}  
  "That~fem.~ is [#faTimah]{.trn2}."
+ [تَانِكَ مُعَلِّمَتَا ٱلْجَارِيَةِ.]{.ar}  
  [tAnika mueallimata -ljAriyati.]{.trn}  
  "Those~dual\ fem.~ are the (female) teachers of the girl."
+ [هَـٰذِهِ رِسَالَتِي.]{.ar}  
  [hApihi risAlatI.]{.trn}  
  "This is my message."

In the first three examples above, the information of the sentence is an indefinite noun or noun-phrase. 
In the fourth example, the information is a proper noun which is definite without [ٱَلْ]{.ar}.
In the last example, the information is a noun-chain whose last noun is definite. Therefore the first noun is definite as well, although, again, without [ٱَلْ]{.ar}.

So you can see that the rule we have stated, that if a pointing noun comes before a noun or phrase that does not begin with [ٱَلْ]{.ar} then the expression is a complete sentence, that this rule applies regardless of whether the noun or phrase is definite or indefinite.

## Using pointing nouns with nouns or phrases beginning with [ٱَلْ]{.ar}

### The pointing noun before the other noun

If a pointing noun comes before a noun that begins with [ٱَلْ]{.ar}, for example,  
[هَـٰذَا ٱَلرَّجُلُ]{.ar}  
[hApA -rrujulu]{.trn},  
then the expression can either:

 i. be a noun phrase: "this man", or
ii. be a complete sentence: "This is the man."

Here is another example:

[هَـٰذَانِ ٱلْمُعَلِّمَانِ ٱلْكَرِيمَانِ.]{.ar}  
[hApAni -lmueallimAni -lkarImAni.]{.trn}  
"These~dual\ masc.~ noble (male) teachers"  
or  
"These~dual\ masc.~ are the noble (male) teachers."

Context would be needed to determine which of the two meanings is intended.

In order to disambiguate the complete sentence from the noun phrase, the detached pronoun is often (but not always) placed between the pointing noun and the information of the sentence thus:

[هَـٰذَانِ هُمَا ٱلْمُعَلِّمَانِ ٱلْكَرِيمَانِ.]{.ar}  
[hApAni huma -lmueallimAni -lkarImAni.]{.trn}  
"These~dual\ masc.~ are the noble (male) teachers"  

Here are some examples in complete sentences:

+ [قَرَأْتُ ذَيْنِكَ ٱلْكِتَابَيْنِ.]{.ar}  
  [qaraEtu paynika -lkitAbayni.]{.trn}  
  "I read those two books."

### The pointing noun after the other noun

If a pointing noun comes after a noun that begins with [ٱَلْ]{.ar}, for example,  
[ٱَلرَّجُلُ هَـٰذَا]{.ar}  
[Earrujulu hApA]{.trn},  
then the expression can only be a noun-phrase: "this man". 

Despite the disambiguity in expressing the noun-phrase with this second method, the first method of placing the pointing noun before the other noun is more commonly used. That is, in order to express "this man",  
[هَـٰذَا ٱَلرَّجُلُ]{.ar}  
[hApA -rrujulu]{.trn}  
is more commonly used than  
[ٱَلرَّجُلُ هَـٰذَا]{.ar}  
[Earrujulu hApA]{.trn}.

Here are some examples in complete sentences:

## Pointing nouns in noun-phrases with definite nouns not beginning with [ٱَلْ]{.ar}

It remains to see how to express pointing nouns in phrases with defiinite nouns not beginning with [ٱَلْ]{.ar}. Such nouns include:

 i. proper nouns, and
ii. the first noun of a noun-chain whose last noun is definite.

In other words, we would like to express phrases like:

+ "that [#fATimah]{.trn2}", used for example in the sentence "This [#fATimah]{.trn2} is tall and that [#fATimah]{.trn2} is short."
+ "these~pl.~ (female) teachers of the girl"
+ "this message of mine"

We know that we cannot place the pointing noun in the beginning of the phrase because, as we saw, that would make it a complete sentence. For example,

+ [تِلْكَ فَاطِمَةُ.]{.ar}  
  [tilka fATimatu.]{.trn}  
  "That~fem.~ is [#faTimah]{.trn2}."
+ [تَانِكَ مُعَلِّمَتَا ٱلْجَارِيَةِ.]{.ar}  
  [tAnika mueallimata -ljAriyati.]{.trn}  
  "Those~dual\ fem.~ are the (female) teachers of the girl."
+ [هَـٰذِهِ رِسَالَتِي.]{.ar}  
  [hApihi risAlatI.]{.trn}  
  "This is my message."

So, we are left with placing the pointing noun at the end of the phrase, thus:

+ [فَاطِمَةُ تِلْكَ.]{.ar}  
  [fATimatu tilka]{.trn}  
  "that~fem.~ [#faTimah]{.trn2}"
+ [مُعَلِّمَتَا ٱلْجَارِيَةِ تَانِكَ]{.ar}  
  [mueallimata -ljAriyati tAnika]{.trn}  
  "those~dual\ fem.~ (female) teachers of the girl"
+ [رِسَالَتِي هَـٰذِهِ]{.ar}  
  [risAlatI hApihi]{.trn}  
  "this message of mine"

Here are some examples in complete sentences:

+ [فَاطِمَةُ هَـٰذِهِ طَوِيلَةٌ وَفَاطِمَةُ تِلْكَ قَصِيرَةٌ.]{.ar}  
  [fATimatu hApihi TawIlatun wafATimatu tilka qaSIratun.]{.trn}  
  "This [#fATimah]{.trn2} is tall and that [#fATimah]{.trn2} is short."

## Using pointing nouns with plural non-rational beings

### REMOVE: In noun-phrases with common nouns

A pointing noun can be used in a noun phrase for expressions like "this man". We can express this using the pointing noun 
[هَـٰذَا]{.ar} [hApA]{.trn} "this~masc.~ one" and the noun being pointed to [ٱَلرَّجُلُ]{.ar} [Earrujulu]{.trn} "the man". We put the two nouns together thus:

[هَـٰذَا ٱَلرَّجُلُ]{.ar}  
[hApA -rrujulu]{.trn}  
"this one, the man" = "this man"

We can also switch the order of the two nouns with the same meaning:

[ٱَلرَّجُلُ هَـٰذَا]{.ar}  
[Earujulu hApA]{.trn}  
"the man, this one" = "this man"

Both of these orders are permissible with the same meaning with the first order ([هَـٰذَا ٱَلرَّجُلُ]{.ar}) being more common.

These noun-phrases are very similar to the ones we learned for desccribers. In fact, we can add a describer to the noun-phrases above:

[هَـٰذَا ٱَلرَّجُلُ ٱلْكَرِيمُ]{.ar}  
[hApA -rrujulu -lkarImu]{.trn}  
"this one, the noble man" = "this noble man"

[ٱَلرَّجُلُ ٱلْكَرِيمُ هَـٰذَا]{.ar}  
[Earujulu -lkarImu hApA]{.trn}  
"the noble man, this one" = "this noble man"

These noun-phrases can be used in sentences too:

[هَـٰذَا ٱَلرَّجُلُ ٱلْكَرِيمُ عَظِيمٌ.]{.ar}  
[hApA -rrujulu -lkarImu eaPImun.]{.trn}  
"This one, the noble man, is great" = "This noble man is great."

Here are some more examples with some of the other pointing nouns:

+ [ضَرَبَتْ هَـٰذِهِ ٱلْمَرْأَةُ تِلْكَ ٱلْجَارِيَةَ.]{.ar}  
+ [قَرَأَ ذَ ٰنِكَ ٱلرَّجُلَانِ تَيْنِكَ ٱلرِّسَالَتَيْنِ ٱلطَّوِيلَتَيْنِ.]{.ar}
+ [هَاتَانِ ٱلْهِرَّتَانِ ٱلصَّغِيرَتَانِ لَكُمَا وَهَـٰذَانِ ٱلْكَلْبَانِ لَنَا.]{.ar}
+ [هَـٰؤُلَاءِ ٱلرِّجَالُ وَأُولَـٰئِكَ ٱلْغِلْمَانُ خَرَجُوا مِنَ ٱلْمَدْرَسَةِ.]{.ar}

### Independent pointing nouns in subject-information sentences

A pointing noun is considered definite even though it does not have [ٱَلْ]{.ar}. It can be used as a subject in subject information sentences. Here are some examples:

[هَـٰذَا رَجُلٌ]{.ar}  
[hApA rajulun]{.trn}  
"This is a man."

Such phrases consist of a pointing noun ("this") and the noun being pointed to ("man"). 

In Arabic, pointing nouns

In Arabic we cannot express a pointed-to noun like "man". We have to say either [رَجُلٌ]{.ar} [rajulun]{.trn} "a man" or [ٱَلرَّجُلُ]{.ar} [Earrujulu]{.trn} "the man". We choose the definite noun [ٱَلرَّجُلُ]{.ar} [Earrujulu]{.trn}

Both the pointing noun and the noun being pointed to are considered definite.

In English, the noun being pointed to does not have the word "the". However, in Arabic the noun being pointed to will have [ٱَلْ]{.ar}. 

Also, in English, the pointing noun comes before the noun being pointed to. In Arabic, the pointing noun may come before or 

### In subject-information sentences

A pointing noun is considered definite even though it does not have [ٱَلْ]{.ar}. It can be used as a subject in subject information sentences. Here are some examples:

+ 
[هَـٰذَا رَجُلٌ]{.ar}   [hApA]{.trn}     
[هَـٰذِهِ]{.ar}   [hApihi]{.trn}   
[هَـٰذَانِ]{.ar}  [hApAni]{.trn}   
[هَاتَانِ]{.ar}  [hAtAni]{.trn}   
[هَـٰؤُلَاءِ]{.ar} [hAEulAEi]{.trn} 
[ذَ ٰلِكَ]{.ar}   [pAlika]{.trn}   
[تِلْكَ]{.ar}    [tilka]{.trn}    
[ذَ ٰنِكَ]{.ar}   [pAnika]{.trn}   
[تَانِكَ]{.ar}   [tAnika]{.trn}   
[أُولَـٰئِكَ]{.ar} [EulAEika]{.trn} 










## TODO

hunA hunAka

hA huwa pA
hA hum EulAEik


