---
title: "Learn Standard Arabic"
subtitle:  "A self-instruction textbook with grammar, vocabulary, and exercises"
author: "Ahmad Sheikh"
date: "`r system('git describe', intern = TRUE)`"
documentclass: book
geometry:
# A4 2 pages per sheet draft
#- paper=a5paper               # a5: 148.5 by 210mm
#- paperwidth=170mm
#- paperheight=244mm
- paperwidth=156mm
- paperheight=234mm
#- paperwidth=6in
#- paperheight=9in
- bindingoffset=16mm
- textwidth=114.8mm           # = (170 - 6)*0.7
- textheight=170.8mm          # = 244 * 0.7
- twoside
mainfont: Brill
mainfontoptions: Numbers=OldStyle
fontsize: 10pt
polyglossia-lang:
  name: en
polyglossia-otherlangs:
  - name: arabic
fontfamilies:
  - name: \arabicfont
    font: Vazirmatn-Light
    options: Script=Arabic,Scale=1.0
  - name: \tradarab
    font: Amiri
    options: Script=Arabic,Scale=1.0
  - name: \tradarabsmall
    font: Amiri
    options: Script=Arabic,Scale=0.75
csquotes: true
output:
  bookdown::gitbook:
    pandoc_args: ["--lua-filter=trn.lua"]
    keep_md: yes
    css: mystyle.css
    number_sections: true
    config:
      toc:
        collapse: section
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: null
      search: yes
      fontsettings:
        theme: white
        family: serif
        size: 2
      sharing:
        facebook: yes
        github: no
        twitter: yes
        linkedin: yes
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
      info: yes
  bookdown::pdf_book:
    number_sections: true
    latex_engine: xelatex
    pandoc_args: ["--lua-filter=trn.lua", "--columns=65"] # for forcing word wrap in pipe tables
    #template: default_latex_template.tex # with Brill font the PDF font is too large and this needs to be uncommented. includes and in_header should then be commented
    includes:
      before_body: "frontpage.tex"
      in_header: "preamble.tex"
    keep_tex: true
---

# Preface {-}
<!--The following line is needed because Preface is an unnumbered chapter and otherwise fancy header will print the header text for the previous chapter "Contents"-->
\markboth{Preface}{Preface}

<!--to allow git describe to run-->
```{r, include=FALSE}
options(tinytex.engine_args = '-shell-escape')
```

```{r, eval=knitr::is_html_output(), echo=FALSE}
  knitr::asis_output('
[بسم الله الرحمن الرحيم]{lang="ar" dir="rtl" style="font-family: Amiri;"}
')
```
```{=latex}
\RL{\tradarab{بسم الله الرحمن الرحيم}}
```
<!--Sunnah are the sayings and actions of the Prophet صلى الله عليه وسلم. Hadith is the text by which we know the sunnah.-->

The primary texts of [#islAm]{.trn2} (the [#qurEAn]{.trn2} and the [#HadIv]{.trn2}) are in Arabic. So too is much of its scholarly literature. However, there is a multitude of Muslims for whom Arabic is not a native language, yet who are familiar enough with English to study textbooks written in this language. The goal of this book is to help them learn Arabic at a beginner's level so that, together with a study of the appropriate expositional texts, they are one step closer to understanding the primary texts in their original language. We hope that this will, if [#allAh]{.trn2} wills, make them feel more connected to the primary texts and their teachings. Furthermore, they can be empowered to study the vast body of Arabic [#islAm]{.trn2}ic literature.

This book is a teaching grammar and not a reference grammar. So, in the initial chapters, topics are presented sequentially at only a basic level, without treating them exhaustively, before moving on to the next topic. Furthermore, since this is a beginner's textbook, only the more common usages are explained.

We have also aimed to make this a self-instruction textbook so that a diligent student should, if [#allAh]{.trn2} wills, be able to study it without an instructor. The target learner is someone who has not been exposed to grammatical terminology like _inflection_, _case_, _mood_, etc. While terminology is necessary for a rigorous non-immersive learning of language, we have tried to steer away from Latin-based terms like _accusative_. Such a term, when first encountered by an uninitiated learner, can be sufficient to deter him from proceeding further. (Learning a language can be hard enough without getting the feeling that your grammar book is accusing you of something.) So we have in some places translated the meaning of Arabic grammar terms to English. In some places, we have used established English grammar terms where the terms are basic enough. We have even, in places, invented terms where we deemed appropriate. The drawback to this non-standard approach, however, is that the student may not be able to immediately relate the terminology he has learned in this book to established terminology in other grammar textbooks. To remedy this to some extent, we provide a glossary in the appendix which maps the grammatical terminology used in this book to other, established, Latin-based and Arabic-based counterparts.

It may also be appropriate to inform the reader that we chose to present a simplified version of Arabic grammar. As such, the grammar presented here may not be entirely consistent with the comprehensive and harmonious framework developed by the Arab grammarians. We chose this approach because we felt that exposing the beginner to complex grammatical details at this stage would be more of a hindrance than a help in learning the language.

```{r, eval=knitr::is_html_output(), echo=FALSE}
  knitr::asis_output('
[a.z.s.]{.smallcaps}  
[lahore, dec.]{.smallcaps} 2021[ce]{.smallcaps}  
[لاهور، جمادى الأولى ١٤٤٣هـ‍]{lang="ar" dir="rtl" style="font-family: Amiri; font-size: 0.75em;"}
')
```
```{=latex}
\medskip
\null\hfill\begin{tabular}{@{}l@{}}
\textsc{a.z.s.}\\
\textsc{lahore, dec. 2021ce}\\
\end{tabular}
\vspace*{-\baselineskip}
\vspace*{2mm}
\begin {flushright}
\RL{\tradarabsmall{لاهور، جمادى الأولى ١٤٤٣ه‍}}
\end {flushright}
```

Test:

This is the _preface_. Test trans [Parf makAn ealayhi pcv machad]{.trn} [◌َ  ◌ً  ◌ِ  ◌ٍ  ◌ُ  ◌ٌ  ◌ْ  ◌ّ  ◌ٰ  ◌ٔ  ◌ٕ  ]{.ar}

[شَيْـَٔيْنِ بَرِيـِٔينَ شَيْـًٔا خَطِيـَٔة]{.ar}

[شَيْءَيْنِ بَرِيءِينَ شَيْءًا خَطِيءَة]{.ar}


this is in [Small Caps]{.smallcaps}.

<!--```{=latex}
\lipsum[1-149]
```-->
<!--```{=latex}
\mainmatter
```-->
