---
title: "Learn Standard Arabic"
subtitle:  "A self-instruction textbook with grammar, vocabulary, and exercises"
author: "Author Names"
date: "`r system('git describe', intern = TRUE)`"
documentclass: book
geometry:
# A4 2 pages per sheet draft
#- paper=a5paper               # a5: 148.5 by 210mm
#- paperwidth=170mm
#- paperheight=244mm
- paperwidth=156mm
- paperheight=234mm
#- paperwidth=6in
#- paperheight=9in
- bindingoffset=16mm
- textwidth=114.8mm           # = (170 - 6)*0.7
- textheight=170.8mm          # = 244 * 0.7
- twoside
mainfont: "Charis SIL"
mainfontoptions: Numbers=OldStyle
fontsize: 10pt
lang: en
otherlangs: [ar]
fontfamilies:
  - name: \arabicfont
    font: Vazirmatn-Light
    options: Script=Arabic,Scale=1.0
  - name: \tradarab
    font: Amiri
    options: Script=Arabic,Scale=1.0
  - name: \tradarabsmall
    font: Amiri
    options: Script=Arabic,Scale=0.75
csquotes: true
output:
  bookdown::gitbook:
    includes:
      before_body: "srchtml/watermark.html"
    pandoc_args: ["--lua-filter=trn.lua"]
    keep_md: yes
    css: mystyle.css
    number_sections: true
    config:
      toc:
        collapse: section
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: null
      search: yes
      fontsettings:
        theme: white
        family: serif
        size: 2
      sharing:
        facebook: yes
        github: no
        twitter: yes
        linkedin: yes
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
      info: yes
  bookdown::pdf_book:
    number_sections: true
    latex_engine: xelatex
    pandoc_args: ["--lua-filter=trn.lua", "--columns=65"] # for forcing word wrap in pipe tables
    #template: default_latex_template.tex # with Brill font the PDF font is too large and this needs to be uncommented. includes and in_header should then be commented
    includes:
      before_body: "srctex/frontpage.tex"
      in_header: "srctex/preamble.tex"
    keep_tex: true
---

# Preface {-}

<!--The following line is needed because Preface is an unnumbered chapter and otherwise fancy header will print the header text for the previous chapter "Contents"-->
\markboth {\textsc{\MakeLowercase{Preface}}}{\textsc{\MakeLowercase{Preface}}}

<!--to allow git describe to run-->
```{r, include=FALSE}
options(tinytex.engine_args = '-shell-escape')
```

<!--
::: {.otherlanguage data-latex="{arabic}" lang="ar"}
[بسم الله الرحمن الرحيم]{.ar}
:::
-->

```{r, eval=knitr::is_html_output(), echo=FALSE}
knitr::asis_output('
[بسم الله الرحمن الرحيم]{lang="ar" dir="rtl" style="font-family: AmiriWeb;"}
')
```

<!-- RTL doesn't work for some reason-->
```{=latex}
\begin{center}
\tradarab{
الرحيم
الرحمن
الله
بسم
}
\end{center}
```
<!--Sunnah are the sayings and actions of the Prophet صلى الله عليه وسلم. Hadith is the text by which we know the sunnah.-->

The primary texts of [#islAm]{.trn2} (the [#qurEAn]{.trn2} and the [#HadIv]{.trn2}) are in Arabic. So too is much of its scholarly literature. However, there is a multitude of Muslims for whom Arabic is not a native language, yet who are familiar enough with English to study textbooks written in this language. The goal of this book is to help them learn Arabic at a beginner's level so that, together with a study of the appropriate expositional texts, they are one step closer to understanding the primary texts in their original language. We hope that this will, if [#allAh]{.trn2} wills, make them feel more connected to the primary texts and their teachings. Furthermore, they can be empowered to study the vast body of Arabic [#islAm]{.trn2}ic literature.

This book is a learning grammar and not a reference grammar. 
So, in the initial chapters, topics are introduced at only a basic level, without treating them exhaustively, before moving on to the next topic. 
That notwithstanding, past the first few chapters, we have generally tried to group content meaningfully for convenient reference during learning.
In addition, since this is a beginner's textbook, only the more common usages are explained.

We have also aimed to make this a self-instruction textbook so that a diligent student should, if [#allAh]{.trn2} wills, be able to study it without an instructor. The target learner is someone who has not been exposed to grammatical terminology like _inflection_, _case_, _mood_, etc. While terminology is necessary for a rigorous non-immersive learning of language, we have tried to steer away from Latin-based terms like _accusative_ and _jussive_. Such terms, when first encountered by an uninitiated learner, may deter from proceeding further. (Learning a language can be hard enough without getting the feeling that your grammar book is accusing you of something!) So we have in some places translated the meaning of Arabic grammar terms to English. In other places, we have used established English grammar terms where the terms are basic enough. We have even, in places, invented terms where we deemed appropriate. The drawback to this non-standard approach, however, is that the student may not be able to immediately relate the terminology he has learned in this book to established terminology in other grammar textbooks. To remedy this to some extent, we provide a glossary in the appendix which maps the grammatical terminology used in this book to other, established, Latin-based and Arabic-based counterparts.

It may also be appropriate to inform the reader that we chose to present a simplified version of Arabic grammar. As such, the grammar presented here may not be entirely consistent with the comprehensive and harmonious framework developed by the Arab grammarians. We chose this approach because we felt that exposing the beginner to complex grammatical details at this stage would be more of a hindrance than a help in learning the language.

This book is a currently work in progress and is produced with the R bookdown package. The code and text for volume\ 1 are open-sourced and developed at 
[github.com/adamiturabi/arabic-tutorial-book](https://github.com/adamiturabi/arabic-tutorial-book).
The typeset output of volume\ 1 is published at 
[adamiturabi.github.io/arabic-tutorial-book/](https://adamiturabi.github.io/arabic-tutorial-book/).

[the authors]{.smallcaps}  

<!--
[a.z.s.]{.smallcaps}  
[lahore, dec.]{.smallcaps} 2021[ce]{.smallcaps}  
[لاهور، جمادى الأولى ١٤٤٣هـ‍]{.ar}

Test:

This is the _preface_. Test trans [Parf makAn ealayhi pcv machad]{.trn} [◌َ  ◌ً  ◌ِ  ◌ٍ  ◌ُ  ◌ٌ  ◌ْ  ◌ّ  ◌ٰ  ◌ٔ  ◌ٕ  ]{.ar}

[شَيْـَٔيْنِ بَرِيـِٔينَ شَيْـًٔا خَطِيـَٔة]{.ar}

[شَيْءَيْنِ بَرِيءِينَ شَيْءًا خَطِيءَة]{.ar}


this is in [Small Caps]{.smallcaps}.


```{r setup ,echo=FALSE}
options(tinytex.engine = "lualatex")
```
-->

# Romanization scheme {-}

\markboth {\textsc{\MakeLowercase{Romanization scheme}}}{\textsc{\MakeLowercase{Romanization scheme}}}

The following romanization scheme is used in this book for the transcription and transliteration of Arabic sounds and letters into the Latin script.

[consonants]{.smallcaps}

<!--
| [ء]{.ar} | [ب]{.ar} | [ت]{.ar} | [ث]{.ar} | [ج]{.ar} | [ح]{.ar} | [خ]{.ar} | [د]{.ar} | [ذ]{.ar} | [ر]{.ar} | [ز]{.ar} | [س]{.ar} | [ش]{.ar} | [ص]{.ar} |
|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-| 
| [E]{.trn2} | [b]{.trn2} | [t]{.trn2} | [v]{.trn2} | [j]{.trn2} | [H]{.trn2} | [x]{.trn2} | [d]{.trn2} | [p]{.trn2} | [r]{.trn2} | [z]{.trn2} | [s]{.trn2} | [c]{.trn2} | [S]{.trn2} |

| [ض]{.ar} | [ط]{.ar} | [ظ]{.ar} | [ع]{.ar} | [غ]{.ar} | [ف]{.ar} | [ق]{.ar} | [ك]{.ar} | [ل]{.ar} | [م]{.ar} | [ن]{.ar} | [ه]{.ar} | [و]{.ar} | [ي]{.ar} | 
|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|:-|
| [D]{.trn2} | [T]{.trn2} | [P]{.trn2} | [e]{.trn2} | [g]{.trn2} | [f]{.trn2} | [q]{.trn2} | [k]{.trn2} | [l]{.trn2} | [m]{.trn2} | [n]{.trn2} | [h]{.trn2} | [w]{.trn2} | [y]{.trn2} | 
-->

|   |   |   |   |   |   |   |   |   |   |   |   |   |   |
|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|
|  [ص]{.ar}  |  [ش]{.ar}  |  [س]{.ar}  |  [ز]{.ar}  |  [ر]{.ar}  |  [ذ]{.ar}  |  [د]{.ar}  |  [خ]{.ar}  |  [ح]{.ar}  |  [ج]{.ar}  |  [ث]{.ar}  |  [ت]{.ar}  |  [ب]{.ar}  |  [ء]{.ar}  |
| [S]{.trn2} | [c]{.trn2} | [s]{.trn2} | [z]{.trn2} | [r]{.trn2} | [p]{.trn2} | [d]{.trn2} | [x]{.trn2} | [H]{.trn2} | [j]{.trn2} | [v]{.trn2} | [t]{.trn2} | [b]{.trn2} | [E]{.trn2} |

|   |   |   |   |   |   |   |   |   |   |   |   |   |   |
|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|
|  [ى]{.ar}  |  [و]{.ar}  |  [ه]{.ar}  |  [ن]{.ar}  |  [م]{.ar}  |  [ل]{.ar}  |  [ك]{.ar}  |  [ق]{.ar}  |  [ف]{.ar}  |  [غ]{.ar}  |  [ع]{.ar}  |  [ظ]{.ar}  |  [ط]{.ar}  |  [ض]{.ar}  |
| [y]{.trn2} | [w]{.trn2} | [h]{.trn2} | [n]{.trn2} | [m]{.trn2} | [l]{.trn2} | [k]{.trn2} | [q]{.trn2} | [f]{.trn2} | [g]{.trn2} | [e]{.trn2} | [P]{.trn2} | [T]{.trn2} | [D]{.trn2}  |

[vowels]{.smallcaps}

|   |   |   |   |   |   |   |   |
|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|
| [◌َ]{.ar}   | [◌ُ]{.ar}   | [◌ِ]{.ar}   | [◌َا]{.ar}  | [◌ُو]{.ar}  | [◌ِي]{.ar}  | [◌َوْ]{.ar}   | [◌َيْ]{.ar}   |
| [a]{.trn2} | [u]{.trn2} | [i]{.trn2} | [A]{.trn2} | [U]{.trn2} | [I]{.trn2} | [aw]{.trn2} | [ay]{.trn2} |


When transcribing example text,
the _italic_ Latin script is used,
no letters are capitalized,
a non-sentence-initial connecting hamzah is transliterated with a hyphen (-),
and the [lAm]{.trn2} of the definite article is not transliterated for sun letters. 
[ة]{.ar} is transcribed as [h]{.trn} at the end of an utterance, and as [t]{.trn} otherwise. 
For example:

<!--[ٱَلشَّجَرَةُ وَٱلْوَرَقَة.]{.ar}  
[Eaccjaratu wa-lwaraqah.]{.trn}.-->
<!--[بِسمِ ٱللَّـٰهِ ٱلرَّحْمَـٰنِ ٱلرَّحِيمِ.]{.ar}  
[bismi -llAhi -rraHmAni -rraHIm.]{.trn}-->
[ٱِلْتَقَطَتْ أُمُّ طَلْحَةَ وَٱلْحَسَنِ وَرَقَةَ ٱلشَّجَرَةِ.]{.ar}  
[EiltaqaTat Eummu TalHata wa-lHasani waraqata -ccajarah.]{.trn}

<!--FIXME: add transcription of assimilation-->
<!--The transcription is typically accurate, except for assimilation that only affects diacritics. This is generally marked neither in the Arabic text nor in the transcription.
For example:

[ٱِعْتَقَدْتُ أَنْ لَمْ يَأْتِ.]{.ar}  
-->


When using Arabic words in the English text,
the regular (non-italic) Latin script is used,
words are transcribed in pausal pronunciation,
word-final [◌ِيّ]{.ar} [-iyy]{.trn} is transcribed as "[I]{.trn2}",
the first letter of a proper noun is capitalized,
word-initial hamzah is not transliterated,
the definite article is always transliterated as "al-" with a hyphen following it,
as shown.
[ة]{.ar} is transcribed as "t" in the middle of an annexation, and as "h" otherwise. 
Accurate romanizations are preferred unless an established and distinctly separate usage is current in English.
Examples:

<!--"[#allAh]{.trn2} revealed [#sUrat al-#fatiHah]{.trn2} in the [#qurEAn]{.trn2}."-->
"[#ealI]{.trn2} and [#usAmah]{.trn2} revised [#sUrat al-#eankabUt]{.trn2} from the [#qurEAn]{.trn2} in the mosque in the month of [#rabIe al-#vAnI]{.trn2}."

"The caliphate moved from [al-#madInah]{.trn2} to Damascus."

