# The Arabic script

Arabic is written right to left. The alphabet traditionally has 28 letters. Some of the sounds are similar to sounds in English but others are very different. Here we will attempt to describe the sounds but we recommend that you learn the correct pronunciation from an experienced Arabic or Quran teacher.

```{r script-alphabet-table, echo = FALSE}
mytable <- read.csv("script_alphabet_table.csv")
knitr::kable(mytable, caption = 'The Arabic alphabet')
```

thaa, dhaal, and Dhaa. Try to pronoun th as in think but bring your tongue farther out between the top and bottom front teeth and have it touch the teeth lighter. Have your lips and cheek in a wide grin.

Then the th sound will be closer to s and this is thaa.

Now try to bring it closer to z but with the tongue in the same place.

Now instead of a wide grin, make your lips small and round, and try to make the th sound again but closer to a z. This is the Dhaa.

The letters ص ض ط ظ can br approximated with mouth rounded. س د ت ذ

## Writing Arabic

### The different forms of the letters

Arabic is written right-to-left, unlike English and most other languages which are written left-to-right. When writing the letters in a word are generally joined to each other, except for six out of the 28 letters. These six letters are [ا، د، ذ، ر، ز، و]{.ar}. They join to the letter before them, like the rest of the letters, but they don't join to the letter following them.

When joining the letters, each letter is modified in order to join to the preceding and following letter. Therefore, there can be up to four shapes for each letter: when the letter is:

1. by itself (isolated).
2. in the beginning of a word,
3. in the middle of a word,
4. in the end of a word

So does that mean that we have to remember $4 \times 28 = 112$ different shapes? Not really. Remember that six of the letters don't join to the following letter. So they only have isolated and end forms. And also, each letter maintains its basic shape and is usually only slightly modified in a common technique to join to the previous and following letter. So if we remember this common technique of modifying the letter shape to join it it will be easy to join the letters.

To explain the common technique of modifying the letters we will take [ب]{.ar} as an example and start with the isolated form:

Isolated form: [ب]{.ar}

To modify this into the end form, we simply join a horizontal line to the right of the letter:

End form: [ـب]{.ar}. 

To get the middle form, we take the end form [ـب]{.ar} and cut off its tail which is at its left, and replace it with a horizontal line. We also move the dot slightly to get:

Middle form: [ـبـ]{.ar}

And finally, to get the beginning form, we take the middle form [ـبـ]{.ar} and remove the horizontal line at the right:

Beginning form: [بـ]{.ar}

Now most of the letters follow this common technique but a few of them are modified a little further in each form. These, more complicated, letters are [ع، غ، ك، ه، ي]{.ar} and you can study them and the rest of the letters in the table below:

Here is a table of all the letters all four positions:


```{r script-alphabet-connceted-forms, echo = FALSE}
mytable <- read.csv("script_alphabet_connected_forms.csv")
knitr::kable(mytable, caption = 'Arabic letters in different positions')
```

You can see that each letter maintains a basic shape and is modified for each of the four positions. 



### Joining the different forms to make a word

Notice that when we modified the isolated form to get to the beginning, middle, and end forms, we added a horizontal line to each or both sides. It is this horizontal line which joines to the horizontal line of the neighboring letter.

As an example, we would like to join the following letters (starting from the right): [م، ع، ش، ر]{.ar} into one word. The first letter is [م]{.ar} so we modify it to its beginning form [مـ]{.ar}. The next two letters are converted to their middle forms [ـعـ، ـشـ]{.ar}. And the last letter [ر]{.ar} is converted to its end form [ـر]{.ar}. Then we join the horizontal lines together and get [مــعــشــر]{.ar}. Usually, when we join letters like this we shorten the horizontal lines so you will generally see the word like this [معشر]{.ar}. (Don't worry right now about how to pronounce this word. Soon, if Allah wills, we will learn how to pronoune and read Arabic.)

So if Arabic letters are always joined, why do we need isolated forms. Shouldn't the beginning, middle and end forms be sufficient? Here's where the no-end-join letters [ا، د، ذ، ر، ز، و]{.ar} come in. As we mentioned these letters don't have beginning and middle forms. So if one of them comes in the beginning of a word, the isolated form will be used. And if it comes in the middle of a word, the end form will be used. Unless there is another no-end-join letter before it, in which case the isolated form will be used.

Even a no-end-join letter can be in the isolated form if it is the last letter in a word and there is a no-end-join letter before it.

Let's take a look at some examples. Join the following groups of letters:

Disjoint | Joined
:------------|:---------
[ذ،ل،ك]{.ar} | [ذلك]{.ar}
[ا،ح،م،د]{.ar} | [احمد]{.ar}
[ر،س،و،ل]{.ar} | [رسول]{.ar}
[و،ز،ي،ر]{.ar} | [وزير]{.ar}
[و،ا،د،ي]{.ar} | [وادي]{.ar}

Notice that in the last example, all the letters were in the isolated form.

**Replace 'join' with 'join' in this chapter**

### Taking care to tell similar letters apart.

Some letters are very similar to each other and only differ in their dots or other slight differences. You should take care to distinguish between these letters. We will describe their similarities and differences here.

The letters [ب]{.ar}, [ت]{.ar}, and [ث]{.ar} differ only in their dots and are otherwise identical in all positions. [ن]{.ar} and [ي]{.ar} are similar in initial and middle positions to [ت]{.ar}, and [ث]{.ar}]{.ar} but differ from them and from each other in isolated and final positions. Compare all five in he table below:

Isolated | End | Middle | Beginnning
:--------|:----|:-------|:----------
[ب]{.ar}| [ـب]{.ar}| [ـبـ]{.ar}| [بـ]{.ar}
[ت]{.ar}| [ـت]{.ar}| [ـتـ]{.ar}| [تـ]{.ar}
[ث]{.ar}| [ـث]{.ar}| [ـثـ]{.ar}| [ثـ]{.ar}
[ن]{.ar}| [ـن]{.ar}| [ـنـ]{.ar}| [نـ]{.ar}
[ي]{.ar}| [ـي]{.ar}| [ـيـ]{.ar}| [يـ]{.ar}

These groups of letters  differ too, only in their dots:

+ [ج]{.ar}, [ح]{.ar}, and [خ]{.ar}
+ [د]{.ar} and [ذ]{.ar}
+ [ر]{.ar} and [ز]{.ar}
+ [س]{.ar} and [ش]{.ar}
+ [ص]{.ar} and [ض]{.ar}
+ [ط]{.ar} and [ظ]{.ar}
+ [ع]{.ar} and [غ]{.ar}

The letters ف and ق are similar in the initial and middle positions except for the dots. But in the isolated and final positions ق goes lower than ف.

Isolated | End | Middle | Beginnning
:--------|:----|:-------|:----------
[ف]{.ar}| [ـف]{.ar}| [ـفـ]{.ar}| [فـ]{.ar}
[ق]{.ar}| [ـق]{.ar}| [ـقـ]{.ar}| [قـ]{.ar}

Be careful also not to confuse [غ]{.ar} and [ف]{.ar} in their middle forms. The loop for [ف]{.ar} is round where it is triangular and flat-topped for [غ]{.ar} (as it is for [ع]{.ar}). Compare their middle forms in the table below:

Isolated | Middle
:--------|:------
[غ]{.ar}| [ـغـ]{.ar}
[ف]{.ar}| [ـفـ]{.ar}

The letters alif ا and laam ل can be confused for each other. Their forms are shown here again for easy comparison:


Isolated
ل
ا
Initial
لـ


End
ـل
ـا
Middle
ـلـ


Thus each letter joins to the letters before and after it through this horizontal line.

When writing, words are written from right to left. Individual words are also separated by spaces just like in English. Letters in a word are joined to each other except if a word contains a no-end-join letter. A no-end-join letter is attached to the letter before it but is not joined to the letter after it.

The no-end-join letters are ا د ذ ر ز و

Because of this, each letter can have up to four forms:
Initial form: This form is used when the letter is in the beginning of a word when it is joined to the following letter. It is also used in the middle of the word if it follows a no-end-join letter. No-end-join letters don’t have initial forms; their isolated forms are used instead.
End form: This is used when the letter is at the end of a word and if it can be joined to the previous letter. All letters have end forms.
Middle form: is used when the letter is in the middle of a word when it joins to the previous letter and the next letter. No-end-join letters don’t have middle forms. Isolated or end forms are used instead depending on whether the letter before the no-end-join letter is itself a no-end-join letter.
alone form: This is the form when the letter is written by itself. It is also used if it is the last letter of a word following a no-end-join letter. If a no-end-join letter occurs in the beginning of a word they will be written in the isolated form instead since they won’t join to the following letter. No-end-join letters will also be in the isolated form in the middle of a word if they are preceded by other no-end-join letters. Examples with isolated forms in bold: 
ح
جار
دار
قرون

Name
Daad
Saad
sheen
seen
zaa
raa
daal
daal
khaa
Haa
jeem
thaa
taa
baa
alif
Isolated
ض
ص
ش
س
ز
ر
د
ذ
خ
ح
ج
ث
ت
ب
ا
Initial
ضـ
صـ
شـ
سـ








خـ
حـ
جـ
ثـ
تـ
بـ


End
ـض
ـص
ـش
ـس
ـز
ـر
ـذ
ـد
ـخ
ـح
ـج
ـث
ـت
ـب
ـا
Middle
ـضـ
ـصـ
ـشـ
ـسـ








ـخـ
ـحـ
ـجـ
ـثـ
ـتـ
ـبـ




 
Isolated
ي
و
ه
ن
م
ل
ك
ق
ف
غ
ع
ظ
ط
Initial
يـ


هـ
نـ
مـ
لـ
كـ
قـ
قـ
غـ
عـ
ظـ
طـ
End
ـي
ـو
ـه
ـن
ـم
ـل
ـك
ـق
ـف
ـغ
ـع
ـظ
ـط
Middle
ـيـ


ـهـ
ـنـ
ـمـ
ـلـ
ـكـ
ـقـ
ـفـ
ـغـ
ـعـ
ـظـ
ـظـ

Take care to distinguish between these letters:

The letters ب ت ث differ only in the dots and are otherwise identical in all positions. ن and ي are similar in initial and middle positions to ب ت ث but differ in isolated and final positions.

The pairs of letters ج ح خ ، د ذ ، ر ز ، س ش ، ص ض ، ط ظ ، ع غ differ too, only in dots..

The letters ف and ق are similar in the initial and middle positions except for the dots. But in the isolated and final positions ق goes lower than ف.

Be careful also not to confuse غ and ف in their middle forms. The loop for ف is round where it is triangular and flat-topped for غ. 

The letters alif ا and laam ل can be confused for each other. Their forms are shown here again for easy comparison:


Isolated
ل
ا
Initial
لـ


End
ـل
ـا
Middle
ـلـ



Looped taa ة
Looped taa ة is similar to regular taa ت in most respects but it is written differently as a ه with two dots above it. It only occurs at the end of a word so it has only an end form and an isolated form. The isolated form is used if the letter before it is a no-end-join letter. Looped taa ة is pronounced exactly as a ت except the end of a sentence when it is pronounced as a ه.
Special forms of letter occurring together
When alif follows a laam in the same word instead of writing لـا  they are written as لا.
Vowels and pronunciation marks.
Vowels
Arabic has six vowels
Three short vowels, not part of the alphabet, but shown with pronunciation marks.
Short-a, transcribed “a” like in English  “mud”, shown with a small diagonal line above the letter like مَـ ma.
“I” like in English “bit”, shown with a small diagonal line under the letter like بِـ bi
“U” like in English “put”, shown with a tiny waw above the letter like فُـ fu.
three long vowels that we’ve already seen which are part of the alphabet:
Long-a, transcribed “aa”, like in English “man”, written with a short-a followed by an unmarked alif eg  ذَا dhaa
“Oo” like in English “moon” written with a short-u followed by an unmarked waaw ذُو
“Ee” like in English “been” written with a short-i followed by an unmarked yaa ذِي dhee 
It also has two short semi-vowels:
“Aw” like in English “goal” shown with a short-a before a waw like لَوْ law. (Note the small circle above the ي called a sukoon which we’ll talk about soon.)
“Ay” like in English “bait” shown with a short-a before a waaw like كَيْ kay.
And it also has two long semi-vowels:
“Aaw” like English “now” shown with an alif before a waw with sukoon. واوْ.
“Aay” like English “bye” shown with an alif before a yaa with sukoon شـايْ.
These last two long semi-vowels are rare and only occur at the end of a sentence.
Sukoon
In Arabic each letter is either pronounced with one of the above vowels or it is pronounced without a vowel, in which case we will put a small circle (use dotted circle instead of tatweel) ـْ, called a sukoon above it. In the word
نِعْمَ ni3ma
The letter ن is pronounced with a short-i, م is pronounced with a short-a, but ع is pronounced without a vowel so it is written ـعْـ.

Similarly, [more examples]
Shaddah
Khfjkhg
Small alif
Sometimes the long-a vowel is written as a small alif above the letter instead of a regular alif. For example هذا haadhaa. This is done only for a few commonly used words that we will learn later if Allah wills. One of these words is الله Allaah, the Arabic name for God.
Long-a vowel written as yaa
The vowel long-a is usually written with an alif. But sometimes, it is written with a yaa instead (but still pronounced as a long-a). This is only occurs for certain words and always at the end of a word. When this occurs we will write the yaa without its dots thus ى and often also put a small alif above it thus .
Tanween
Take a look at these words:
سَعْدٌ sa3dun، ضَرْبًا Darban، دُعَاءً du3aa'an, سَعَةً sa3atan
We have shown three new pronunciation marks ـٌ ـً ـٍ These are pronounced like their short vowels but with a “n” sound at the end. The “n” sound is shown, not by a ن, but by the doubled short vowel pronunciation marks. This occurs for grammatical reasons that we will learn later if Allah wills.

We will call these pronunciation marks a-tanween (a-n), i-tanween, and u-tanween. Collectively they are called (n-marks) and only occur at the end of a word.

Note that ضربًا Darban has an extra alif at the end, but this alif is not pronounced. It is added only as a spelling rule which is: a-tanween is always followed by a silent alif, except if the a-tanween is on a looped taa ة. There are also additional rules when writing a-tanween with hamza which we will explain later.

Note also that it is perfectly normal to have a word that ends with a ن that has a doubled vowel mark like غَبَنٌ ghabanun.
Hamza
Hamza ء is not traditionally considered part of the alphabet but it is very much a letter. Its called a “glottal stop” in English. You can hear it if you say “a apple” quickly. When you stop your breath momentarily between “a” and “apple” that is the sound that hamza makes. In fact it also occurs in the beginning of “a apple” when you start to say “a”. In transcribing we will represent it with ‘. Note that this is different than out representation for ع which is `.

In Arabic writing hamza can either be written by itself ء or it can be written “sitting” on alif, waw or ي thus أ ؤ ئ . It can also be written under an alif إ. In this case the ا و or ي are called the seat of the hamza. Note that when sitting on ي the dots of the yaa will not be written so depending on position, it can appear ئ ئـ ـئ ـئـ.

// maybe: If hamza follow by alif ءا comes in the beginning of a word, it is written as آ instead. 
// In hamza spelling rules also explain a-tanween

But whether it appears by itself ء or with a seat the pronunciation is the same. The choice of whether to write it by itself or with a “seat” is only a spelling rule. However this rule is quite complicated so we will explain it in a later chapter. For now we will only give the rule for initial hamza:

When hamza occurs at the beginning of a word it will be written as:
As أَ with short-a, e.g., أَحْمَدُ ‘aHmadu
As أُ with short-u, eg
As إِ with short-i
As آ with long-a, eg آنَسْتُ ‘aanastu. Note that instead of a hamza (dotted circle with hamza) we wrote a madda symbol (dotted circle with madda).
As أُو with long-a, eg أُوجِرَ ‘oojira
As إِي with a long-i eg إِي

Besides this we will limit the words used with hamza until we explain the rules fully so hopefully it will not be too confusing.
Connecting hamza
Some words in arabic begin with a sukoon. When this occurs a connecting hamza (written as a tiny صـ on an alif) is put before it. If this word comes in the beginning of the sentence the connecting alif is pronounced as a hamza. Otherwise this connecting hamza is not pronounced and the word is connected to the previous word in pronunciation.

In this tutorial we will transcribe the connecting hamza with a hyphen - when it is in the middle of a sentence.

Examples
‘alwaladu sa3eedun ٱلْوَلَدُ سَعِيدٌ
Daraba -lwaladu ‘aHmada NOT Daraba ‘alwaladu aHmada

When one word ends in a long vowel and the next word begins with a connecting hamza, the long vowel becomes a short vowel in pronunciation (but in writing the long vowel’s letter is retained.) For example

‘Ilaa masjidin BUT Dhahaba ‘ila -lmasjidi NOT dhahaba ‘ilaa -lmasjidi NOR ‘ilaa ‘almasjidi
‘Illaa zaydan BUT ‘Illa -lmuSalleena NOT `illaa -lmuSalleena NOR ‘illaa ‘almuSalleena.
Fee baytin BUT Fi -lbayt NOT fee -lbayt NOR fee ‘albayt
Dhoo sa3atin BUT Dhu -lqarnayni NOT dhoo -lqarnayni NOR dhoo ‘alqarnayni 

Sometimes a connecting hamza can occurs after an n-mark. In these cases will will pronounce the n-mark and follow it with a short-i:

Zaydani -lkabeeru NOT zaydan ‘alkabeera
Pausing at the end of a sentence.
Difference between Quranic script and Standard Arabic script.
In printed Qurans the script will differ somewhat from the one we’ve shown here, especially in vowels. The reasons for this are beyond the scope of this tutorial. But here we’ll just give a few examples and note that these differences are only found in printed Qurans. For standard written Arabic, the script explained in this chapter is used.

إبرهم ابراهم 'ibraheem
ربو ribaa
Madda ءانست aanastu حآفين Haaffeena


The script and spelling rules presented so far show how Standard Arabic is written in regular books and other written material.
The Quran is also written in Arabic but due to reasons that beyond the scope of this tuorial, it’s script often differs slightly from Standard Arabic’s. For example
Madda
Dagger alif on waw

