# The Arabic script

## The Arabic alphabet

The Arabic alphabet traditionally has 28 letters. There is also one additional letter called [hamzah]{.trn}, that was originally only sounded and not written. Together with [hamzah]{.trn} there are 29 letters.

The following table shows all the letters in the Arabic alphabet.

```{r script-alphabet-table, echo = FALSE}
mytable <- read.csv("script_alphabet_table.csv")
knitr::kable(mytable, caption = 'The Arabic alphabet')
```

### Pronunciation notes

Some of the sounds are similar to sounds in English but others are very different. Here we will attempt to describe the sounds but we recommend that you learn the correct pronunciation from an experienced Arabic or Quran teacher.

thaa, dhaal, and Dhaa. Try to pronoun th as in think but bring your tongue farther out between the top and bottom front teeth and have it touch the teeth lighter. Have your lips and cheek in a wide grin.

Then the th sound will be closer to s and this is thaa.

Now try to bring it closer to z but with the tongue in the same place.

Now instead of a wide grin, make your lips small and round, and try to make the th sound again but closer to a z. This is the Dhaa.

The letters ص ض ط ظ can br approximated with mouth rounded. س د ت ذ

## Joining Arabic letters

### The different forms of the letters

Arabic is written right-to-left, unlike English and most other languages which are written left-to-right. When writing the letters in a word are generally joined to each other, except for six out of the 28 letters. These six letters are [ا، د، ذ، ر، ز، و]{.ar}. They join to the letter before them, like the rest of the letters, but they don't join to the letter following them.

When joining the letters, each letter is modified in order to join to the preceding and following letter. Therefore, there can be up to four shapes for each letter: when the letter is:

1. by itself (isolated).
2. in the beginning of a word,
3. in the middle of a word,
4. in the end of a word

So does that mean that we have to remember $4 \times 28 = 112$ different shapes? Not really. Remember that six of the letters don't join to the following letter. So they only have isolated and end forms. And also, each letter maintains its basic shape and is usually only slightly modified in a common technique to join to the previous and following letter. So if we remember this common technique of modifying the letter shape to join it it will be easy to join the letters.

To explain the common technique of modifying the letters we will take [ب]{.ar} as an example and start with the isolated form:

Isolated form: [ب]{.ar}

To modify this into the end form, we simply join a horizontal line to the right of the letter:

End form: [ـب]{.ar}. 

To get the middle form, we take the end form [ـب]{.ar} and cut off its tail which is at its left, and replace it with a horizontal line. We also move the dot slightly to get:

Middle form: [ـبـ]{.ar}

And finally, to get the beginning form, we take the middle form [ـبـ]{.ar} and remove the horizontal line at the right:

Beginning form: [بـ]{.ar}

Now most of the letters follow this common technique but a few of them are modified a little further in each form. These, more complicated, letters are [ع، غ، ك، ه، ي]{.ar} and you can study them and the rest of the letters in the table below:

Here is a table of all the letters all four positions:


```{r script-alphabet-connceted-forms, echo = FALSE}
mytable <- read.csv("script_alphabet_connected_forms.csv")
knitr::kable(mytable, caption = 'Arabic letters in different positions')
```

You can see that each letter maintains a basic shape and is modified for each of the four positions. 

### Joining the different forms to make a word

Notice that when we modified the isolated form to get to the beginning, middle, and end forms, we added a horizontal line to each or both sides. It is this horizontal line which joines to the horizontal line of the neighboring letter.

As an example, we would like to join the following letters (starting from the right): [م-ع-ش-ر]{.ar} into one word. The first letter is [م]{.ar} so we modify it to its beginning form [مـ]{.ar}. The next two letters are converted to their middle forms [ـعـ، ـشـ]{.ar}. And the last letter [ر]{.ar} is converted to its end form [ـر]{.ar}. Then we join the horizontal lines together and get [مــعــشــر]{.ar}. Usually, when we join letters like this we shorten the horizontal lines so you will generally see the word like this [معشر]{.ar}.

So if Arabic letters are always joined, why do we need isolated forms. Shouldn't the beginning, middle and end forms be sufficient? Here's where the no-end-join letters [ا، د، ذ، ر، ز، و]{.ar} come in. As we mentioned these letters don't have beginning and middle forms. So if one of them comes in the beginning of a word, the isolated form will be used. And if it comes in the middle of a word, the end form will be used. Unless there is another no-end-join letter before it, in which case the isolated form will be used.

The letter after a no-end-join letter will be in its beginning form even though it is in the middle of a word. But if it is the last letter in the word then it will take its isolated form.

Let's take a look at some examples. Join the following groups of letters:

Disjoint | Joined
:------------|:---------
[ذ-ل-ك]{.ar} | [ذلك]{.ar}
[ا-ح-م-د]{.ar} | [احمد]{.ar}
[ر-س-و-ل]{.ar} | [رسول]{.ar}
[و-ز-ي-ر]{.ar} | [وزير]{.ar}
[ر-ا-ز-ق]{.ar} | [رازق]{.ar}

Notice that in the last example, all the letters were in the isolated form.

### Exercises

## Taking care to tell similar letters apart.

Some letters are very similar to each other and only differ in their dots or other slight differences. You should take care to distinguish between these letters. We will describe their similarities and differences here.

The letters [ب]{.ar}, [ت]{.ar}, and [ث]{.ar} differ only in their dots and are otherwise identical in all positions. [ن]{.ar} and [ي]{.ar} are similar in initial and middle positions to [ب]{.ar}, [ت]{.ar}, and [ث]{.ar} but differ from them and from each other in isolated and final positions. Compare all five in the table below:

Isolated | End | Middle | Beginnning
:--------|:----|:-------|:----------
[ب]{.ar}| [ـب]{.ar}| [ـبـ]{.ar}| [بـ]{.ar}
[ت]{.ar}| [ـت]{.ar}| [ـتـ]{.ar}| [تـ]{.ar}
[ث]{.ar}| [ـث]{.ar}| [ـثـ]{.ar}| [ثـ]{.ar}
[ن]{.ar}| [ـن]{.ar}| [ـنـ]{.ar}| [نـ]{.ar}
[ي]{.ar}| [ـي]{.ar}| [ـيـ]{.ar}| [يـ]{.ar}

These groups of letters  differ too, only in their dots:

+ [ج]{.ar}, [ح]{.ar}, and [خ]{.ar}
+ [د]{.ar} and [ذ]{.ar}
+ [ر]{.ar} and [ز]{.ar}
+ [س]{.ar} and [ش]{.ar}
+ [ص]{.ar} and [ض]{.ar}
+ [ط]{.ar} and [ظ]{.ar}
+ [ع]{.ar} and [غ]{.ar}

The letters [ف]{.ar} and [ق]{.ar} are similar in the initial and middle positions except for the dots. But in the isolated and final positions, the tail of [ق]{.ar} goes lower than that of [ف]{.ar}.

Isolated | End | Middle | Beginnning
:--------|:----|:-------|:----------
[ف]{.ar}| [ـف]{.ar}| [ـفـ]{.ar}| [فـ]{.ar}
[ق]{.ar}| [ـق]{.ar}| [ـقـ]{.ar}| [قـ]{.ar}

Be careful also not to confuse [غ]{.ar} and [ف]{.ar} in their middle forms. The loop for [ف]{.ar} is round where it is triangular and flat-topped for [غ]{.ar} (as it is for [ع]{.ar}). Compare their middle forms in the table below:

Isolated | Middle
:--------|:------
[غ]{.ar}| [ـغـ]{.ar}
[ف]{.ar}| [ـفـ]{.ar}

The letters [Ealif]{.trn} [ا]{.ar} and [lAm]{.trn} [ل]{.ar} can be confused for each other. Their forms are shown here again for easy comparison:

Isolated | End | Middle | Beginnning
:--------|:----|:-------|:----------
[ا]{.ar}| [ـا]{.ar}| none | none
[ل]{.ar}| [ـل]{.ar}| [ـلـ]{.ar}| [لـ]{.ar}

## Vowels and pronunciation marks.

In Arabic each letter in a word will have a vowel after it unless if there is a vowel right before it in which case it may or may not have avowel after it.

### Short Vowels

Arabic has six vowels. There are three short vowels which don't have letters in the alphabet. Instead they are shown with pronunciation marks:

1.  [a]{.trn} like in English “mud”, written with an [a]{.trn}-mark which is a small diagonal line above the letter like [مَـ]{.ar} [ma]{.trn}.
1.  [i]{.trn} like in English “bit”, written with an [i]{.trn}-mark which is a small diagonal line under the letter like [بِـ]{.ar} [bi]{.trn}.
1.  [u]{.trn} like in English “put”, written with an [u]{.trn}-mark which is a tiny waw above the letter like [فُـ]{.ar} [fu]{.trn}.

Examples of words with short vowels:

+ [فَتَحَ]{.ar} [fataHa]{.trn}
+ [عَمِلَ]{.ar} [eamila]{.trn}
+ [قُتِلَ]{.ar} [qutila]{.trn}

### Long Vowels

There are also three long vowels which are part of the alphabet:

1.  [A]{.trn} generally written with an unmarked [ا]{.ar} [Ealif]{.trn} and with the preceding letter having an [a]{.trn}-mark. Example [مَا]{.ar} [mA]{.trn}. This vowel is mostly pronounced like the vowel in English “man”. If however, it comes after these letters [خ،ر،ص،ض،ط،ظ،غ،ق]{.ar} it is pronounced like English "awe".
1.  [I]{.trn} like in English “been” written with an unmarked [ي]{.ar} [yAE]{.trn} with the preceding letter having an [i]{.trn}-mark. Example [فِي]{.ar} [fI]{.trn}.
1.  [U]{.trn} like in English “moon” written with an unmarked [و]{.ar} [wAw]{.trn} with the preceding letter having an [u]{.trn}-mark. Example [ذُو]{.ar} [pU]{.trn}.

Examples of words with long and short vowels:

+ [هَارُونُ]{.ar} [hArUnu]{.trn}
+ [كَذَا]{.ar} [kapA]{.trn}
+ [سَرَادِيبَ]{.ar} [sarAdIba]{.trn}

#### [A]{.trn} vowel written with a small [Ealif]{.trn}

Sometimes the [A]{.trn} vowel is written as a small small [Ealif]{.trn} instead of a regular [Ealif]{.trn}. This is done only for a few commonly used words. Here are some examples:

+ [هَـٰذَا]{.ar} [hApA]{.trn}
+ [ذَ﻿ٰلِكَ]{.ar} [pAlika]{.trn}

#### [A]{.trn} vowel written with a [yAE]{.trn}

The [A]{.trn} vowel is generally written with an [Ealif]{.trn} as we have seen above. But for some words, instead of an [Ealif]{.trn}, it is written with a [yAE]{.trn}. This only occurs for certain words and even then only at the end of the word. When this happens, we will write the [yAE]{.trn} without its dots and write a small [Ealif]{.trn} above it, like this [ىٰ]{.ar}. Here are some examples:

+ [عَلَىٰ]{.ar} [ealA]{.trn}
+ [رَمَىٰ]{.ar} [ramA]{.trn}

### Round mark

As we have seen above if an Arabic letter has a vowel after it it will take one of the three pronunciation marks [ـَ ـِِ ـُ]{.ar}. If, however, there is no vowel after the letter we will put a round mark on it [ـْ]{.ar}. This mark can only occur if there is a short vowel before the letter. Examples:

+ [كَمْ]{.ar} [kam]{.trn}
+ [مُنْذُ]{.ar} [munpu]{.trn}
+ [مِنْهُمْ]{.ar} [minhum]{.trn}
+ [مِنْهَا]{.ar} [minhA]{.trn}

### Semi-vowels

Arabic has two short semi-vowels:

1. [aw]{.trn} like in English “goal”. This is written with a [wAw]{.trn} with a round-mark on it and a [A]{.trn} vowel before it. Example [لَوْ]{.ar} [law]{.trn}. 
2. [ay]{.trn} like in English “bait”. This is written with a [yAE]{.trn} with a round-mark on it and a [A]{.trn} vowel before it. Example [كَيْ]{.ar} [kay]{.trn}. 
Examples with short semi-vowels:

+ [وَيْحَكَ]{.ar} [wayHaka]{.trn}
+ [غَيْرُهُ]{.ar} [gayruhu]{.trn}
+ [قَوْلُهُ]{.ar} [qawluhu]{.trn}

It also has two long semi-vowels:

1. [Aw]{.trn} like in English “cow”. This is written with a [wAw]{.trn} with a round-mark on it and a [A]{.trn} vowel before it. Example [وَاوْ]{.ar} [wAw]{.trn}. 
2. [Ay]{.trn} like in English “bye”. This is written with a [yAE]{.trn} with a round-mark on it and a [A]{.trn} vowel before it. Example [شَايْ]{.ar} [cAy]{.trn}.

These long semi-vowels are rare and may only occur at the end of a sentence.

### Doubled letters

It is possible for a letter to have a round-mark followed by the same letter with a short or long vowel. For example, [qattala]{.trn}. Here the first [t]{.trn} has a round-mark and the second [t]{.trn} has an [a]{.trn} mark. We would expect this word to be written in Arabic like this [قَتْتَلَ]{.ar}. But in fact it is written with one [ت]{.ar} with a tiny doubling mark [ّ ]{.ar} above it along with the vowel mark, like this [قَتَّلَ]{.ar}. When pronouncing this word, stop at and stress the doubled letter [qattala]{.trn} and make sure it does not sound like the undoubled letter in [قَتَلَ]{.ar} [qatala]{.trn}. Examples with doubled letters:

+ [كَبَّرَ]{.ar} [kabbara]{.trn}
+ [كَبِّرْ]{.ar} [kabbir]{.trn}
+ [حَدُّهُ]{.ar} [Hadduhu]{.trn}
+ [فَعَّالَ]{.ar} [faeeAla]{.trn}
+ [سِكِّينُ]{.ar} [sikkInu]{.trn}. Note that the [i]{.trn}-mark is below the doubling mark but above the letter [ك]{.ar}. This is the most common way to write this, although having the [i]{.trn}-mark below the letter is also sometimes done as well. (In this case, the doubling mark will still be above the letter.)
+ [سَفُّودُ]{.ar} [saffUdu]{.trn}
+ [ضَالِّينَ]{.ar} [DAllIna]{.trn}
+ [مُزَّمِّلُ]{.ar} [muzzammilu]{.trn}

## Special characters

### Looped taa [ة]{.ar}

Looped taa [ة]{.ar} is a special letter which is merged from two letters of the alphabet. It is a taa [ت]{.ar} but it is written a [ه]{.ar} with two dots above it. Looped taa [ة]{.ar} is pronounced exactly as a [ت]{.ar} except if it is at the end of a sentence in which case it is pronounced as a [ه]{.ar}. It only occurs at the end of a word so it has only an end form and an isolated form (in case the letter before it is a no-end-join letter). 

Examples:


### Hamza

The letter [hamzah]{.trn} [ء]{.ar} is not traditionally considered part of the alphabet but it is very much a letter. Its called a “glottal stop” in English. You can hear it if you say “a apple” quickly. When you stop your breath momentarily between “a” and “apple” that is the sound that hamza makes. In fact it also occurs in the beginning of “a apple” when you start to say “a”. In transcribing we will represent it with ‘. Note that this is different than out representation for ع which is `.

In Arabic writing hamza can either be written "standalone" by itself [ء]{.ar} or it can be written “seated” on [alif]{.trn} [أ]{.ar} (also under [alif]{.trn} [إ]{.ar}), [wAw]{.trn} [ؤ]{.ar} or [yAE]{.trn} [ئ]{.ar}. When [hamzah]{.trn} is seated the [alif]{.trn}, [wAw]{.trn}, and [yAE]{.trn} are called the seat of the hamza. 

Here are some of notes about writing [hamzah]{.trn} when standalone and seated:

+ When [hamzah]{.trn} is standalone [ء]{.ar}, if it is followed by an [alif]{.trn}: [ءا]{.ar}, it written as [آ]{.ar} if [hamzah]{.trn} is in the beginning or middle of a word. If it is at the end of a word, or was originally at the end but there is a suffix attached to it, then is written as [ءَا]{.ar}.

+ If standalone [hamzah]{.trn} comes between two letters that are joined, then it is written above the line that joins them, for example: [خَطِيـَٔةٌ]{.ar} [xaTIEatun]{.trn}. In this word, the [yAE]{.trn} [ي]{.ar} joins to the looped [tAE]{.trn} [ة]{.ar}. So the standalone [hamzah]{.trn} is written above the joining line like so [ــٔـ]{.ar}.

+ When [hamzah]{.trn} is seated on [alif]{.trn}, if it carries an [i]{.trn}-mark or [in]{.trn}-mark, it is written below the [alif]{.trn}: [إِ]{.ar}, [إٍ]{.ar}. Otherwise, it is written above the [alif]{.trn}: [أَ، أً، أُ، أٌ، أْ]{.ar}.

+ When [hamzah]{.trn} is seated on [yAE]{.trn} [ئ]{.ar} the dots of the [yAE]{.trn} are no longer written. Here's how it will appear in different positions:

  Isolated | End | Middle | Beginnning
  :--------|:----|:-------|:----------
  [ئ]{.ar}| [ـئ]{.ar}| [ـئـ]{.ar}| [ئـ]{.ar}

So how do we know when to write [hamzah]{.trn} standalone and when seated? And how do we choose between its three different seats? There are a set of rules that we need to follow in order to correctly write [hamza]{.trn}:

1. If [hamzah]{.trn} occurs in the beginning of a word:
   a. If the [hamzah]{.trn} carries a long-[A]{.trn} vowel, it is written standalone followed by an [alif]{.trn} and written as [آ]{.ar}, for example [آمَنَ]{.ar} [EAmana]{.trn}.
   b. If the [hamzah]{.trn} carries any other vowel, it is written seated on an [alif]{.trn}, and is marked with the appropriated vowel mark, for example [أَسْلَمَ]{.ar} [Easlama]{.trn}, [أُرِيدُ]{.ar} [EurIdu]{.trn}, [إِسْلَامُ]{.ar} [EislAmu]{.trn}, [إِيمَانُ]{.ar} [EImAnu]{.trn}, [أُوخِذَ]{.ar} [EUxipa]{.trn}. 

2. If [hamzah]{.trn} occurs in the middle of a word:
   a. If there is a long vowel or semi-vowel before the [hamzah]{.trn}:
      i. If the [hamzah]{.trn} is after a long-[I]{.trn} vowel or [ay]{.trn} semi-vowel, then the hamzah will be written standalone. Examples:  
      [هَيْـَٔةُ]{.ar} [hayEatu]{.trn}, [خَطِيـَٔةُ]{.ar} [xaTIEatu]{.trn}  
      [مُسِيـُٔونَ]{.ar} [musIEUna]{.trn}, [مُسِيـَٔانِ]{.ar} [musIEAni]{.trn}, [مُسِيـِٔينَ]{.ar} [musIEIna]{.trn}, [مُسِيـَٔيْنِ]{.ar} [musIEayni]{.trn}  
      [شَيْـُٔهُ]{.ar} [cayEuhu]{.trn}, [شَيْـَٔهُ]{.ar} [cayEahu]{.trn}, [شَيْـِٔهِ]{.ar} [cayEihi]{.trn}, [شَيْـَٔانِ]{.ar} [cayEAni]{.trn}, [شَيْـَٔيْنِ]{.ar} [cayEayni]{.trn}  
      [مَجِيـُٔهُ]{.ar} [majIEuhu]{.trn}, [مَجِيـَٔهُ]{.ar} [majIEahu]{.trn}, [مَجِيـِٔهِ]{.ar} [majIEihi]{.trn}  
      ii. If the [hamzah]{.trn} is after a long-[U]{.trn} vowel or [aw]{.trn} semi-vowel, then:  
          If the [hamzah]{.trn} has an [i]{.trn}-mark it is written seated on [yAE]{.trn}. Examples:
          [سُوئِهِ]{.ar} [sUEihi]{.trn}, [ضَوْئِهِ]{.ar} [DawEihi]{.trn}  
          Otherwise, the [hamzah]{.trn} is written standalone. Examples:
          [سُوءَه]{.ar} [sUEahu]{.trn}, [سُوءَانِ]{.ar} [sUEAni]{.trn},  [تَوْءَمُ]{.ar} [tawEamu]{.trn}, [ضَوْءَهُ]{.ar} [DawEahu]{.trn},  [ضَوْءَانِ]{.ar} [DawEAni]{.trn},  [سُوءُهُ]{.ar} [sUEuhu]{.trn},  [يَسُوءُونَ]{.ar} [yasUEUna]{.trn}  
      iii. If the [hamzah]{.trn} is after a long [A]{.trn}-vowel, then:
           If the [hamzah]{.trn} has an [i]{.trn}-mark it is written seated on [yAE]{.trn}. Example:
           [سَائِلُ]{.ar} [sAEilu]{.trn}.  
           If the [hamzah]{.trn} has an [u]{.trn}-mark it is written seated on [wAw]{.trn}. Example: [تَسَاؤُلُ]{.ar} [tasAEulu]{.trn}.  
           Otherwise, when the [hamzah]{.trn} has an [a]{.trn}-mark, it is written standalone. Example: [سَاءَلَ]{.ar} [sAEala]{.trn}, [قِرَاءَاتُ]{.ar} [qirAEAtu]{.trn}.
   b. If the letter before the [hamzah]{.trn} has a null-vowel mark and is not [wAw]{.trn} or [yAE]{.trn}, then:
      i. If the [hamzah]{.trn} was originally at the end of the word, but a suffix has been attached to the word, then the [hamzah]{.trn} will be written standalone. Examples: [عِبْـَٔانِ]{.ar} [eibEAni]{.trn}, [عِبْـَٔييْنِ]{.ar} [eibEayni]{.trn}, [بُطْـَٔهُ]{.ar} [buTEahu]{.trn}, [بُطْـُٔهُ]{.ar} [buTEuhu]{.trn}, [بُطْـِٔهِ]{.ar} [buTEihi]{.trn}. ([انِ]{.ar}, [يْنِ]{.ar}, [هُ]{.ar}, and [هِ]{.ar} are suffixes).
      ii. Otherwise, if the [hamzah]{.trn} is originally at the middle of the word, then it is written seated on [wAw]{.trn} if it has an [u]{.trn}-mark and [yAE]{.trn} if it has an [i]{.trn}-mark. Examples: [مَسْؤُولُ]{.ar} [masEUlu]{.trn}, [تَرْئِيسُ]{.ar} [tarEIsu]{.trn}.  
          If the [hamzah]{.trn} has an [a]{.trn} mark then:  
          If it is followed by a long-[A]{.trn} vowel represented by an [alif]{.trn}, the [hamzah]{.trn} is standalone followed by the [alif]{.trn} and the combination is written as [آ]{.ar}. Example: [مِرْآةُ]{.ar} [mirEAtu]{.trn}, [ظَمْآنُ]{.ar} [PamEAnu]{.trn}.  
          Otherwise, if there is no [alif]{.trn} after the [hamzah]{.trn}, the [hamzah]{.trn} is written seated on [alif]{.trn}. Examples: [مَسْأَلَةُ]{.ar} [masEalatu]{.trn}, , [مَرْأَةُ]{.ar} [marEatu]{.trn}.
   c. If the [hamzah]{.trn} has a null-vowel mark, then it is written seated on [alif]{.trn} if it has an [a]{.trn}-mark, [wAw]{.trn} if it has an [u]{.trn}-mark, and [yAE]{.trn} if it has an [i]{.trn}-mark. Examples: [كَأْسُ]{.ar} [kaEsu]{.trn}, [بِئْسَ]{.ar} [biEsa]{.trn}, [سُؤْلَكَ]{.ar} [suElaka]{.trn}.
   d. Otherwise, only if the above conditions are not satisfied, then compare the vowel marks of the [hamzah]{.trn} and the letter before it:
      i. If either vowel mark is an [i]{.trn}-mark then the [hamzah]{.trn} will be written on a [yAE]{.trn}. Examples: [سُئِلَ]{.ar} [suEila]{.trn}, [يَئِسَ]{.ar} [yaEisa]{.trn}, [مُتَّكِئِينَ]{.ar} [muttakiEIna]{.trn}.
      ii. If neither vowel mark is an [i]{.trn}-mark, and at least one of the vowel marks is a [u]{.trn}-mark, then the [hamzah]{.trn} will be written on a [wAw]{.trn}. Examples: [سُؤَالُ]{.ar} [suEAlu]{.trn}, [رُؤُوسُ]{.ar} [ruEUsu]{.trn}, [لُؤَيْمُ]{.ar} [luEaymu]{.trn}.
      iii. Otherwise, if both of the vowel marks are [a]{.trn}-marks, then:  
           If the [hamzah]{.trn} is followed by a long-[A]{.trn} vowel represented by an [alif]{.trn}, the [hamzah]{.trn} is written standalone. Examples: [شَنَآنُ]{.ar} [canaEAnu]{.trn}.
           Otherwise the [hamzah]{.trn} will be written on an [alif]{.trn}. Examples: [سَأَلَ]{.ar} [saEala]{.trn}, [رَأَىٰ]{.ar} [raEA]{.trn}.

3. If [hamzah]{.trn} is at the end of a word, disregard the vowel mark on it and consider only the letter before the [hamzah]{.trn}.
   a. If there is a long vowel ([A]{.trn}, [I]{.trn}, [U]{.trn}) or a semi-vowel ([aw]{.trn}, [ay]{.trn}) before it then the [hamzah]{.trn} will be written standalone. Examples: [دُعَاءُ]{.ar} [dueAEu]{.trn}, [سُوءُ]{.ar} [sUEu]{.trn}, [جِيءَ]{.ar} [jIEa]{.trn}, [ضَوْءَ]{.ar} [DawEa]{.trn}, [شَيْءَ]{.ar} [cayEa]{.trn}.
   b. Otherwise, if the previous letter has a null-vowel mark, the [hamzah]{.trn} will again be standalone. Examples: [بُطْءُ]{.ar} [buTEu]{.trn}, [عِبْءُ]{.ar} [eibEu]{.trn}, [شَطْءُ]{.ar} [caTEu]{.trn}
   c. Otherwise, if the previous letter has an:
      i. [a]{.trn}-mark, the [hamzah]{.trn} is written seated on [alif]{.trn}. Example [يَهْدَأُ]{.ar} [yahdaEu]{.trn}, [مُبْتَدَأُ]{.ar} [mubtadaEu]{.trn}.
      ii. [i]{.trn}-mark, the [hamzah]{.trn} is written seated on [yAE]{.trn}. Example [يُهَدِّئُ]{.ar} [yuhaddiEu]{.trn}.
      iii. [u]{.trn}-mark, the [hamzah]{.trn} is written seated on [wAw]{.trn}.
     
   , the [hamzah]{.trn} is written seated on [alif]{.trn} if the vowel before is [wAw]{.trn} if it has an [u]{.trn}-mark and [yAE]{.trn} if it has an [i]{.trn}-mark 
      i. If the [hamzah]{.trn} has an [a]{.trn}-mark, the [hamzah]{.trn} will be written standalone. For example, [تَسَاءَلَ]{.ar} [tasAEala]{.trn}, [خَطِيـَٔتَهُ]{.ar} [xaTIEatahu]{.trn}, [سُوءَهُ]{.ar} [sUEahu]{.trn} [تَوْءَمْ]{.ar} [tawEam]{.trn}, [هَيْـَةُ]{.ar} [hayEatu]{.trn}. If the [hamzah]{.trn} is followed by a long [A]{.trn} vowel, the standalone [hamzah]{.trn} will followed by an [alif]{.trn} [ءَا]{.ar}. This time it will not be written as [آ]{.ar}. For example, [خَطِيـَٔاتُ]{.ar} [xaTIEAtu]{.trn}, [شَيْـَٔانِ]{.ar} [cayEAni]{.trn}, [قِرَاءَاتُ]{.ar} [qirAEAtu]{.trn}, [سُوءَانِ]{.ar} [sUEAni]{.trn}.
      ii. If the [hamzah]{.trn} has an [i]{.trn}-mark, the [hamzah]{.trn} will be written seated on a [yAE]{.trn}. For example, [سَائِلُ]{.ar} [sAEilu]{.trn}, 

 If a prefix is attached to the beginning of the word, it will not affect the writing of the [hamzah]{.trn}. For e
 If hamzah is at end of word and standalone, and suffix is added, the hamza remains standalone if the letter before the hamzah joins to the letter after the hamzah. Otherwise, the hamza may become seated depending on the rule for middle hamza

// maybe: If hamza follow by alif ءا comes in the beginning of a word, it is written as آ instead. 
// In hamza spelling rules also explain a-tanween

FIXME: explain all rules now but mention don't need to memorize yet. Can refer bak to this page.

But whether it appears by itself ء or with a seat the pronunciation is the same. The choice of whether to write it by itself or with a “seat” is only a spelling rule. However this rule is quite complicated so we will explain it in a later chapter. For now we will only give the rule for initial hamza:

When hamza occurs at the beginning of a word it will be written as:
As أَ with short-a, e.g., أَحْمَدُ ‘aHmadu
As أُ with short-u, eg
As إِ with short-i
As آ with long-a, eg آنَسْتُ ‘aanastu. Note that instead of a hamza (dotted circle with hamza) we wrote a madda symbol (dotted circle with madda).
As أُو with long-a, eg أُوجِرَ ‘oojira
As إِي with a long-i eg إِي

Besides this we will limit the words used with hamza until we explain the rules fully so hopefully it will not be too confusing.

##### Exercises

#### Connecting hamza

Some words in arabic begin with a sukoon. When this occurs a connecting hamza (written as a tiny صـ on an alif) is put before it. If this word comes in the beginning of the sentence the connecting alif is pronounced as a hamza. Otherwise this connecting hamza is not pronounced and the word is connected to the previous word in pronunciation.

In this tutorial we will transcribe the connecting hamza with a hyphen - when it is in the middle of a sentence.

Examples
‘alwaladu sa3eedun [ٱَلْوَلَدُ سَعِيدٌ]{.ar}
Daraba -lwaladu ‘aHmada NOT Daraba ‘alwaladu aHmada

When one word ends in a long vowel and the next word begins with a connecting hamza, the long vowel becomes a short vowel in pronunciation (but in writing the long vowel’s letter is retained.) For example

‘Ilaa masjidin BUT Dhahaba ‘ila -lmasjidi NOT dhahaba ‘ilaa -lmasjidi NOR ‘ilaa ‘almasjidi
‘Illaa zaydan BUT ‘Illa -lmuSalleena NOT `illaa -lmuSalleena NOR ‘illaa ‘almuSalleena.
Fee baytin BUT Fi -lbayt NOT fee -lbayt NOR fee ‘albayt
Dhoo sa3atin BUT Dhu -lqarnayni NOT dhoo -lqarnayni NOR dhoo ‘alqarnayni 

Sometimes a connecting hamza can occurs after an n-mark. In these cases will will pronounce the n-mark and follow it with a short-i:

Zaydani -lkabeeru NOT zaydan ‘alkabeera

## Joining [lAm]{.trn} and [Ealif]{.trn}

When the letter [Ealif]{.trn} follows [lAm]{.trn} we would expect them to be joined like this [ل+ا]{.ar} &rarr; [لـا]{.ar}. But actually, they are joined in a special way 

[ل+ا]{.ar} &rarr; [لا]{.ar}

If we add [hamzah]{.trn} to the [Ealif]{.trn} or add another [lAm]{.trn}, here is how they will be written and pronounced in words:

Disjoint | Joined | Transcription
:--------|:-------|:-------------
[لَ + ا]{.ar}| [لَا]{.ar}| [lA]{.trn}
[ٱَ + لْ + إِ + يمَانُ]{.ar}| [ٱَلْإِيمَانُ]{.ar}| [alEImAnu]{.trn}
[ٱَ + لْ + آ + فَةُ]{.ar}| [ٱَلْآفَةُ]{.ar}| [alEAfatu]{.trn}
[أَ + لَّ + ا]{.ar}| [أَلَّا]{.ar}| [EallA]{.trn}
[لِ + لِ + ٱ + نْتِقَالِ]{.ar}| [لِلِٱنْتِقَالِ]{.ar}| [li li-ntiqAli]{.trn}

## [n]{.trn}-marks

We will now show three new pronunciation marks, collectively called [n]{.trn}-marks.

1. [an]{.trn}-mark [ ً]{.ar}
1. [in]{.trn}-mark [ ٍ]{.ar}
1. [un]{.trn}-mark [ ٌ]{.ar}

These [n]{.trn}-marks may only occur on a letter at the end of a word. They are pronounced as a short vowel ([a]{.trn}, [i]{.trn}, or [u]{.trn}) followed by an [n]{.trn}. 

As a spelling rule, if a word ends with an [an]{.trn} mark, we will generally add a silent [Ealif]{.trn} after it, for example [سَالِمًا]{.ar} [sAliman]{.trn}. This happens for all words except if the word ends with one of these three endings:

1. a looped [tAE]{.trn} [ة]{.ar}, for example [غَاضِبَةً]{.ar} [gADibatan]{.trn}, not [غَاضِبَةًا]{.ar} or [غَاضِبَتًا]{.ar}.

2. [Ealif]{.trn} followed by an unseated [hamzah]{.trn} [اء]{.ar}, for example [دَاءً]{.ar} [dAEan]{.trn} not [دَاءًا]{.ar}.

3. A [hamzah]{.trn} seated on an [Ealif]{.trn} [أ]{.ar}, for example [مُبْتَدَأً]{.ar} [mubtadaEan]{.trn} not [مُبْتَدَأًا]{.ar}.

Here are some examples of words with [n]{.trn}-marks:

+ [سَعْدٌ]{.ar}  [saedun]{.trn}
+ [ضَرْبًا]{.ar} [Darban]{.trn}
+ [قَاضٍ]{.ar}  [qADin]{.trn}
+ [سَعَةً]{.ar}  [saeatan]{.trn}
+ [دُعَاءً]{.ar} [dueAEan]{.trn}
+ [ٱِمْرَأً]{.ar} [imraEan]{.trn}
+ [شَيْئًا]{.ar} [cayEan]{.trn}
+ [سُوءًا]{.ar} [sUEan]{.trn}
+ [غَبَنٌ]{.ar} [gabanun]{.trn}

## Pausing at the end of a sentence.

## Difference between Quranic script and Standard Arabic script.

In printed Qurans the script will differ somewhat from the one we’ve shown here, especially in vowels. The reasons for this are beyond the scope of this tutorial. But here we’ll just give a few examples and note that these differences are only found in printed Qurans. For standard written Arabic, the script explained in this chapter is used.

إبرهم ابراهم 'ibraheem
ربو ribaa
Madda ءانست aanastu حآفين Haaffeena


The script and spelling rules presented so far show how Standard Arabic is written in regular books and other written material.
The Quran is also written in Arabic but due to reasons that beyond the scope of this tuorial, it’s script often differs slightly from Standard Arabic’s. For example
Madda
Dagger alif on waw

